---
title: "Other explorative analysis"
author: "Mikko Arvas"
date: "19.7.2021"
output: pdf_document
---


# Summary

Various explorations for the three Finnish Cohorts (FinDonor, FINRISK97, Health2000). 

```{r setup, , message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
#library(GGally)
library(ggfortify)
#library(psych)
#library(knitr)
#library(lubridate)
#library(ggbeeswarm)

# echo "rmarkdown::render('explore_data.Rmd', clean=TRUE,output_format='pdf_document',output_file='../results/explore_data_FIN.pdf')" | R --slave

```

# Data loading and preparation

```{r}
#Get the non filtered data produced by low_ferritin_data.Rmd
load("../results/summary_all_cohorts.rdata")
```


# CRP

```{r}

#Mark anemic

summary_all_cohorts <- summary_all_cohorts %>% 
  filter(!is.na(Hb)) %>% 
  filter(!is.na(CRP)) %>%
  filter(!is.na(Ferritin)) %>% 
  mutate(
    CRP=CRPori,
    #Region = as.factor(Region),
    Sex=as.factor(Sex),
    Menstruation=as.factor(Menstruation),
  anemic = case_when(
    Hb < 135 & Sex == 'Men' ~ TRUE,
    Hb < 118 & Sex == 'Women' ~ TRUE,
    TRUE ~ FALSE
  ),
  CRP = case_when(
    CRP == 0 ~ 0.001,
    TRUE ~ CRP
  )
)

summary_all_cohorts %>% 
  group_by(anemic,Sex) %>% 
  summarise(
    maxHb=max(Hb),
    n = n()
    )

```

## CRP in anemic men only

```{r}
temp <- summary_all_cohorts %>%  filter(!is.na(anemic)) %>% filter(Sex == 'Men') %>% filter(anemic == TRUE)

p <- ggplot(temp)
p <- p + geom_point(aes(x=CRP, y=Ferritin,col=Age))
p <- p + scale_y_log10() + scale_x_log10() 
p <- p + facet_wrap(~Cohort)
p <- p + geom_hline(yintercept =15, col="red") + geom_hline(yintercept =30, col="red")
p <- p + theme(axis.text.x=element_text(angle=90, hjust=1))
p

```

## Ferritin vs CRP in all cohorts and groups 

```{r}
temp <- summary_all_cohorts %>%  filter(!is.na(anemic)) 
temp2 <- summary_all_cohorts %>%  filter(!is.na(anemic)) %>% filter(anemic == TRUE)


p <- ggplot(temp)
p <- p + geom_point(aes(x=CRP, y=Ferritin, col=Age),alpha=0.5)
p <- p + geom_point(aes(x=CRP, y=Ferritin),data=temp2,shape=1,col="green")
p <- p + geom_smooth(aes(x=CRP, y=Ferritin),data=temp2,col="green",method = "lm")
p <- p + scale_y_log10() + scale_x_log10() 
p <- p + facet_grid(Group~Cohort)
p <- p + geom_hline(yintercept =15, col="red",linetype=2) + geom_hline(yintercept =30, col="red",linetype=2)
p <- p + geom_vline(xintercept =2.9, col="red",linetype=2)
p <- p + geom_text(aes(x=0.01,y=15,label='15'))
p <- p + geom_text(aes(x=0.03,y=30,label='30'))
p <- p + geom_text(aes(x=2.9,y=1,label='2.9'))
p <- p + theme(axis.text.x=element_text(angle=90, hjust=1))
p <- p + geom_smooth(aes(x=CRP, y=Ferritin), col="black",alpha=0.5)
p <- p + xlab("CRP mg/l") + ylab("Ferritiini ug/l")
p


```

```{r}
ggsave2(filename="../results/ferritin_vs_crp_with_anemia.pdf")
ggsave2(filename="../results/ferritin_vs_crp_with_anemia.png")
```

# How do CRP==0 look?

```{r}
CRP_0_true <- summary_all_cohorts %>%
  filter(Cohort =='HEALTH00') %>% 
  mutate(
    CRP_0 = case_when(
      CRP == 0.001 ~ TRUE,
      TRUE ~ FALSE
    )
  ) %>% filter(CRP_0==TRUE) %>% 
  select(-CRP_0,
         -ID,
         -TwoYearsFromStartCount_FB,
         -DaysToPreviousFB,
        # -IronComplienceNumeric,
         -Cohort,
         #-RedMeat_n,
         -CRPori
         )

CRP_0_false <- summary_all_cohorts %>%
  filter(Cohort =='HEALTH00') %>% 
  mutate(
    CRP_0 = case_when(
      CRP == 0.001 ~ TRUE,
      TRUE ~ FALSE
    )
  ) %>% filter(CRP_0==FALSE) %>% 
  select(-CRP_0,
         -ID,
         -TwoYearsFromStartCount_FB,
         -DaysToPreviousFB,
        # -IronComplienceNumeric,
         -Cohort,
        # -RedMeat_n,
         -CRPori
         )
  

summary(CRP_0_true %>% filter(Sex=="Men"))

```

```{r}
summary(CRP_0_false %>% filter(Sex=="Men"))
```

No obvious differences. Although as expected BMI slightly smaller in the CRP_0 group.

```{r}
summary(CRP_0_true %>% filter(Sex=="Women"))
```

```{r}
summary(CRP_0_false %>% filter(Sex=="Women"))
```
No obvious differences. Although as expected BMI slightly smaller in the CRP_0 group, and ferritin and age too. Anyhow the CRP=0 data appears real.


# Donation history for press


```{r}
# load FinDonor demographic data
load("../data/r02ds.donorData.PRID.rdata")
donor_demo <- output
rm(output)

tmp <- donor_demo %>% 
  dplyr::select(releaseID,DonationCount_FB) %>% 
  left_join(
    summary_all_cohorts %>% dplyr::select(ID, Group, Sex, Age,TwoYearsFromStartCount_FB,DaysToPreviousFB),
    by = c("releaseID" = "ID")
    ) %>% 
  mutate(
    Group = case_match(Group,
      "Pre_menopause_women" ~ "Pre-menopausal",
      "Post_menopause_women" ~ "Post-menopausal",
      "Men" ~"Men"
    )
  )

dim(tmp)
tmp <- na.omit(tmp)
dim(tmp)

```


```{r}

p <- ggplot(tmp,aes(x=TwoYearsFromStartCount_FB,col="white", fill=Group))
p <- p + stat_bin(binwidth = 1)
p <- p + facet_grid(Group ~.,scales = "free_y")
p <- p + xlab("Count of whole blood donations 2 years \n before FinDonor enrollment visit")
p <- p + ylab("Count of persons")
p <- p +  
scale_color_manual(values = c( "#00BFFF",  "#de2d26", "#ff85a2" ),
                     limits = c( "Men",  "Post-menopausal", "Pre-menopausal" )) +
  scale_fill_manual(values = c( "#00BFFF",  "#de2d26", "#ff85a2" ),
                     limits = c( "Men",  "Post-menopausal", "Pre-menopausal" )) +
theme_classic() +
  theme(legend.position = "none",
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour="black"),
        #axis.title.y = element_blank(),
        axis.title.x = element_text(size = 10),
        axis.text = element_text(size = 8),
        panel.grid.major.y = element_blank())

p
ggplot2::ggsave("../results/figures/donation_history_2y.png",device = "png", height = 12, width = 11, units = "cm" )

```



```{r}

p <- ggplot(tmp,aes(x=DonationCount_FB,col="white", fill=Group))
p <- p + stat_bin(binwidth = 5,)
p <- p + facet_grid(Group ~.,scales = "free_y")
p <- p + xlab("Count of life time whole blood donations \n before FinDonor enrollment visit")
p <- p + ylab("Count of persons")
p <- p +  #scale_x_log2() +
scale_color_manual(values = c( "#00BFFF",  "#de2d26", "#ff85a2" ),
                     limits = c( "Men",  "Post-menopausal", "Pre-menopausal" )) +
  scale_fill_manual(values = c( "#00BFFF",  "#de2d26", "#ff85a2" ),
                     limits = c( "Men",  "Post-menopausal", "Pre-menopausal" )) +
theme_classic() +
  theme(legend.position = "none",
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour="black"),
        #axis.title.y = element_blank(),
        axis.title.x = element_text(size = 10),
        axis.text = element_text(size = 8),
        panel.grid.major.y = element_blank())

p
ggplot2::ggsave("../results/figures/donation_history.png",device = "png", height = 12, width = 11, units = "cm" )

```
