---
title: "Ferritin model prediction"
author: "Mikko Arvas"
date: "25.04.2021"
output: pdf_document

---


# Summary


```{r setup, , message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
# echo "rmarkdown::render('ferritin_model_prediction.Rmd', clean=TRUE,output_format='pdf_document',output_file='../results/ferritin_model_prediction_FIN.pdf')" | R --slave
library(tidyverse)
library(brms)
library(bayesplot)

```


# Data loading and preparation 

Load data

```{r}
load("../data/ID_data_regression_cohorts.rdata")
all_cohorts <- regression_cohorts

all_cohorts <- all_cohorts %>% 
  filter(!is.na(CRP)) %>% 
  filter(! CRP >= 30)
```

# All cohorts together 

## Data
```{r}
# men
## menstruation and childbirth variables removed 
regression_data_men <- all_cohorts %>% 
  filter(Group == "Men") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, Cohort, iron_deficiency, Smoking, iron_complience, Group, Region, Weight, RedMeat_n) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# premenopausal women
regression_data_pre <- all_cohorts %>% 
  filter(Group == "Pre_menopause_women") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, Cohort, iron_deficiency, Smoking, iron_complience, Group, Region, Weight, RedMeat_n, Menstruation, PreviousChildbirth) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# postmenopausal womem 
## menstruation variable removed 
regression_data_post <- all_cohorts %>% 
  filter(Group == "Post_menopause_women") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, Cohort, iron_deficiency, Smoking, iron_complience, Group, Region, Weight, RedMeat_n, Menstruation, PreviousChildbirth) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 


```

# Summaries 


```{r men}
regression_data_men <- regression_data_men %>% select(
  log_ferritin,
  iron_deficiency,
  Age, 
  BMI, 
  Smoking, 
  iron_complience,
  donation_count_2, 
  donation_count,
  log_last_donation, 
  log_CRP,
  #Region,
  Weight,
  #RedMeat_n,
  Cohort
) %>% mutate(
  Cohort = as.factor(Cohort)
)

summary(regression_data_men)

```

THERE SHOULD BE NO NAs AT THIS POINT!!!


```{r pre}
regression_data_pre <- regression_data_pre %>% select(
  log_ferritin,
  iron_deficiency,
  Age, 
  BMI, 
  Smoking, 
  iron_complience,
  donation_count_2, 
  donation_count,
  log_last_donation, 
  log_CRP,
  #Region,
  Weight,
  #RedMeat_n,
  Cohort,
  Menstruation,
  PreviousChildbirth
) %>% mutate(
  Cohort = as.factor(Cohort),
  Menstruation = as.factor(Menstruation)
) 

summary(regression_data_pre)

```

```{r post}
regression_data_post <- regression_data_post %>% select(
  log_ferritin,
  iron_deficiency,
  Age, 
  BMI, 
  Smoking, 
  iron_complience,
  donation_count_2, 
  donation_count,
  log_last_donation, 
  log_CRP,
  #Region,
  Weight,
  #RedMeat_n,
  Cohort,
  #Menstruation,
  PreviousChildbirth
)  %>% mutate(
  Cohort = as.factor(Cohort)
  # Menstruation = as.factor(Menstruation)
)

summary(regression_data_post)

```

# Ferritin models with donor history



```{r men_b}
#men
file <- "../results/bmodels/pred/linear_men_b"

linear_men_b <- brm(log_ferritin ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        iron_complience + 
                        donation_count_2 + 
                        donation_count +
                        log_last_donation + 
                        log_CRP +
                        #Region + 
                        Weight +
                        #RedMeat_n +
                        Cohort, 
                      data = regression_data_men, 
                      #family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(linear_men_b)

```

```{r  pre_b}
# premenopausal women
file <- "../results/bmodels/pred/linear_pre_b"

linear_pre_b <- brm(log_ferritin ~ 
                      Age + 
                        BMI + 
                        Smoking + 
                        iron_complience + 
                        donation_count_2 + 
                        donation_count +
                        log_last_donation + 
                        log_CRP +
                        #Region + 
                        Weight +
                        #RedMeat_n +
                        Cohort +
                        Menstruation +
                        PreviousChildbirth, 
                      data = regression_data_pre,
                      #family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(linear_pre_b)


```

```{r  post_b}
# postmenopausal women
file <- "../results/bmodels/pred/linear_post_b"

linear_post_b <- brm(log_ferritin ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        iron_complience + 
                        donation_count_2 + 
                        donation_count +
                        log_last_donation + 
                        log_CRP +
                        #Region + 
                        Weight +
                        #RedMeat_n +
                        Cohort +
                        #Menstruation +   # will not work as all observations are no_period
                        PreviousChildbirth, 
                     data = regression_data_post, 
                     #family = bernoulli(),
                     file=file, 
                     cores = 4,
                     iter= 10000
)

summary(linear_post_b)
```

## Diagnostics

### Trace

```{r , message=FALSE}
mcmc_plot(linear_men_b, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(linear_pre_b, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(linear_post_b, type = "trace")
```

Chains do not seem to get stuck.

### ACF bars

```{r}
mcmc_plot(linear_men_b, type = "acf_bar")
```


```{r}
mcmc_plot(linear_pre_b, type = "acf_bar")
```


```{r}
mcmc_plot(linear_post_b, type = "acf_bar")
```

Autocorrelation drops nicely.

### Rhat

```{r}
mcmc_plot(linear_men_b, type = "rhat")
```

```{r}
mcmc_plot(linear_pre_b, type = "rhat")
```

```{r}
mcmc_plot(linear_post_b, type = "rhat")
```
https://mc-stan.org/rstan/reference/Rhat.html


"The Rhat function produces R-hat convergence diagnostic, which compares the between- and within-chain estimates for model parameters and other univariate quantities of interest. If chains have not mixed well (ie, the between- and within-chain estimates don't agree), R-hat is larger than 1. We recommend running at least four chains by default and only using the sample if R-hat is less than 1.05."


# Ferritin models with donor history

```{r men_b_nc}
#men
file <- "../results/bmodels/pred/linear_men_b_nc"

linear_men_b_nc <- brm(log_ferritin ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        iron_complience + 
                        # donation_count_2 + 
                        # donation_count +
                        # log_last_donation + 
                        log_CRP +
                        #Region + 
                        Weight +
                        #RedMeat_n +
                        Cohort, 
                      data = regression_data_men, 
                      #family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(linear_men_b_nc)

```

```{r  pre_b_nc}
# premenopausal women
file <- "../results/bmodels/pred/linear_pre_b_nc"

linear_pre_b_nc <- brm(log_ferritin ~ 
                      Age + 
                        BMI + 
                        Smoking + 
                        iron_complience + 
                        # donation_count_2 + 
                        # donation_count +
                        # log_last_donation + 
                        log_CRP +
                        #Region + 
                        Weight +
                        #RedMeat_n +
                        Cohort +
                        Menstruation +
                        PreviousChildbirth, 
                      data = regression_data_pre,
                      #family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(linear_pre_b_nc)


```

```{r  post_b_nc}
# postmenopausal women
file <- "../results/bmodels/pred/linear_post_b_nc"

linear_post_b_nc <- brm(log_ferritin ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        iron_complience + 
                        # donation_count_2 + 
                        # donation_count +
                        # log_last_donation + 
                        log_CRP +
                        #Region + 
                        Weight +
                        #RedMeat_n +
                        Cohort +
                        #Menstruation +   # will not work as all observations are no_period
                        PreviousChildbirth, 
                     data = regression_data_post, 
                     #family = bernoulli(),
                     file=file, 
                     cores = 4,
                     iter= 10000
)

summary(linear_post_b_nc)
```



## Diagnostics

### Trace

```{r , message=FALSE}
mcmc_plot(linear_men_b_nc, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(linear_pre_b_nc, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(linear_post_b_nc, type = "trace")
```

Chains do not seem to get stuck.

### ACF bars

```{r}
mcmc_plot(linear_men_b_nc, type = "acf_bar")
```


```{r}
mcmc_plot(linear_pre_b_nc, type = "acf_bar")
```


```{r}
mcmc_plot(linear_post_b_nc, type = "acf_bar")
```

Autocorrelation drops nicely.

### Rhat

```{r}
mcmc_plot(linear_men_b_nc, type = "rhat")
```

```{r}
mcmc_plot(linear_pre_b_nc, type = "rhat")
```

```{r}
mcmc_plot(linear_post_b_nc, type = "rhat")
```
https://mc-stan.org/rstan/reference/Rhat.html


"The Rhat function produces R-hat convergence diagnostic, which compares the between- and within-chain estimates for model parameters and other univariate quantities of interest. If chains have not mixed well (ie, the between- and within-chain estimates don't agree), R-hat is larger than 1. We recommend running at least four chains by default and only using the sample if R-hat is less than 1.05."


# Predict

```{r}

linear_men_b_p <- predict(linear_men_b)
linear_men_b_p_nc <- predict(linear_men_b_nc)
linear_pre_b_p <- predict(linear_pre_b)
linear_pre_b_p_nc <- predict(linear_pre_b_nc)
linear_post_b_p <- predict(linear_post_b)
linear_post_b_p_nc <- predict(linear_post_b_nc)
```

```{r}

pred_men<- tibble(
    fer_pred = exp(1)^linear_men_b_p[, "Estimate"],
    fer = exp(1)^regression_data_men$log_ferritin,
    group="Men",
    model="History"
)

pred_post<- tibble(
    fer_pred = exp(1)^linear_post_b_p[, "Estimate"],
    fer = exp(1)^regression_data_post$log_ferritin,
    group="Post",
    model="History"
)

pred_pre<- tibble(
    fer_pred = exp(1)^linear_pre_b_p[, "Estimate"],
    fer = exp(1)^regression_data_pre$log_ferritin,
    group="Pre",
    model="History"
)

pred_men_nc<- tibble(
    fer_pred = exp(1)^linear_men_b_p_nc[, "Estimate"],
    fer = exp(1)^regression_data_men$log_ferritin,
    group="Men",
    model="No_history"
)

pred_post_nc<- tibble(
    fer_pred = exp(1)^linear_post_b_p_nc[, "Estimate"],
    fer = exp(1)^regression_data_post$log_ferritin,
    group="Post",
    model="No_history"
)

pred_pre_nc<- tibble(
    fer_pred = exp(1)^linear_pre_b_p_nc[, "Estimate"],
    fer = exp(1)^regression_data_pre$log_ferritin,
    group="Pre",
    model="No_history"
)


data_sim <- bind_rows(
  pred_men,
  pred_post,
  pred_pre,
  pred_men_nc,
  pred_post_nc,
  pred_pre_nc
)


p <- ggplot(data_sim,aes(y=fer_pred,x=fer))
p <- p + geom_point(alpha=0.1)
p <- p + geom_smooth(method = "lm")
p <- p + geom_smooth(method = "gam")
p <- p + facet_grid(model~group)
p <- p + scale_x_continuous(trans='log2') + scale_y_continuous(trans='log2')
p <- p + geom_vline(aes(xintercept=15),color="red")
p <- p + geom_hline(aes(yintercept=15),color="red")
p
```

```{r}
cors <- data_sim %>% 
  group_by(group, model) %>%  
  summarise(
    r2 = round(summary(lm(fer_pred~fer))$r.squared, 2)
    )
cors
```

