---
title: "Ferritin median change simulation NL"
author: "Mikko Arvas"
date: "11.12.2023"
output:
  html_document:
    df_print: paged
---


# Summary

Simulate effect of changing ferritin median and difference between genpop and donors.


```{r setup, , message = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(broom)
library(colorspace)
# echo "rmarkdown::render('ferritin_model_shift_sim.Rmd', clean=TRUE,output_format='pdf_document',output_file='../results/ferritin_model_shift_sim.pdf')" | R --slave
```




# Data loading and preparation 

Load data

```{r}
load("../data/ID_data_regression_cohorts_genpop_combined.rdata")
all_cohorts <- regression_cohorts


```



# All cohorts together 

## Data
```{r}
regression_data_men <- all_cohorts %>% 
  filter(Group == "Men") %>% 
  dplyr::select(ID, log_ferritin, Age, donation_count, log_last_donation, Cohort, Group, blood_vol) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         blood_vol = scale(blood_vol, scale = FALSE)[,1],
         donation_count_2 = donation_count^2
         ) 



```


   *Make fake Dutch data that includes new donors start*

```{r}

# Take male donors and change their history to be as in genpop

fake_new_donor <- regression_data_men %>% 
  filter(Cohort == "FINDONOR")

thl <- regression_data_men %>% 
  filter(Cohort == "THL")

#like genpop
dc <- sample(thl$donation_count,nrow(fake_new_donor),replace = TRUE)
ll <- sample(thl$log_last_donation,nrow(fake_new_donor),replace = TRUE)
dc2 <- sample(thl$donation_count_2,nrow(fake_new_donor),replace = TRUE)
#like genpop but no iron deficient
lf <- sample(thl$log_ferritin[thl$log_ferritin >3.9 ]
             ,nrow(fake_new_donor),replace = TRUE)

fake_new_donor <- fake_new_donor %>% 
  mutate(
    log_ferritin = lf,
    Cohort = "NEWDONOR",
    donation_count =dc  ,
    log_last_donation = ll ,
    donation_count_2 = dc2, 
  )

regression_data_men_nl <- bind_rows(regression_data_men, fake_new_donor)
#instead of this you would just assing
# regression_data_men_nl <- regression_data_men


```


   *Make fake Dutch data that includes new donors stop*




## Test models 



```{r }

linear_men <- lm(log_ferritin ~ 
                        Age + 
                        donation_count_2 + 
                        donation_count +
                        log_last_donation + 
                        blood_vol+
                        Cohort, 
                      data = regression_data_men_nl, 
)

summary(linear_men)

```


```{r}
confint(linear_men)
```




```{r}

p <- ggplot(data=regression_data_men_nl)
p <- p +  geom_histogram(aes(x=log_ferritin))
p <- p +  facet_wrap(~Cohort)
p
```

```{r}
regression_data_men_nl %>% 
  group_by(Cohort) %>% 
  summarise(
    med_fer = exp(median(log_ferritin))
    )

```


# Test the effect of changing the median of the cohorts


```{r}

# Reduce genpop to be like Finnish pre-menopausal
# in Blood donors NL median = 55
# in Gen pop NL median = 134

#
bdf <- 26 /55 # 
gpf <-  26 /134 
ndf <-   26 / 127 

###Change the Cohort variable values to match NL###
addfer <- tibble(Cohort = c("FINDONOR","THL","NEWDONOR"),
       ferfac = c(bdf,gpf,ndf)
       )

regression_data_men_as_fi <- regression_data_men_nl %>%
  right_join(addfer,by=c("Cohort"="Cohort")) %>% 
  mutate(
    log_ferritin = log(exp(log_ferritin) *  ferfac)
      )
    
  
  
regression_data_men_as_fi %>% 
  group_by(Cohort) %>% 
  summarise(
    med_fer = exp(median(log_ferritin))
    )

```


```{r}

tmp <- bind_cols( 
  regression_data_men_nl %>% select(log_ferritin) %>% rename(original = log_ferritin),
    regression_data_men_as_fi %>% select(log_ferritin, Cohort) %>% rename(like_dutch_men = log_ferritin),
  )

p <- ggplot(tmp)
p <- p + geom_point(aes(y=exp(like_dutch_men),x=exp(original),color=Cohort))
p
```



```{r}
tmp <- bind_rows(
  regression_data_men_nl %>% select(Cohort,log_ferritin) %>%  mutate(set="original"),
  regression_data_men_as_fi %>% select(Cohort,log_ferritin) %>% mutate(set="like_finn_pre")
)

tmp_sum <- tmp %>%  group_by(Cohort,set) %>% 
  summarise(
    median = median(log_ferritin)
  )

p <- ggplot(data=tmp)
p <- p +  geom_histogram(aes(x=log_ferritin),alpha=0.5,binwidth = 0.1)
p <- p + geom_vline(data=tmp_sum,aes(xintercept=median))
p <- p + facet_grid(Cohort~set)
p
```

# Forest plot with new donors 


```{r  }
#

linear_men_NL <- lm(log_ferritin ~ 
                      Age + 
                        donation_count_2 + 
                        donation_count +
                        log_last_donation + 
                        Cohort +
                        blood_vol,
                      data = regression_data_men_nl,
  )

summary(linear_men_NL)


```

```{r  }
#

linear_men_as_fi <- lm(log_ferritin ~ 
                      Age + 
                        donation_count_2 + 
                        donation_count +
                        log_last_donation + 
                        Cohort +
                        blood_vol,
                      data = regression_data_men_as_fi,
  )

summary(linear_men_as_fi)


```




```{r}
confint(linear_men_NL)
```

```{r}
confint(linear_men_as_fi)
```




```{r}
regressor_values <- c(
 "Age" = "Age (5 years)",
   "log_last_donation" = "Days since last donation (log, /log(2))",
    "donation_count_2" = "Donations in last 2 years (quadratic)",
    "donation_count" = "Donations in last 2 years",
 
 ###Change the Cohort variable values to match NL###
 
    "CohortTHL" = "General population",
    "CohortNEWDONOR" = "New donors"
)


intervals_pre <- bind_cols(
  regressor= rownames(confint(linear_men_as_fi)),
  confint(linear_men_as_fi),
  m=summary(linear_men_as_fi)$coefficients[,"Estimate"],
  set = "as_pre_fi"
  ) %>% 
  rename(ll = `2.5 %`,hh=`97.5 %`)


intervals_men_NL <- bind_cols(
  regressor= rownames(confint(linear_men_NL)),
  confint(linear_men_NL),
  m=summary(linear_men_NL)$coefficients[,"Estimate"],
  set = "original"
  ) %>% 
  rename(ll = `2.5 %`,hh=`97.5 %`)

intervals_combined <- bind_rows(intervals_pre,intervals_men_NL) %>% 
  filter(! (regressor == '(Intercept)' )) %>% 
  mutate(regressor = plyr::revalue(regressor, regressor_values)) 


```


## Plot

```{r , fig1, fig.height = 10, fig.width = 15 ,warning=FALSE}

orlower <- -1
orupper <- 1

# Truncate ORs for plotting
plotdata_wnd <- intervals_combined %>% 
  mutate(
    ll = case_when(
      ll < orlower ~ orlower,
      TRUE ~ ll
    ),
    hh = case_when(
      hh > orupper ~ orupper,
      TRUE ~ hh
    )
  ) %>% mutate(is.sig = ll < 0 & hh < 0 | ll > 0 & hh > 0)


pos <- position_nudge(y=as.numeric(as.factor(plotdata_wnd$set))/5 -0.1) 

p <- ggplot(plotdata_wnd, aes(x = m, y = regressor, color = set)) + 
  geom_point(position = pos,
             aes(shape = is.sig), 
             size=4) +
  geom_linerange(aes(xmin = ll, xmax = hh), 
                 position = pos,
                 size = 1, 
                 linetype = 1
                  ) +
    scale_shape_manual(values = c(1,16)) +
  xlim(orlower,orupper)  +   
  geom_vline(xintercept = 0 , color = "grey", linetype = "dotted",
             size = 1) +
  theme(axis.title = element_text(size = 14),
        strip.text =  element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.background = element_rect(fill = "white"),
        strip.background.x = element_blank(),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14)
  ) +
  guides(shape = "none") +
  theme_classic() +
  xlab("Coefficient") +
  theme(text = element_text(size = 20)) +
  theme(axis.title.y = element_blank())

p

```


# Forest plot with out new donors


```{r  }
#

linear_men_NL <- lm(log_ferritin ~ 
                      Age + 
                        donation_count_2 + 
                        donation_count +
                        log_last_donation + 
                        Cohort +
                        blood_vol,
###Change the Cohort variable values to match NL###
                      data = regression_data_men_nl %>% filter(Cohort!="NEWDONOR"),
  )

summary(linear_men_NL)


```

```{r }
#

linear_men_as_fi <- lm(log_ferritin ~ 
                      Age + 
                        donation_count_2 + 
                        donation_count +
                        log_last_donation + 
                        Cohort +
                        blood_vol,
###Change the Cohort variable values to match NL###
                      data = regression_data_men_as_fi %>% filter(Cohort!="NEWDONOR"),
  )

summary(linear_men_as_fi)


```




```{r}
confint(linear_men_NL)
```

```{r}
confint(linear_men_as_fi)
```




```{r}
regressor_values <- c(
 "Age" = "Age (5 years)",
   "log_last_donation" = "Days since last donation (log, /log(2))",
    "donation_count_2" = "Donations in last 2 years (quadratic)",
    "donation_count" = "Donations in last 2 years",
 
###Change the Cohort variable values to match NL###

    "CohortTHL" = "General population"
   # "CohortNEWDONOR" = "New donors"
)


intervals_pre <- bind_cols(
  regressor= rownames(confint(linear_men_as_fi)),
  confint(linear_men_as_fi),
  m=summary(linear_men_as_fi)$coefficients[,"Estimate"],
  set = "as_pre_fi"
  ) %>% 
  rename(ll = `2.5 %`,hh=`97.5 %`)


intervals_men_NL <- bind_cols(
  regressor= rownames(confint(linear_men_NL)),
  confint(linear_men_NL),
  m=summary(linear_men_NL)$coefficients[,"Estimate"],
  set = "original"
  ) %>% 
  rename(ll = `2.5 %`,hh=`97.5 %`)

intervals_combined <- bind_rows(intervals_pre,intervals_men_NL) %>% 
  filter(! (regressor == '(Intercept)' )) %>% 
  mutate(regressor = plyr::revalue(regressor, regressor_values)) 


```


## Plot

```{r , fig2, fig.height = 10, fig.width = 15 ,warning=FALSE}

orlower <- -1
orupper <- 1

# Truncate ORs for plotting
plotdata_wond <- intervals_combined %>% 
  mutate(
    ll = case_when(
      ll < orlower ~ orlower,
      TRUE ~ ll
    ),
    hh = case_when(
      hh > orupper ~ orupper,
      TRUE ~ hh
    )
  ) %>% mutate(is.sig = ll < 0 & hh < 0 | ll > 0 & hh > 0)

pos <- position_nudge(y=as.numeric(as.factor(plotdata_wond$set))/5 -0.1) 

p <- ggplot(plotdata_wond, aes(x = m, y = regressor, color = set)) + 
  geom_point(position = pos,
             aes(shape = is.sig), 
             size=4) +
  geom_linerange(aes(xmin = ll, xmax = hh), 
                 position = pos,
                 size = 1, 
                 linetype = 1
                  ) +
  
  scale_shape_manual(values = c(1,16)) +
  xlim(orlower,orupper)  +   
  geom_vline(xintercept = 0 , color = "grey", linetype = "dotted",
             size = 1) +
  theme(axis.title = element_text(size = 14),
        strip.text =  element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.background = element_rect(fill = "white"),
        strip.background.x = element_blank(),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14)
  ) +
  guides(shape = "none") +
  theme_classic() +
  xlab("Coefficient") +
  theme(text = element_text(size = 20)) +
  theme(axis.title.y = element_blank())

p

```






#  Test the effect of changing the median of the cohorts across different values


```{r}

max_f_fer <- 1
#Test changing from the original values down to Finnish pre levels
#gpf <-  26 /134 
min_f_fer <- gpf
length.out <- 10

# Generate length.out equally spaced values for each variable
values_var1 <- seq(min_f_fer, max_f_fer, length.out = length.out)

# Create a grid of all combinations of these values
INPUT <- expand.grid(donor = values_var1, gp = values_var1)

data <- regression_data_men_nl



simres <- INPUT %>% 
  mutate(
    newdata = map2(donor,gp, ~ {
      data %>%
        mutate(
          log_ferritin = case_when(
###Change the Cohort variable values to match NL###
            Cohort == "FINDONOR" ~ log(exp(log_ferritin) * .x),
            Cohort == "THL" | Cohort == "NEWDONOR"  ~ log(exp(log_ferritin) * .y)
          )
        )
    }
    )
  )
   

fermodel <- function(df) {
  lm(log_ferritin ~   Age + 
       donation_count_2 + 
       donation_count +
       log_last_donation + 
       Cohort +
       blood_vol,
     data = df)
}

simres <- simres %>% 
  mutate(
    newmodel = map(newdata, fermodel)
  )
  
tidy_coef <- function(x) {
  tidy(x,conf.int = TRUE)
}



simres <- simres %>%  
  mutate(
    # Use tidy to extract coefficients and confidence intervals
    coefficients = map(newmodel, tidy_coef),
    # Unnest the coefficients column
    coefficients = map(coefficients, ~ select(.x, term, estimate, conf.low, conf.high)),
    # Separate coefficients into columns
    
###Change the Cohort variable values to match NL###

    GPestimate = map_dbl(coefficients, ~ filter(.x, term == "CohortTHL")$estimate),
    GPconf_low = map_dbl(coefficients, ~ filter(.x, term == "CohortTHL")$conf.low),
    GPconf_high = map_dbl(coefficients, ~ filter(.x, term == "CohortTHL")$conf.high),
    NDestimate = map_dbl(coefficients, ~ filter(.x, term == "CohortNEWDONOR")$estimate),
    NDconf_low = map_dbl(coefficients, ~ filter(.x, term == "CohortNEWDONOR")$conf.low),
    NDconf_high = map_dbl(coefficients, ~ filter(.x, term == "CohortNEWDONOR")$conf.high)
  )
  
head(simres)
```

```{r}
tail(simres)
```


```{r}


med_fer <- data %>% 
  group_by(Cohort) %>% 
  mutate(ferritin = exp(log_ferritin)) %>% 
  summarise(medianfer = median(ferritin))
med_fer
```



## Plot

```{r}



tmp <- pivot_longer(
  simres %>% select(GPestimate,NDestimate,donor,gp),cols=c(GPestimate,NDestimate )
  ) %>% 
  mutate(
###Change the Cohort variable values to match NL###
    Cohort = case_when(
      name == "GPestimate" ~ "THL",
      name == "NDestimate" ~ "NEWDONOR"
    )
  ) %>% inner_join(med_fer,by=c("Cohort" = "Cohort") )

p <- ggplot(tmp)
p <- p + geom_point(
  aes(y=value, 
      x=donor * med_fer$medianfer[1],
      color=gp  * medianfer 
      ),data=tmp
      
  )
p <- p + xlab("Donor cohort i.e. reference median ferritin (ng/l)") + ylab("General population coefficient estimate")
p <- p + scale_colour_continuous(name="Other \ncohort median\nferritin (ng/l)")
p <- p + geom_text(aes(x = 26,  y =plotdata_wnd$m[6],label="as_pre_fi_pg"),col="red")
p <- p + geom_text(aes(x = 55,  y =plotdata_wnd$m[13],label="original_gp"),col="red")
p <- p + geom_point(aes(x = 26,  y =plotdata_wnd$m[6]),col="red")
p <- p + geom_point(aes(x = 55,  y =plotdata_wnd$m[13]),col="red")
p <- p + facet_grid(. ~ Cohort)
p

```


# Test with out new donors


```{r}

max_f_fer <- 1
#Test changing from the original values down to Finnish pre levels
#gpf <-  26 /134 
min_f_fer <- gpf
length.out <- 10

# Generate length.out equally spaced values for each variable
values_var1 <- seq(min_f_fer, max_f_fer, length.out = length.out)

# Create a grid of all combinations of these values
INPUT <- expand.grid(donor = values_var1, gp = values_var1)

###Change the Cohort variable values to match NL###
data <- regression_data_men_nl %>% filter(Cohort != "NEWDONOR")



simres <- INPUT %>% 
  mutate(
    newdata = map2(donor,gp, ~ {
      data %>%
        mutate(
          log_ferritin = case_when(
###Change the Cohort variable values to match NL###
            Cohort == "FINDONOR" ~ log(exp(log_ferritin) * .x),
            Cohort == "THL" ~ log(exp(log_ferritin) * .y)
          )
        )
    }
    )
  )
   

fermodel <- function(df) {
  lm(log_ferritin ~   Age + 
       donation_count_2 + 
       donation_count +
       log_last_donation + 
       Cohort +
       blood_vol,
     data = df)
}

simres <- simres %>% 
  mutate(
    newmodel = map(newdata, fermodel)
  )
  
tidy_coef <- function(x) {
  tidy(x,conf.int = TRUE)
}



simres <- simres %>%  
  mutate(
    # Use tidy to extract coefficients and confidence intervals
    coefficients = map(newmodel, tidy_coef),
    # Unnest the coefficients column
    coefficients = map(coefficients, ~ select(.x, term, estimate, conf.low, conf.high)),
    # Separate coefficients into columns
    
###Change the Cohort variable values to match NL###

    GPestimate = map_dbl(coefficients, ~ filter(.x, term == "CohortTHL")$estimate),
    GPconf_low = map_dbl(coefficients, ~ filter(.x, term == "CohortTHL")$conf.low),
    GPconf_high = map_dbl(coefficients, ~ filter(.x, term == "CohortTHL")$conf.high),
  )
  
head(simres)
```

```{r}
tail(simres)
```


```{r}


med_fer <- data %>% 
  group_by(Cohort) %>% 
  mutate(ferritin = exp(log_ferritin)) %>% 
  summarise(medianfer = median(ferritin))
med_fer
```



## Plot

```{r}



tmp <- pivot_longer(
  simres %>% select(GPestimate,donor,gp),cols=c(GPestimate)
  ) %>% 
  mutate(
###Change the Cohort variable values to match NL###
    Cohort = case_when(
      name == "GPestimate" ~ "THL",
     )
  ) %>% inner_join(med_fer,by=c("Cohort" = "Cohort") )

p <- ggplot(tmp)
p <- p + geom_point(
  aes(y=value, 
      x=donor * med_fer$medianfer[1],
      color=gp  * medianfer 
      ),data=tmp
      
  )
p <- p + xlab("Donor cohort i.e. reference median ferritin (ng/l)") + ylab("General population coefficient estimate")
p <- p + scale_colour_continuous(name="General population\ncohort median\nferritin (ng/l)")
p <- p + geom_text(aes(x = 26,  y =plotdata_wnd$m[6],label="as_pre_fi_pg"),col="red")
p <- p + geom_text(aes(x = 55,  y =plotdata_wnd$m[13],label="original_gp"),col="red")
p <- p + geom_point(aes(x = 26,  y =plotdata_wnd$m[6]),col="red")
p <- p + geom_point(aes(x = 55,  y =plotdata_wnd$m[13]),col="red")
#p <- p + facet_grid(. ~ Cohort)
p

```


# Manipulate donation histories


```{r}



```

