---
title: "Ferritin model subgroup analysis"
author: "Sofie Ekroos"
date: "25.11.2021"
output: pdf_document

---


# Summary

This code is based on Mikko Arvas' Bayesian model for ferritin, "Ferritin model" (https://github.com/FRCBS/iron_measurement_comparisons/blob/master/src/ferritin_model_bayes.Rmd). In this analysis we analyse ferritin levels in pre-menopausal women.

Version 2: blood donation variables removed completely for general population.

All variables:

Age  
BMI  
Smoking
iron_complience (only blood donor cohort)
donation_count_2 (only blood donor cohort)
donation_count (only blood donor cohort)
log_last_donation (only blood donor cohort)
log_CRP 
Region (not in blood donor cohort)
Weigh
Menstruation
RedMeat_n (not in h2000 cohort)
PreviousChildbirth


```{r setup, , message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
# echo "rmarkdown::render('ferritin_model_subgroup_analysis_bayes.Rmd', clean=TRUE,output_format='pdf_document',output_file='../results/bmodels_subgroup/version1_ferritin_model_subgroup_analysis/version1_ferritin_model_subgroup_analysis_bayes.pdf')" | R --slave
library(tidyverse)
library(brms)
library(bayesplot)
```



PART 2: BAYESIAN LINEAR REGRESSION

# Data loading and preparation 

Load data

```{r}
load("../data/ID_data_regression_cohorts.rdata")
all_cohorts <- regression_cohorts %>% filter(Group == "Pre_menopause_women") 

all_cohorts <- all_cohorts %>% 
  filter(!is.na(CRP)) %>% 
  filter(! CRP >= 30)
```


# Build subgroups of pre-menopausal women (three cohorts: FinDonor, FinRisk97, Health2000) 

## Data

```{r}
# FinDonor
regression_data_blood <- all_cohorts %>% 
  filter(Cohort == "FINDONOR") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, iron_deficiency, Smoking, iron_complience, Region, Weight, Menstruation, PreviousChildbirth, RedMeat_n) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         Weight = scale(Weight, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# FinRisk97
regression_data_fr1997 <- all_cohorts %>% 
  filter(Cohort == "FINRISK97") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, iron_deficiency, Smoking, iron_complience, Region, Weight, Menstruation, PreviousChildbirth, RedMeat_n) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         Weight = scale(Weight, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# Health2000 
regression_data_h2000 <- all_cohorts %>% 
  filter(Cohort == "HEALTH00") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, iron_deficiency, Smoking, iron_complience, Region, Weight, Menstruation, PreviousChildbirth, RedMeat_n) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         Weight = scale(Weight, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 


```


## Models 

Running the diagnostics in this code, we can see that chains do not mix properly in the general population cohorts due to the variables iron_complience and donation_count_2. The observations for these two cohorts were imputed due to missing data (donation interval between 3-6 years, 0 donations in the past two years and iron supplementation as no supplementation taken, see low_ferritin_data). For this reason, we removed these two varibles for the general population cohorts. 

### FinDonor

```{r blood}
file <- "../results/bmodels_subgroup/version2_ferritin_model_subgroup_analysis/linear_blood_b"

linear_blood_b <- brm(log_ferritin ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        iron_complience + 
                        donation_count_2 + 
                        donation_count +
                        log_last_donation + 
                        log_CRP +
                        #Region + 
                        Weight +
                        Menstruation +
                        RedMeat_n + 
                        PreviousChildbirth, 
                      data = regression_data_blood, 
                      #family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(linear_blood_b)

```

### FinRisk97

```{r  fr1997}
file <- "../results/bmodels_subgroup/version2_ferritin_model_subgroup_analysis/linear_fr1997_b"

linear_fr1997_b <- brm(log_ferritin ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        #iron_complience + 
                        #donation_count_2 + 
                        #donation_count +
                        #log_last_donation + 
                        log_CRP +
                        Region + 
                        Weight +
                        Menstruation +
                        RedMeat_n + 
                        PreviousChildbirth, 
                      data = regression_data_fr1997,
                      #family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(linear_fr1997_b)


```

### Health2000

```{r  h2000}
file <- "../results/bmodels_subgroup/version2_ferritin_model_subgroup_analysis/linear_h2000_b"

linear_h2000_b <- brm(log_ferritin ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        #iron_complience + 
                        #donation_count_2 + 
                        #donation_count +
                        #log_last_donation + 
                        log_CRP +
                        Region + 
                        Weight +
                        Menstruation +
                        #RedMeat_n + (all NA answers)
                        PreviousChildbirth, 
                     data = regression_data_h2000, 
                     #family = bernoulli(),
                     file=file, 
                     cores = 4,
                     iter= 10000
)

summary(linear_h2000_b)
```

## Diagnostics

### Trace

```{r , message=FALSE}
mcmc_plot(linear_blood_b, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(linear_fr1997_b, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(linear_h2000_b, type = "trace")
```

Chains do not seem to get stuck.

### ACF bars

```{r}
mcmc_plot(linear_blood_b, type = "acf_bar")
```


```{r}
mcmc_plot(linear_fr1997_b, type = "acf_bar")
```


```{r}
mcmc_plot(linear_h2000_b, type = "acf_bar")
```

Autocorrelation drops nicely.

### Rhat

```{r}
mcmc_plot(linear_blood_b, type = "rhat")
```

```{r}
mcmc_plot(linear_fr1997_b, type = "rhat")
```

```{r}
mcmc_plot(linear_h2000_b, type = "rhat")
```
https://mc-stan.org/rstan/reference/Rhat.html


"The Rhat function produces R-hat convergence diagnostic, which compares the between- and within-chain estimates for model parameters and other univariate quantities of interest. If chains have not mixed well (ie, the between- and within-chain estimates don't agree), R-hat is larger than 1. We recommend running at least four chains by default and only using the sample if R-hat is less than 1.05."

## Forest plots

```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(linear_blood_b, 
         type = "areas",
         prob = 0.95,
         prob_est="median"
         #transformations = "exp"
         ) +
  geom_vline(xintercept = 0, color = "grey")

#p <- p + scale_x_log10()
#Ferritin needs to be removed
# plot.OR<-  tidy_logit_pre_menop_women %>%filter(!term=="log_Ferritin_beginning") %>%  dplyr::select(OR) %>% mutate(y=rev(2:(nrow(tidy_logit_pre_menop_women)))) %>% bind_rows(tibble(OR=NA,y=1 ))
# p <- p + geom_point(aes(y=y,x=OR),data=plot.OR, col="red")
p
```


```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(linear_fr1997_b, 
         type = "areas",
         prob=0.95,
         prob_est="median"
         #transformations = "exp"
        ) +
  geom_vline(xintercept = 0, color = "grey")

#p <- p + scale_x_log10()
p
```


```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(linear_h2000_b, 
         type = "areas",
         prob = 0.95,
         prob_est="median"
         #transformations = "exp"
         ) +
  geom_vline(xintercept = 0, color = "grey")

#p <- p + scale_x_log10()
#Ferritin needs to be removed
# plot.OR<-  tidy_logit_pre_menop_women %>%filter(!term=="log_Ferritin_beginning") %>%  dplyr::select(OR) %>% mutate(y=rev(2:(nrow(tidy_logit_pre_menop_women)))) %>% bind_rows(tibble(OR=NA,y=1 ))
# p <- p + geom_point(aes(y=y,x=OR),data=plot.OR, col="red")
p
```

### All groups together

```{r}
regressor_values <- c(
 "Age" = "Age (5 years)",
   "log_CRP" = "CRP (log transformed)",
   "BMI" = "BMI",
   "log_last_donation" = "Days since last donation (log, /log(2))",
   "Smokingyes" ="Smoker",
    "iron_complience" = "Iron supplementation",
    "donation_count_2" = "Donations in last 2 years (quadratic)",
    "Menstruationregular_period" = "Regular periods",
    "Weight" = "Weight (10 kg)",
    "Region" = "Region",
    "RedMeat_n" = "Red meat",
    "PreviousChildbirth" = "Given birth in the past",
    "RegionTYKS" = "Region (Turku university hospital)",
    "RegionTAYS" = "Region (Tampere university hospital)",
    "RegionOYS" = "Region (Oulu university hospital)",
    "RegionKYS" = "Region (Kuopio university hospital)",
    "donation_count" = "Donations in last 2 years"
 
)
```




```{r }

intervals_blood <- mcmc_intervals_data(
  linear_blood_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median"
  #transformations = "exp"
  ) %>% mutate(Cohort='FINDONOR')

intervals_fr1997 <- mcmc_intervals_data(
  linear_fr1997_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median"
  #transformations = "exp"
  ) %>% mutate(Cohort='FINRISK97') %>% 
  arrange(m) # how to order the variables?

intervals_h2000 <- mcmc_intervals_data(
  linear_h2000_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median"
  #transformations = "exp"
  ) %>% mutate(Cohort='HEALTH00')

intervals_combined <- rbind(intervals_blood,intervals_fr1997,intervals_h2000) %>% 
  filter(! (parameter == 'b_Intercept' | parameter == 'lp__' | parameter == 'sigma')) %>% 
  mutate(parameter = gsub("b_","",parameter)) %>% 
  mutate(regressor = plyr::revalue(parameter, regressor_values)) %>% 
  select(-parameter)
  #arrange(m)


```


```{r , fig1, fig.height = 10, fig.width = 15 ,warning=FALSE}

orlower <- -1
orupper <- 1

# Truncate ORs for plotting
plotdata <- intervals_combined %>% 
  mutate(
    ll = case_when(
      ll < orlower ~ orlower,
      TRUE ~ ll
    ),
    hh = case_when(
      hh > orupper ~ orupper,
      TRUE ~ hh
    )
  )

#theme_set(bayesplot::theme_default())
pos <- position_nudge(y=as.numeric(as.factor(plotdata$Cohort))/5 -0.1) 

p <- ggplot(plotdata, aes(x = m, y = regressor, color = Cohort)) + 
  geom_point(position = pos,
             size=4) +
  geom_linerange(aes(xmin = ll, xmax = hh), 
                 position = pos,
                 size = 1, 
                 linetype = 1
                 ) +
  scale_color_manual(values = c(  "#E84A5F", "#99B898", "#FECEAB" ),
                     limits = c("FINDONOR",  "FINRISK97","HEALTH00")) +
  scale_fill_manual(values = c(  "#E84A5F", "#99B898", "#FECEAB" ),
                    limits = c("FINDONOR",  "FINRISK97","HEALTH00")) +
  scale_shape_manual(values = c(1,16)) +
  xlim(orlower,orupper)  +   
  geom_vline(xintercept = 0 , color = "grey", linetype = "dotted",
             size = 1) +
  theme(axis.title = element_text(size = 14),
        strip.text =  element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.background = element_rect(fill = "white"),
        strip.background.x = element_blank(),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14)
  ) +
  guides(shape = "none") +
  theme_classic() +
  ylab("Coefficient")  + 
  ggtitle("Ferritin model (premenopausal women, version 2)") 

p


```

```{r}


ggplot2::ggsave(filename = "../results/figures/subgroup_analysis/version2_forest_subgroup_analysis_bayes_ferritin.pdf",
                plot = p,
                width = 35,
                height = 25,
                dpi = 600,
                units = "cm")

# save ORs for comparison with other models 
OR_linear_subgroup_pre <- plotdata
save(OR_linear_subgroup_pre, file = "../data/OR_ferritin_model_subgroup_analysis.rdata")


```

