---
title: "Low ferritin model"
author: "Mikko Arvas"
date: "15.22.2021"
output: pdf_document

---


# Summary


```{r setup, , message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
# echo "rmarkdown::render('low_ferritin_model_bayes.Rmd', clean=TRUE,output_format='pdf_document',output_file='../results/low_ferritin_model_bayes.pdf')" | R --slave
library(tidyverse)
library(brms)
library(bayesplot)
```



PART 2: BAYESIAN LOGISTIC REGRESSION

# Data loading and preparation 

Load data

```{r}
load("../data/ID_data_regression_cohorts.rdata")
all_cohorts <- regression_cohorts
```


# All cohorts together 

## Data
```{r}
# men
## menstruation variable removed 
regression_data_men <- all_cohorts %>% 
  filter(Group == "Men") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, Cohort, iron_deficiency, Smoking, iron_complience, Group) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# premenopausal women
regression_data_pre <- all_cohorts %>% 
  filter(Group == "Pre_menopause_women") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, Menstruation, Cohort, iron_deficiency, Smoking, iron_complience, Group) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# postmenopausal womem 
## menstruation variable removed 
regression_data_post <- all_cohorts %>% 
  filter(Group == "Post_menopause_women") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, Cohort, iron_deficiency, Smoking, iron_complience, Group) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 


```

## Models 



```{r men}
#men
file <- "../results/bmodels/logit_men_b"

logistic_men_b <- brm(iron_deficiency ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        iron_complience + 
                        donation_count_2 + 
                        log_last_donation + 
                        log_CRP +
                        Cohort, 
                      data = regression_data_men, 
                      family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(logistic_men_b)

```

```{r  pre}
# premenopausal women
file <- "../results/bmodels/logit_pre_b"

logistic_pre_b <- brm(iron_deficiency ~ Age + 
                        BMI + 
                        Smoking + 
                        iron_complience + 
                        donation_count_2 + 
                        log_last_donation + 
                        log_CRP +
                        Cohort + 
                        Menstruation, 
                      data = regression_data_pre,
                      family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(logistic_pre_b)


```

```{r  post}
# postmenopausal women
file <- "../results/bmodels/logit_post_b"

logistic_post_b <- brm(iron_deficiency ~ Age + 
                       BMI + 
                       Smoking + 
                       iron_complience + 
                       donation_count_2 + 
                       log_last_donation + 
                       log_CRP +
                       Cohort, 
                     data = regression_data_post, 
                     family = bernoulli(),
                     file=file, 
                     cores = 4,
                     iter= 10000
)

summary(logistic_post_b)
```

## Diagnostics

### Trace

```{r , message=FALSE}
mcmc_plot(logistic_men_b, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(logistic_pre_b, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(logistic_post_b, type = "trace")
```

Chains do not seem to get stuck.

### ACF bars

```{r}
mcmc_plot(logistic_men_b, type = "acf_bar")
```


```{r}
mcmc_plot(logistic_pre_b, type = "acf_bar")
```


```{r}
mcmc_plot(logistic_post_b, type = "acf_bar")
```

Autocorrelation drops nicely.

### Rhat

```{r}
mcmc_plot(logistic_men_b, type = "rhat")
```

```{r}
mcmc_plot(logistic_pre_b, type = "rhat")
```

```{r}
mcmc_plot(logistic_post_b, type = "rhat")
```
https://mc-stan.org/rstan/reference/Rhat.html


"The Rhat function produces R-hat convergence diagnostic, which compares the between- and within-chain estimates for model parameters and other univariate quantities of interest. If chains have not mixed well (ie, the between- and within-chain estimates don't agree), R-hat is larger than 1. We recommend running at least four chains by default and only using the sample if R-hat is less than 1.05."

## Forest plots

```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(logistic_men_b, 
         type = "areas",
         prob = 0.95,
         transformations = "exp") +
  geom_vline(xintercept = 1, color = "grey")

p <- p + scale_x_log10()
#Ferritin needs to be removed
# plot.OR<-  tidy_logit_pre_menop_women %>%filter(!term=="log_Ferritin_beginning") %>%  dplyr::select(OR) %>% mutate(y=rev(2:(nrow(tidy_logit_pre_menop_women)))) %>% bind_rows(tibble(OR=NA,y=1 ))
# p <- p + geom_point(aes(y=y,x=OR),data=plot.OR, col="red")
p
```


```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(logistic_pre_b, 
         type = "areas",
         prob=0.95,
         prob_est="median",
         transformations = "exp"
        ) +
  geom_vline(xintercept = 1, color = "grey")

p <- p + scale_x_log10()
p
```


```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(logistic_post_b, 
         type = "areas",
         prob = 0.95,
         transformations = "exp") +
  geom_vline(xintercept = 1, color = "grey")

p <- p + scale_x_log10()
#Ferritin needs to be removed
# plot.OR<-  tidy_logit_pre_menop_women %>%filter(!term=="log_Ferritin_beginning") %>%  dplyr::select(OR) %>% mutate(y=rev(2:(nrow(tidy_logit_pre_menop_women)))) %>% bind_rows(tibble(OR=NA,y=1 ))
# p <- p + geom_point(aes(y=y,x=OR),data=plot.OR, col="red")
p
```

### All groups together

```{r }

intervals_men <- mcmc_intervals_data(
  logistic_men_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median",
  transformations = "exp"
  ) %>% mutate(Group='Men')

intervals_pre <- mcmc_intervals_data(
  logistic_pre_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median",
  transformations = "exp"
  ) %>% mutate(Group='Pre') %>% 
  arrange(m) # how to order the variables?

intervals_post <- mcmc_intervals_data(
  logistic_post_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median",
  transformations = "exp"
  ) %>% mutate(Group='Post')

intervals_combined <- rbind(intervals_pre,intervals_post,intervals_men) %>% 
  filter(! (parameter == 'exp(b_Intercept)' | parameter == 'exp(lp__)')) %>% 
  mutate(parameter = gsub("\\)","",gsub("exp\\(b_","",parameter))) #%>% 
  #arrange(m)


```


```{r , warning=FALSE}

orlower <- 1/10
orupper <- 10

# Truncate ORs for plotting
temp <- intervals_combined %>% 
  mutate(
    ll = case_when(
      ll < orlower ~ orlower,
      TRUE ~ ll
    ),
    hh = case_when(
      hh > orupper ~ orupper,
      TRUE ~ hh
    )
  )

theme_set(bayesplot::theme_default())
pos <- position_nudge(y=as.numeric(as.factor(temp$Group))/10 -0.1) 

p <- ggplot(temp, aes(x = m, y = parameter, color = Group)) + 
  geom_point(position = pos) +
  geom_linerange(aes(xmin = ll, xmax = hh), position = pos) +
 scale_color_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                     limits = c("Men",  "Post","Pre" )) +
     scale_fill_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                     limits = c("Men",  "Post","Pre" )) +
 scale_shape_manual(values = c(1,16)) +
scale_x_log10(limits=c(orlower,orupper))  + #how to get the lines that drop out of the visible?  
  geom_vline(xintercept = 1 , color = "grey", linetype = "dotted",
             size = 1)

p


```

```{r}


ggplot2::ggsave(filename = "../results/forest_all_cohorts_b.pdf",
                plot = p,
                width = 35,
                height = 18,
                dpi = 600,
                units = "cm")


```

