---
title: "Low ferritin model"
author: "Mikko Arvas"
date: "15.11.2021"
output: pdf_document

---


# Summary

Version 3: removed region variable

```{r setup, , message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
# echo "rmarkdown::render('low_ferritin_model_bayes_version2.Rmd', clean=TRUE,output_format='pdf_document',output_file='../results/low_ferritin_model_bayes/low_ferritin_model_bayes_version2.pdf')" | R --slave
library(tidyverse)
library(brms)
library(bayesplot)
```

PART 2: BAYESIAN LOGISTIC REGRESSION

# Data loading and preparation 

Load data

```{r}
load("../data/ID_data_regression_cohorts.rdata")
all_cohorts <- regression_cohorts

all_cohorts <- all_cohorts %>% 
  filter(!is.na(CRP)) %>% 
  filter(! CRP >= 30)
```

# All cohorts together 

## Data
```{r}
# men
## menstruation and childbirth variables removed 
regression_data_men <- all_cohorts %>% 
  filter(Group == "Men") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, Cohort, iron_deficiency, Smoking, iron_complience, Group, Region, Weight, RedMeat_n) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# premenopausal women
regression_data_pre <- all_cohorts %>% 
  filter(Group == "Pre_menopause_women") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, Cohort, iron_deficiency, Smoking, iron_complience, Group, Region, Weight, RedMeat_n, Menstruation, PreviousChildbirth) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# postmenopausal womem 
## menstruation variable removed 
regression_data_post <- all_cohorts %>% 
  filter(Group == "Post_menopause_women") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, Cohort, iron_deficiency, Smoking, iron_complience, Group, Region, Weight, RedMeat_n, Menstruation, PreviousChildbirth) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 


```

## Models 


```{r men}
#men
file <- "../results/bmodels/version3/logit_men_b"

logistic_men_b <- brm(iron_deficiency ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        iron_complience + 
                        donation_count_2 + 
                        donation_count +
                        log_last_donation + 
                        log_CRP +
                        #Region + 
                        Weight +
                        #RedMeat_n +
                        Cohort, 
                      data = regression_data_men, 
                      family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(logistic_men_b)

```

```{r  pre}
# premenopausal women
file <- "../results/bmodels/version3/logit_pre_b"

logistic_pre_b <- brm(iron_deficiency ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        iron_complience + 
                        donation_count_2 + 
                        donation_count +
                        log_last_donation + 
                        log_CRP +
                        #Region + 
                        Weight +
                        #RedMeat_n +
                        Cohort +
                        Menstruation +
                        PreviousChildbirth, 
                      data = regression_data_pre,
                      family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(logistic_pre_b)


```

```{r  post}
# postmenopausal women
file <- "../results/bmodels/version3/logit_post_b"

logistic_post_b <- brm(iron_deficiency ~
                       Age + 
                        BMI + 
                        Smoking + 
                        iron_complience + 
                        donation_count_2 + 
                        donation_count +
                        log_last_donation + 
                        log_CRP +
                        #Region + 
                        Weight +
                        #RedMeat_n +
                        Cohort +
                        #Menstruation +   # will not work as all observations are no_period
                        PreviousChildbirth, 
                     data = regression_data_post, 
                     family = bernoulli(),
                     file=file, 
                     cores = 4,
                     iter= 10000
)

summary(logistic_post_b)
```

## Diagnostics

### Trace

```{r , message=FALSE}
mcmc_plot(logistic_men_b, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(logistic_pre_b, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(logistic_post_b, type = "trace")
```

Chains do not seem to get stuck.

### ACF bars

```{r}
mcmc_plot(logistic_men_b, type = "acf_bar")
```


```{r}
mcmc_plot(logistic_pre_b, type = "acf_bar")
```


```{r}
mcmc_plot(logistic_post_b, type = "acf_bar")
```

Autocorrelation drops nicely.

### Rhat

```{r}
mcmc_plot(logistic_men_b, type = "rhat")
```

```{r}
mcmc_plot(logistic_pre_b, type = "rhat")
```

```{r}
mcmc_plot(logistic_post_b, type = "rhat")
```
https://mc-stan.org/rstan/reference/Rhat.html


"The Rhat function produces R-hat convergence diagnostic, which compares the between- and within-chain estimates for model parameters and other univariate quantities of interest. If chains have not mixed well (ie, the between- and within-chain estimates don't agree), R-hat is larger than 1. We recommend running at least four chains by default and only using the sample if R-hat is less than 1.05."

## Forest plots

```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(logistic_men_b, 
         type = "areas",
         prob = 0.95,
         transformations = "exp") +
  geom_vline(xintercept = 1, color = "grey")

p <- p + scale_x_log10()
#Ferritin needs to be removed
# plot.OR<-  tidy_logit_pre_menop_women %>%filter(!term=="log_Ferritin_beginning") %>%  dplyr::select(OR) %>% mutate(y=rev(2:(nrow(tidy_logit_pre_menop_women)))) %>% bind_rows(tibble(OR=NA,y=1 ))
# p <- p + geom_point(aes(y=y,x=OR),data=plot.OR, col="red")
p
```


```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(logistic_pre_b, 
         type = "areas",
         prob=0.95,
         prob_est="median",
         transformations = "exp"
        ) +
  geom_vline(xintercept = 1, color = "grey")

p <- p + scale_x_log10()
p
```


```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(logistic_post_b, 
         type = "areas",
         prob = 0.95,
         transformations = "exp") +
  geom_vline(xintercept = 1, color = "grey")

p <- p + scale_x_log10()
#Ferritin needs to be removed
# plot.OR<-  tidy_logit_pre_menop_women %>%filter(!term=="log_Ferritin_beginning") %>%  dplyr::select(OR) %>% mutate(y=rev(2:(nrow(tidy_logit_pre_menop_women)))) %>% bind_rows(tibble(OR=NA,y=1 ))
# p <- p + geom_point(aes(y=y,x=OR),data=plot.OR, col="red")
p
```

### All groups together

```{r}
regressor_values <- c(
 "Age" = "Age (5 years)",
   "log_CRP" = "CRP (log transformed)",
   "BMI" = "BMI",
   "log_last_donation" = "Days since last donation (log, /log(2))",
   "Smokingyes" ="Smoker",
    "iron_complience" = "Iron supplementation",
    "donation_count_2" = "Donations in last 2 years (quadratic)",
    "Menstruationregular_period" = "Regular periods",
    "Weight" = "Weight (10 kg)",
    "PreviousChildbirth" = "Given birth in the past",
    "donation_count" = "Donations in last 2 years",
    "CohortHEALTH00" = "Cohort (Health 2000)",
    "CohortFINRISK97" = "Cohort (FinRisk 1997)"

)
```




```{r }

intervals_men <- mcmc_intervals_data(
  logistic_men_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median",
  transformations = "exp"
  ) %>% mutate(Group='Men')

intervals_pre <- mcmc_intervals_data(
  logistic_pre_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median",
  transformations = "exp"
  ) %>% mutate(Group='Pre-menopausal women') %>% 
  arrange(m) # how to order the variables?

intervals_post <- mcmc_intervals_data(
  logistic_post_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median",
  transformations = "exp"
  ) %>% mutate(Group='Post-menopausal women')

intervals_combined <- rbind(intervals_pre,intervals_post,intervals_men) %>% 
  filter(! (parameter == 'exp(b_Intercept)' | parameter == 'exp(lp__)')) %>% 
  mutate(parameter = gsub("\\)","",gsub("exp\\(b_","",parameter))) %>% 
  mutate(regressor = plyr::revalue(parameter, regressor_values)) %>% 
  select(-parameter)
  #arrange(m)


```


```{r , fig1, fig.height = 10, fig.width = 15 ,warning=FALSE}

orlower <- 1/10
orupper <- 10

# Truncate ORs for plotting
plotdata <- intervals_combined %>% 
  mutate(
    ll = case_when(
      ll < orlower ~ orlower,
      TRUE ~ ll
    ),
    hh = case_when(
      hh > orupper ~ orupper,
      TRUE ~ hh
    )
  )

#theme_set(bayesplot::theme_default())
pos <- position_nudge(y=as.numeric(as.factor(plotdata$Group))/5 -0.1) 

p <- ggplot(plotdata, aes(x = m, y = regressor, color = Group)) + 
  geom_point(position = pos,
             size=4) +
  geom_linerange(aes(xmin = ll, xmax = hh), 
                 position = pos,
                 size = 1, 
                 linetype = 1
                 ) +
  scale_color_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                     limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
  scale_fill_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                    limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
  scale_shape_manual(values = c(1,16)) +
  scale_x_log10(limits=c(orlower,orupper))  + #how to get the lines that drop out of the visible?  
  geom_vline(xintercept = 1 , color = "grey", linetype = "dotted",
             size = 1) +
  theme(axis.title = element_text(size = 14),
        strip.text =  element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.background = element_rect(fill = "white"),
        strip.background.x = element_blank(),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14)
  ) +
  guides(shape = "none") +
  theme_classic() +
  ylab("Coefficient") 

p


```

```{r}


ggplot2::ggsave(filename = "../results/figures/low_ferritin_model/version3_forest_all_cohorts_bayes.pdf",
                plot = p,
                width = 35,
                height = 25,
                dpi = 600,
                units = "cm")

# save ORs for comparison with other models 
OR_logistic_bayes <- plotdata
save(OR_logistic_bayes, file = "../data/OR_low_ferritin_model_bayes.rdata")


```



## Compare to bootstrapped



```{r  fig2, fig.height = 15, fig.width = 10 ,warning=FALSE}
file <- "../results/bootstraps/plotdata_filtered.rdata"
if(file.exists(file)) {
  load(file)
  temp <- plotdata_filtered %>% select(-p.value,-is_sig) %>% 
    rename(ll = Bca_inf,
           hh = Bca_sup,
           m = OR
    ) %>%  mutate(
      Method = "Bayes"
    ) %>%  rename(Group=group)
  
  temp <- plotdata %>%  select(regressor,ll,hh,m,Group) %>% 
    mutate(Method="Boot") %>% 
  bind_rows(temp)
  
  temp <- temp %>% 
    mutate(
      ll = case_when(
        ll < orlower ~ orlower,
        TRUE ~ ll
      ),
      hh = case_when(
        hh > orupper ~ orupper,
        TRUE ~ hh
      )
    )
  
  pos <- position_nudge(y=as.numeric(as.factor(temp$Method))/5 -0.1) 
  
  p <- ggplot(temp, aes(x = m, y = regressor, color = Method)) + 
     geom_point(position = pos,
                size=4) +
    geom_linerange(aes(xmin = ll, xmax = hh), 
                   position = pos,
                   size = 1 
    ) +
    # scale_color_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
    #                    limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
    # scale_fill_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
    #                   limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
    #scale_shape_manual(values = c(1,16)) +
    scale_x_log10(limits=c(orlower,orupper))  + #how to get the lines that drop out of the visible?  
    geom_vline(xintercept = 1 , color = "grey", linetype = "dotted",
               size = 1) +
    theme(axis.title = element_text(size = 14),
          strip.text =  element_text(size = 14),
          legend.text = element_text(size = 12),
          strip.background = element_rect(fill = "white"),
          strip.background.x = element_blank(),
          axis.text.x = element_text(size = 14),
          axis.text.y = element_text(size = 14)
    ) +
    guides(shape = "none") +
    theme_classic() +
    facet_grid(Group~.) +
    ylab("Coefficient") 
  
  p
  
  
}
```


```{r}
if(file.exists(file)) {

ggplot2::ggsave(filename = "../results/figures/forest_all_cohorts_bayes_vs_boot.pdf",
                plot = p,
                width = 35,
                height = 30,
                dpi = 600,
                units = "cm")
}

```