---
title: "Low ferritin model"
author: "Sofie Ekroos"
date: "19.7.2021"
output: github_document
bibliography: index.bib
---

Based on the original code written by Muriel Lobier and Mikko Arvas (https://github.com/FRCBS/iron_levels_of_blood_donors/blob/master/src/index.Rmd and https://github.com/FRCBS/changes_in_donor_health/blob/master/src/explorative_analysis.Rmd)

# Summary

This document includes all codes necessary to run the analysis of and produce the figures for the three Finnish Cohorts (FinDonor, FINRISK97, Health2000). The code allows the user to:

- describe the cohorts and compute the prevalence of iron deficiency
- identify co-variants of iron deficiency using multivariable linear regression 

```{r setup, , message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# library(janitor)
library(tableone)
library(stargazer)
library(gridExtra)
library(cowplot)
library(broom)
library(GGally)
library(ggfortify)
# library(sjPlot)
library(knitr)
library(lubridate)
#library(HDInterval) !! package ‘HDIinterval’ is not available (for R version 3.6.3)!!!
library(car)
#library(relaimpo)
library(sfsmisc)
library(MASS)
# library(kableExtra)
#library(brms)
#library(rstan)
#library(rms)
#library(modelr)
#library(tidybayes)
library(ordinal)
library(sjmisc)
#library(sjlabelled)
# library(lazerhawk)
library(epiDisplay)
library(safeBinaryRegression)
library(epitools)
library(ez)
library(ggbeeswarm)
library(tidymodels)
```

# Data loading and preparation

```{r , warning=FALSE}

#load FinDonor biomarker data
load("../data/r02.fd.bd.all.rdata")
indiv_donations_data <- output
rm(output)

#load FinDonor demographic data
load("../data/r02ds.donorData.rdata")
donor_demo <- output
rm(output)

#THL data
load("../data/thldata.rdata")
#thldalta.rdata contains all five THL cohorts, extract FINRISK97 and Health2000 from the others 
fr1997 <- thldata$fr1997
h2000 <- thldata$h2000
rm(thldata)
```

## Data preparation

There are `r indiv_donations_data %>% distinct(donor) %>% nrow()` donors enrolled in the study. 

There are `r indiv_donations_data %>%  filter(!is.na(Ferritin) & !is.na(Hb_v) ) %>% distinct(donor) %>% nrow()` left once we remove donors that have no Ferritin or no Hb measurement for any donation event. 


```{r , warning=FALSE}
#FINDONOR
## Assign all blood donors to FINDONOR cohort
indiv_donations_data$Cohort <- c("FINDONOR")

## Get sex for each blood donor
blood_data_summary <- indiv_donations_data %>% 
  group_by(donor) %>% 
  summarise(Sex=first(gender)) 

## Get values for first study donation with the required measurements donation (regardless of donation type)
## We remove events with no Ferritin and hb_v to be inline with from when the nb of previous donations were counted.
blood_values_init <- indiv_donations_data %>% 
  filter(!is.na(Ferritin) & !is.na(Hb_v) ) %>% 
  group_by(donor) %>% 
  filter(date == min(date)) %>% 
  dplyr::select(age, Ferritin, Hb_v, CRP, DaysToPreviousFB, donor, Cohort) %>% 
  ungroup()


blood_data_summary <- blood_data_summary %>% 
  full_join(blood_values_init, by = "donor") 


blood_data_summary <- donor_demo %>%
  inner_join(blood_data_summary, by = "donor") %>%
  mutate(all_study_FB_donation_count = NonFinnDonorDonationCount_FB + YesFinnDonorDonationCount_FB) 

#FINRISK97 
## Assign all FinRisk97 paprticipants to FINRISK97 cohort
fr1997$Cohort <- c("FINRISK97")

## Change term for sex from binary 1,2 to men, women
fr1997$SUKUP <- gsub("1", "Men", fr1997$SUKUP)
fr1997$SUKUP <- gsub("2", "Women", fr1997$SUKUP)

## Get sex for each participant
fr1997_summary <- fr1997 %>%
  filter(!is.na(FERRITIN)) %>%
  group_by(RELEASE_ID) %>%
  summarise(Sex=first(SUKUP))

## remove n/a:s for ferritin (NOT FOR HB!!!! We only have data from about 400 participants, all male) 
fr1997_remove_na <- fr1997 %>%
  filter(!is.na(FERRITIN)) %>%
  group_by(RELEASE_ID) %>%
  dplyr::select(IKA, FERRITIN, HGB, CRP, RELEASE_ID, K129, BMI, KY100_22, Q69, KY163, GRAVID, Cohort) %>%
  ungroup()

fr1997_summary <- fr1997_summary %>%
  full_join(fr1997_remove_na, by = "RELEASE_ID")

#HEALTH2000
## Assign all Health2000 participants to HEALTH2000 cohort

h2000$Cohort <- c("HEALTH00")

## Change term for sex from binary 1,2 to men, women
h2000$SP2 <- gsub("1", "Men", h2000$SP2)
h2000$SP2 <- gsub("2", "Women", h2000$SP2)  

## Get sex for each participant
h2000_summary <- h2000 %>%
  filter(!is.na(FERRITIINI)) %>%
  group_by(RELEASE_ID) %>%
  summarise(Sex=first(SP2))

## remove n/a:s for ferritin and Hb
  h2000_remove_na <- h2000 %>%
    filter(!is.na(FERRITIINI) & !is.na(B_Hb)) %>%
    group_by(RELEASE_ID) %>%
    dplyr::select(IKA2, FERRITIINI, B_Hb, CRP, RELEASE_ID, BD03, BMII_BMI.x, FB05, synnytys, BD07, Cohort) %>%
    ungroup()
  
  h2000_summary <- h2000_summary %>%
    full_join(h2000_remove_na, by = "RELEASE_ID")

  
# Rename variables to maintain cohesion between the cohorts
## Participant ID 
names(blood_data_summary)[names(blood_data_summary) == "donor"] <- "ID"
names(fr1997_summary)[names(fr1997_summary) == "RELEASE_ID"] <- "ID"
names(h2000_summary)[names(h2000_summary) == "RELEASE_ID"] <- "ID"
## Age
names(fr1997_summary)[names(fr1997_summary) == "IKA"] <- "Age"
names(h2000_summary)[names(h2000_summary) == "IKA2"] <- "Age"
## Hemoglobin
names(blood_data_summary)[names(blood_data_summary) == "Hb_v"] <- "Hb"
names(fr1997_summary)[names(fr1997_summary) == "HGB"] <- "Hb"
names(h2000_summary)[names(h2000_summary) == "B_Hb"] <- "Hb"
## Ferritin
names(fr1997_summary)[names(fr1997_summary) == "FERRITIN"] <- "Ferritin"
names(h2000_summary)[names(h2000_summary) == "FERRITIINI"] <- "Ferritin"
## Menstruation
names(blood_data_summary)[names(blood_data_summary) == "QR79"] <- "Menstruation"
names(fr1997_summary)[names(fr1997_summary) == "K129"] <- "Menstruation"
names(h2000_summary)[names(h2000_summary) == "BD03"] <- "Menstruation"
## BMI
names(h2000_summary)[names(h2000_summary) == "BMII_BMI.x"] <- "BMI"
## Smoking
names(blood_data_summary)[names(blood_data_summary) == "QR54"] <- "Smoking"
names(fr1997_summary)[names(fr1997_summary) == "Q69"] <- "Smoking"
names(h2000_summary)[names(h2000_summary) == "FB05"] <- "Smoking"
## Red meat
names(blood_data_summary)[names(blood_data_summary) == "QR40"] <- "RedMeat"
names(fr1997_summary)[names(fr1997_summary) == "KY100_22"] <- "RedMeat"


# ## Iron supplements !!!! STILL MISSING THIS PART, THE CODING FOR IRON SUPPLEMENTATION DOESN'T ADD UP WITH THE VARIABLE CODES IN EXCEL !!!!
# names(blood_data_summary)[names(blood_data_summary) == "QR36"] <- "GivenIronSupplements"
# names(blood_data_summary)[names(blood_data_summary) == "QR37"] <- "TookIronSupplements"


## History of childbirth
names(blood_data_summary)[names(blood_data_summary) == "QR83"] <- "PreviousChildbirth"
names(fr1997_summary)[names(fr1997_summary) == "KY163"] <- "PreviousChildbirth"
names(h2000_summary)[names(h2000_summary) == "synnytys"] <- "PreviousChildbirth"
## Current pregnancy
names(fr1997_summary)[names(fr1997_summary) == "GRAVID"] <- "CurrentPregnancy"
names(h2000_summary)[names(h2000_summary) == "BD07"] <- "CurrentPregnancy"


```

## Demographic group specification and assignment

Assign women to period/no period groups

FINDONOR:
There are `r blood_data_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>%  nrow()` women with no  answer to the question regarding their menstrual status. Of these, `r blood_data_summary %>% filter(sex == "Women" & is.na(Menstruation) & age > 65) %>%  nrow()` are older than 65. We impute these to  no period so they can be included in the post-menopausal women. Other women are all under 55, thus their menstrual status cannot be ascertained. 
Do you still have regular periods? (1=reg, 2=irreg, 3=no)

FINRISK97:
There are `r fr1997 %>% filter(Sex == "Women" & is.na(Menstruation)) %>%  nrow()` women with no  answer to the question regarding their menstrual status. Of these, `r blood_data_summary %>% filter(Sex == "Women" & is.na(Menstruation) & age > 65) %>%  nrow()` are older than 65. We impute these to  no period so they can be included in the post-menopausal women. Other women are all under 55, thus their menstrual status cannot be ascertained. 
Do you still menstruate? (1=reg, 2=irreg, 3=no)

HEALTH2000:
There are `r h2000 %>% filter(Sex == "Women" & is.na(Menstruation)) %>%  nrow()` women with no  answer to the question regarding their menstrual status (do you have periods nowadays?). Of these, `r blood_data_summary %>% filter(Sex == "Women" & is.na(Menstruation) & age > 65) %>%  nrow()` are older than 65. We impute these to  no period so they can be included in the post-menopausal women. Other women are all under 55, thus their menstrual status cannot be ascertained.
BD03 = do you have periods nowadays? (1=reg, 2=irreg, 3=no)

```{r}
# FINDONOR
older_findonor_donors <- filter(blood_data_summary, 
                       Sex == "Women" & is.na(Menstruation) & Age > 55)$ID

blood_data_summary <- blood_data_summary %>% 
  mutate(Menstruation = ifelse(ID %in% older_findonor_donors, "no_period", as.character(Menstruation))) 

#FINRISK97
older_fr1997 <- filter(fr1997_summary, 
                       Sex == "Women" & is.na(Menstruation) & Age > 55)$ID

fr1997_summary <- fr1997_summary %>% 
  mutate(Menstruation = ifelse(ID %in% older_fr1997, "no_period", as.character(Menstruation)))

# my own old code for referece, this is likely wrong
# older_fr1997 <- fr1997_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>%  nrow()
# 
# fr1997_summary <- fr1997_summary %>% 
#   mutate(Menstruation = ifelse(ID %in% older_fr1997, "3", as.character(Menstruation))) 

#HEALTH2000
older_h2000 <- filter(h2000_summary,
                      Sex == "Women" & is.na(Menstruation) & Age > 55)$ID

h2000_summary <- h2000_summary %>% 
  mutate(Menstruation = ifelse(ID %in% older_h2000, "no_period", as.character(Menstruation)))

# my own old code for referece, this is likely wrong
# older_h2000 <- h2000_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>%  nrow()
# 
# h2000_summary <- h2000_summary %>% 
#   mutate(Menstruation = ifelse(ID %in% older_h2000, "3", as.character(Menstruation))) 



```


Female study participants with no menstruation response after imputation are removed: (FinDonor: `r blood_data_summary %>% filter(sex == "Women" & is.na(Menstruation)) %>% nrow()` donors.)


```{r}
#FINDONOR
findonor_nb_women_no_menstruation_response <- blood_data_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>% nrow()


## remove female donors with no menstruation response
blood_data_summary_final <- blood_data_summary %>% 
  mutate(mens_ok_blood = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & !is.na(Menstruation)  ~ "Women",
    TRUE ~ "NA")) %>% 
  filter(mens_ok_blood != "NA") 

## We define the women's groups: 
findonor_n_women_removed <- blood_data_summary_final %>% 
  mutate(Group = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & (Menstruation == "no_period" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "no_period" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "irregular_period" | Menstruation == "regular_period") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>% 
  group_by(Group) %>% 
  filter(Group ==  "Women_pre_menop_no_mens") %>% 
  nrow()


#FINRISK97
finrisk_nb_women_no_menstruation_response <- fr1997_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>% nrow()


## remove female FinRisk97 participants with no menstruation response
fr1997_summary_final <- fr1997_summary %>% 
  mutate(mens_ok_fr1997 = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & !is.na(Menstruation)  ~ "Women",
    TRUE ~ "NA")) %>% 
  filter(mens_ok_fr1997 != "NA") 

## We define the women's groups:
fr1997_n_women_removed <- fr1997_summary_final %>%
  mutate(Group = case_when(
    Sex == "Men" ~ "Men",
    Sex == "Women" & (Menstruation == "3" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "3" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "1" | Menstruation == "2") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>%
  group_by(Group) %>%
  filter(Group ==  "Women_pre_menop_no_mens") %>%
  nrow()

#HEALTH2000
h2000_nb_women_no_menstruation_response <- h2000_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>% nrow()


## remove female Health2000 participants with no menstruation response
h2000_summary_final <- h2000_summary %>% 
  mutate(mens_ok_h2000 = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & !is.na(Menstruation)  ~ "Women",
    TRUE ~ "NA")) %>% 
  filter(mens_ok_h2000 != "NA") 

## We define the women's groups:
h2000_n_women_removed <- h2000_summary_final %>%
  mutate(Group = case_when(
    Sex == "Men" ~ "Men",
    Sex == "Women" & (Menstruation == "3" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "3" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "1" | Menstruation == "2") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>%
  group_by(Group) %>%
  filter(Group ==  "Women_pre_menop_no_mens") %>%
  nrow()


```

We now define the following groups:

+ pre-menopausal: regular or irregular menstruation reported
+ post-menopausal: no menstruation reported and age equal to > 45
+ donors younger than 45 and no reported menstruation are excluded (`r n_women_removed` donors)



```{r}
#FINDONOR:
## We define the women's groups: 
blood_data_summary_final <- blood_data_summary_final %>% 
  mutate(Group = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & (Menstruation == "no_period" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "no_period" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "irregular_period" | Menstruation == "regular_period") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>% 
  filter(Group != "Women_pre_menop_no_mens" & Group != "NA") %>% 
  droplevels() %>% 
  mutate(Group = ordered(Group, levels =  c("Pre_menopause_women", "Post_menopause_women", "Men"))) 


#FINRISK97:
## We define the women's groups: 
fr1997_summary_final <- fr1997_summary_final %>% 
  mutate(Group = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & (Menstruation == "3" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "3" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "1" | Menstruation == "2") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>% 
  filter(Group != "Women_pre_menop_no_mens" & Group != "NA") %>% 
  droplevels() %>% 
  mutate(Group = ordered(Group, levels =  c("Pre_menopause_women", "Post_menopause_women", "Men"))) 

#HEALTH2000:
## We define the women's groups: 
h2000_summary_final <- h2000_summary_final %>% 
  mutate(Group = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & (Menstruation == "3" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "3" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "1" | Menstruation == "2") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>% 
  filter(Group != "Women_pre_menop_no_mens" & Group != "NA") %>% 
  droplevels() %>% 
  mutate(Group = ordered(Group, levels =  c("Pre_menopause_women", "Post_menopause_women", "Men"))) 


```

In order to later merge the three cohorts into one table, we need to make colums for the variables the different cohorts don't take into consideration. 

```{r}
# We make new columns for donation data for the general population cohorts
fr1997_summary_final$TwoYearsFromStartCount_FB <- 0
h2000_summary_final$TwoYearsFromStartCount_FB <- 0

# We then  assign a randomized donation interval for 3-6 years for the general population cohorts (at this point ferritin levels out, see Lobier et al.)
fr1997_summary_final <- fr1997_summary_final %>% mutate(DaysToPreviousFB = round(runif(nrow(fr1997_summary_final), min = 3*365, max = 6*365) ,0))
h2000_summary_final <- h2000_summary_final %>% mutate(DaysToPreviousFB = round(runif(nrow(h2000_summary_final), min = 3*365, max = 6*365) ,0))

# Add columns for the variables that are missing from the cohorts (CurrentPregnancy for FinDonor, RedMeat for Health2000 and the both iron supplement variables for both the general population cohorts) and add values to them (no for pregnancy in FinDonor, n/a for the rest)
blood_data_summary_final$CurrentPregnancy <- 0
fr1997_summary_final$GivenIronSupplements <- NA
fr1997_summary_final$TookIronSupplements <- NA
h2000_summary_final$GivenIronSupplements <- NA
h2000_summary_final$TookIronSupplements <- NA
h2000_summary_final$RedMeat <- NA
```

## Remove cohort participants with missing data

### Donation history

We remove `r blood_data_summary_final %>% filter(is.na(DaysToPreviousFB)) %>% nrow()`, that is donors who have not donated previously (they are missing the nb of days since last FB donation variable). As there is no information on blood donation in the general population cohorts this is not done for them. 

```{r}
new_donors_data <- blood_data_summary_final %>% 
  filter(is.na(DaysToPreviousFB)) %>% 
    mutate(group = case_when(Group == "Pre_menopause_women" ~ "Pre-menopausal women",
                   Group == "Post_menopause_women" ~ "Post-menopausal women",
                   Group == "Men" ~ "Men" ,
                   TRUE~ "NA"),
         Group = ordered(Group)) 

blood_data_summary_final <- blood_data_summary_final %>% 
  drop_na(DaysToPreviousFB)
```

### BMI

We remove `r cohort_name %>% filter(is.na(BMI)) %>% nrow()` participants for whom we do not have the BMI data:

```{r}
#### FinDonor
nb_removed_blood <- findonor_nb_women_no_menstruation_response +
                            blood_data_summary_final %>% filter(is.na(BMI)) %>% nrow() 

blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!is.na(BMI))

#### FinRisk97
nb_removed_fr97 <- finrisk_nb_women_no_menstruation_response +
                            fr1997_summary_final %>% filter(is.na(BMI)) %>% nrow() 

fr1997_summary_final <- fr1997_summary_final %>% 
  filter(!is.na(BMI))

#### Health2000
nb_removed_h2000 <- h2000_nb_women_no_menstruation_response +
                            h2000_summary_final %>% filter(is.na(BMI)) %>% nrow() 

h2000_summary_final <- h2000_summary_final %>% 
  filter(!is.na(BMI))
```

### Smoking

We remove `r cohort_name %>% filter(is.na(Smoking)) %>% nrow()` donors  who  did not answer the smoking question:

```{r}
#### FinDonor
nb_removed_blood <- nb_removed_blood +
                            blood_data_summary_final %>% filter(is.na(Smoking)) %>% nrow() 

blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!is.na(Smoking))

#### FinRisk97
##### NOTE: REMOVING N/As FOR SMOKING REMOVES >3000 OBSERVATIONS 
nb_removed_fr97 <- nb_removed_fr97 +
                            fr1997_summary_final %>% filter(is.na(Smoking)) %>% nrow() 

fr1997_summary_final <- fr1997_summary_final %>% 
  filter(!is.na(Smoking))

#### Health2000
nb_removed_h2000 <- nb_removed_h2000 +
                            h2000_summary_final %>% filter(is.na(Smoking)) %>% nrow() 

h2000_summary_final <- h2000_summary_final %>% 
  filter(!is.na(Smoking))


```

### Previous childbirth

We remove `r cohort_name %>% filter((group != "Men" & is.na(QR83)))%>% nrow()` female donors who did not answer the question about childbirth: 

```{r}
#### FinDonor
nb_removed_blood <- nb_removed_blood +
                            blood_data_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth))) %>% nrow 

blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!(Group != "Men" & is.na(PreviousChildbirth)))

#### FinRisk97
nb_removed_fr97 <- nb_removed_fr97 +
                            fr1997_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth))) %>% nrow 

fr1997_summary_final <- fr1997_summary_final %>% 
  filter(!(Group != "Men" & is.na(PreviousChildbirth)))

#### Health2000
nb_removed_h2000 <- nb_removed_h2000 +
                            h2000_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth))) %>% nrow 

h2000_summary_final <- h2000_summary_final %>% 
  filter(!(Group != "Men" & is.na(PreviousChildbirth)))

```

### Current pregnancy

We remove `r cohort_name %>% filter((group != "Men" & is.na(QR83)))%>% nrow()` female donors who did not answer the pregnancy question: 

```{r}
#### FinDonor
nb_removed_blood <- nb_removed_blood +
                            blood_data_summary_final %>% filter((Group != "Men" & is.na(CurrentPregnancy))) %>% nrow 

blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!(Group != "Men" & is.na(CurrentPregnancy)))

#### FinRisk97
nb_removed_fr97 <- nb_removed_fr97 +
                            fr1997_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth))) %>% nrow 

fr1997_summary_final <- fr1997_summary_final %>% 
  filter(!(Group != "Men" & is.na(PreviousChildbirth)))

#### Health2000
nb_removed_h2000 <- nb_removed_h2000 +
                            h2000_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth))) %>% nrow 

h2000_summary_final <- h2000_summary_final %>% 
  filter(!(Group != "Men" & is.na(PreviousChildbirth)))
```


### Iron supplementation 

We remove `r blood_data_summary_final %>% mutate(iron_comp_c = ifelse(iron_supp == FALSE, 0, iron_comp_c )) %>%  filter(is.na(iron_comp_c)) %>% nrow()` donors that did not answer properly to both questions if applicable.

For the modelling, we impute a 0 (no supplementation) to iron_comp_c when the donor reports not being offered iron supplementation.

This is not done for the general population cohorts, as they were not asked about iron supplementation and as such all observation are input as n/a. 

```{r}
nb_removed_blood <- nb_removed_blood +
                            blood_data_summary_final %>% 
                              mutate(iron_comp_c = ifelse(iron_supp == FALSE, 0, iron_comp_c )) %>% 
                              filter(is.na(iron_comp_c)) %>% 
                              nrow() 

blood_data_summary_final <- blood_data_summary_final %>% 
  mutate(iron_comp_c = ifelse(iron_supp == FALSE, 0, iron_comp_c )) %>% 
  filter(!is.na(iron_comp_c))
```

### Red meat intake 
      
RedMeat = how often do you eat meat?
NOTE: THIS NEEDS TO BE REDONE, THE CODING FOR MEAT INTAKE DIFFERS BETWEEN FINDONOR/FINRISK 

FinDonor: 1=can't say, 2=none, 3=once a week or less, 4=1-3 times a week, 5=4-6 times a week, 6=once a day
FinRisk97: 1=more seldom than once a month or not at all, 2) once or twice a month, 3) once a week 4) twice a week or more often, 5) almost daily, 6) daily 7) several times a day


```{r message=FALSE, warning=FALSE}

# donors_to_remove <- blood_data_summary_final %>% 
#   dplyr::select(ID, Group, RedMeat) %>% 
#   gather(key= question, value = answer, -ID,-Group) %>% 
#   filter(is.na(answer)) %>% 
#   dplyr::select(Group, ID) %>% 
#   distinct(ID, Group) 
# 
# 
# blood_data_summary_final <- blood_data_summary_final %>% 
#   filter(!ID %in% donors_to_remove$Donor) 
# 
# nb_removed_blood <- nb_removed_blood +
#                             donors_to_remove %>% nrow()


```

### Total number of donors removed for missing questionnaire answers

The total number of donors removed because of missing questionnaire data is:
`r nb_removed_blood`
`r nb_removed_fr97`
`r nb_removed_h2000`






## Remove donors with extreme physiological measures

As decided previously, we remove data according the following criteria:

* BMI > 50
* CRP > 30
* Ferritin > 400 

This amounts to `r cohort_name %>% filter(BMI >= 50 | CRP >= 30 | Ferritin >= 400) %>% nrow()` donors that are removed. 


```{r}
# FinDonor
blood_data_summary_final <- blood_data_summary_final %>% 
   filter(BMI < 50 & CRP < 30 & Ferritin < 400)

# FinRisk97
fr1997_summary_final <- fr1997_summary_final %>% 
   filter(BMI < 50 & CRP < 30 & Ferritin < 400)

#Health2000
h2000_summary_final <- h2000_summary_final %>% 
   filter(BMI < 50 & CRP < 30 & Ferritin < 400)

```

## Final group N for the three cohorts

```{r}
# FinDonor
blood_data_summary_final %>% 
   mutate(Group = dplyr::recode(Group, Pre_menopause_women = "Pre-menopausal women",
         Post_menopause_women = "Post-menopausal women")) %>% 
  group_by(Group) %>% 
  summarise ( N = n()) %>% 
  kable() 

# FinRisk97
fr1997_summary_final %>% 
   mutate(Group = dplyr::recode(Group, Pre_menopause_women = "Pre-menopausal women",
         Post_menopause_women = "Post-menopausal women")) %>% 
  group_by(Group) %>% 
  summarise ( N = n()) %>% 
  kable() 

# Health2000
h2000_summary_final %>% 
   mutate(Group = dplyr::recode(Group, Pre_menopause_women = "Pre-menopausal women",
         Post_menopause_women = "Post-menopausal women")) %>% 
  group_by(Group) %>% 
  summarise ( N = n()) %>% 
  kable() 

 
```


## Merging cohorts into one table 

For the code to work all three cohorts need to be merged into one single table. Before this can be done variables the variables selected need to have the same name.  

```{r}
### NOTE: DATA ON IRON SUPPLEMENTS ARE STILL MISSING DUE TO DIFFERENCE IN CODING IN THE EXCEL FILE, ADDED COLUMN WITH N/A DATA JUST TO TRY THIS OUT, REMOVE AND FIX CODE WHEN I'VE FIGURED OUT WHAT THE CORRECT CODE IS
blood_data_summary_final$TookIronSupplements <- NA
blood_data_summary_final$GivenIronSupplements <- NA
  
### mutate smoking and pregnancy part - is this correct? Should this be done for the other variables that aren't working currently? 
blood_data_summary_final <- blood_data_summary_final %>% 
  dplyr::select(ID, Group, Sex, Age, TwoYearsFromStartCount_FB, DaysToPreviousFB,
                Ferritin, Hb, BMI, Menstruation, Smoking, RedMeat, GivenIronSupplements, TookIronSupplements, PreviousChildbirth, 
                CurrentPregnancy, Cohort) %>% 
    mutate(Smoking = ifelse(Smoking == "no", "no", "yes"),
         Smoking = ordered(Smoking, levels = c( "no",  "yes")))

### mutate smoking and pregnancy part - is this correct?
fr1997_summary_final <- fr1997_summary_final %>% 
  dplyr::select(ID, Group, Sex, Age, TwoYearsFromStartCount_FB, DaysToPreviousFB,
                Ferritin, Hb, BMI, Menstruation, Smoking, RedMeat, GivenIronSupplements, TookIronSupplements, PreviousChildbirth, 
                CurrentPregnancy, Cohort)
       mutate(Smoking = ifelse(Smoking == "no", "no", "yes"),
       Smoking = ordered(Smoking, levels = c( "no",  "yes")),
       CurrentPregnancy = ordered(CurrentPregnancy, levels = c( "no",  "yes")))

### mutate smoking and pregnancy part - is this correct?
h2000_summary_final <- h2000_summary_final %>% 
  dplyr::select(ID, Group, Sex, Age, TwoYearsFromStartCount_FB, DaysToPreviousFB,
                Ferritin, Hb, BMI, Menstruation, Smoking, RedMeat, GivenIronSupplements, TookIronSupplements, PreviousChildbirth, 
                CurrentPregnancy, Cohort)
         mutate(Smoking = ifelse(Smoking == "no", "no", "yes"),
         Smoking = ordered(Smoking, levels = c( "no",  "yes")),
         CurrentPregnancy = ordered(CurrentPregnancy, levels = c( "no",  "yes")))


# merge the cohorts

summary_all_cohorts <- rbind(blood_data_summary_final, fr1997_summary_final, h2000_summary_final)



  

```

## Final group N for the summarized table

```{r}
summary_all_cohorts %>% 
   mutate(Group = dplyr::recode(Group, Pre_menopause_women = "Pre-menopausal women",
         Post_menopause_women = "Post-menopausal women")) %>% 
  group_by(Group) %>% 
  summarise ( N = n()) %>% 
  kable() 
 
```


## Recoding of variables

We use a linear scale from 1 to 4 for red meat consumption and mapped the 6 possible response the 4 codes. 

* red meat:
    * "never" ~ 1,
    * "less_than_once_weekly" ~ 2,
    * "1.3_week" or "4.6_week"~ 3,
    * "daily" or "several_daily" ~ 4,
    
    NOTE 1: DIFFERENT CODING IN FINDONOR REPORT, WHY IS THIS DIFFERENT? 1=can't say, 2=none, 3=once a week or less, 4=1-3 times a week, 
    5=4-6 times a week, 6=once a day
    NOTE 2: FINRISK97 USES A DIFFERENT CONSUMPTION SCALE, THIS NEEDS TO BE RECODED: 1=more seldom than once a month or not at all, 2) once
    or twice a month, 3) once a week 4) twice a week or more often, 5) almost daily, 6) daily 7) several times a day


* Smoking:
    * "no" ~"no"
    * "sometimes" or "daily"  ~ "yes"
   
   NOTE: DIFFERENT CODING EARLIER, WHICH ONE SHOULD BE USED? 
   
* QR83: have you given birth ?

```{r}
# Recode food and non-alcoholic beverages
meat <- summary_all_cohorts %>% 
  dplyr::select(ID, RedMeat) %>% 
  gather(key = key, value =  value, -ID) %>% 
  mutate(linear_value = case_when(value == "never" ~ 1,
    value == "less_than_once_weekly" ~ 2,
    value %in% c("1.3_week" , "4.6_week") ~ 3,
    TRUE   ~ 4 )) %>% # "daily" | "several_daily"
  dplyr::select(-value) %>% 
  mutate(key = paste(key, "n", sep = "_")) %>% 
  spread(key = key, value = linear_value)

summary_all_cohorts <- summary_all_cohorts %>% 
  full_join(meat, by = "donor") 

```

# Results - Regression analyses

## Regressions on sTfR and ferritin  

We analyzed the data with multiple linear regression. We first ran Ordinary Least Square (OLS) regressions and their diagnostics. If the diagnostics identified problematic observations (e.g., outliers, influential observations), we also ran robust regression regressions [@yu2017robust]. Robust regression is based on less assumptions than OLS regression and is less sensitive to influential observations. It  gives more accurate estimates of regression coefficients in the presence of problematic observations. Using robust regression leads to a more principled approach to regression analysis and avoids potential confounds introduced by selective removal of influential observations. As the rlm command from the MASS package does not compute directly the *t* and *p* values, we used the f.robftest function from the sfsmisc package. Practical examples of using robust regression can be found in the literature [@davison2017platelet].  


In addition, we used relative importance analysis to estimate the average proportion of variance in the outcome variable explained by each co-variate [@Groemping2006]. Relative iportance was computed using the pmvd method which assigns 0 to the relative importance of a regressor if it is non significant in the complete model. .

The regressors were entered as follows in the regression models for sTfR and ferritin:

* Continuous variables were entered as continuous predictors
    * age, number of donations in last two years, nb of days since last donation, CRP, BMI
  
* Binary variables were entered as categorical (but coded as numerical for relative importance):
    * Smoking, pregnancy
  
* Ordinal variables were coded on a linear scale and entered as continuous [(Reference)](https://www3.nd.edu/~rwilliam/stats3/OrdinalIndependent.pdf)
    * Iron supplementation 
    * All dietary intake
  
We compute both coefficients and standardized coefficients (with all dependent and independent variables standardized) for robust regression and only coefficients for OLS regression. We present robust coefficients and their bootstrapped BCa 95% confidence intervals in Table 1 (ferritin) and  Table 2 (sTfR), robust standardized coefficients, their bootstrapped distribution and BCa 95% confidence intervals in Figure 2.A and non-standardized OLS regression coefficients in Supplementary Table 1 (Ferritin) and 2 (sTfR). 

## Pre-processing

### Rename and transform

```{r}
regression_cohorts <- summary_all_cohorts %>% 
    rename(don_ct = TwoYearsFromStartCount_FB,
            last_don = DaysToPreviousFB,
            iron = iron_comp_c) %>% 
     mutate(don_ct_2 = don_ct^2,
            log_ferr = log(Ferritin),
            log_last_don = log(last_don)/log(2),
            log_CRP = log(CRP)) 

```

Data transformations:

* Log transformed variables: 
    * Ferritin
    * sTfR
    * CRP 
  
* Age is divided by 5 to simplify coeff interpretation
* Number of days before donation is transformed as $log\_last\_don = log(last\_don)/log(2)$ to help with the interpretation of coefficients. 
    * An increase in one of the transformed variable is equivalent to a doubling of the number of days since last donation.

* Standardization:
    * Standardized coefficients: 
        * All dependent and independent variables were standardized
      
    * Coefficients:
        * All dependent variables entered as continuous varaibles are centered but not scaled. 

## Pre-menopausal women
### Data pre-processing

We center all variables entered as continuous.

```{r}
# Center variables

test_data_std <- blood_data_summary_reg %>% 
  filter(group == "Pre_menopause_women") %>% 
  dplyr::select(donor ,red_meat_n) %>% 
  gather(key = key, value = value, -donor) %>% 
  group_by(key) %>% 
  mutate(value = scale(value, scale = FALSE)[,1]) %>% 
  spread(key = key, value = value)
  
test_data_std <- blood_data_summary_reg %>% 
  filter(group == "Pre_menopause_women") %>% 
  dplyr::select(donor, log_ferr, TransferrinR, age, log_CRP, BMI, don_ct,log_last_don,
                smoking, pregnancy, iron, log_sTfR) %>%
  mutate(age = age / 5, 
         age = scale(age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         don_ct = scale(don_ct, scale = FALSE)[,1],
         log_last_don = scale(log_last_don, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         iron = scale(iron, scale = FALSE)[,1],
         don_ct_2 = don_ct^2) %>% 
  full_join(test_data_std, by = "donor")  %>% 
  dplyr::select(-donor)

```


### Mutiple regression - ferritin as outcome

#### Correlograms

```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "age", "log_CRP",  "don_ct", "don_ct_2",
                    "log_last_don"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "pregnancy",  "red_meat_n", "smoking", "iron", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```


# Correlogram for age

```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "age"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram log CRP 

```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "log_CRP"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram donation count 

```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "don_ct"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram donation count 2 

```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "don_ct_2"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram log last donation

```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "log_last_don"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram pregnancy (QR83=have you given birth? yes/no)

```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "pregnancy"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram red meat

```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "red_meat_n"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram smoking 

```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "smoking"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram iron supplements 

```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "iron"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram BMI

```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```








# PART 3: LOGISTICAL REGRESSION






<!-- ################ START OF FITTING FOR LOGISTICAL REGRESSION, SOME PARTS DUPLICATED IN LATER CODE!!!! -->


<!-- # Summary -->

<!-- This file includes the code for computing the regression results regarding the predictors of improving of self-related health during the FinDonor study for the explorative analysis parts. Models containing as many predictor variables as possible are fitted to data and confidence intervals for coefficients computed with bootstrapping. -->

<!-- # Load the data -->

<!-- We load the data from hypothesis based regressions and initial data analysis and combine them together.  -->

<!-- ```{r loading} -->
<!-- #Load questionnaire data recoded to change -->
<!-- load(file = "../results/changedata.rdata")  -->


<!-- #Load data used in hypothesis regressions -->
<!-- load(file = "../results/hypregdata.rdata") -->


<!-- #We cannot know if our "Reduced","Stable","Increased" factor levels are equally spaced. Hence make the variables unordered. -->
<!-- #changedata %>% mutate_at(funs(ordered), l1) %>% mutate_each_(funs(as.numeric), l2) -->
<!-- changedata <- changedata %>% mutate_if(is.ordered, function(x){factor(x,ordered = FALSE)}) -->
<!-- regression_data_men <- inner_join(changedata,reg_data$men,by=c("donor"="donor")) -->
<!-- regression_data_post_menop_women <- inner_join(changedata,reg_data$womenpost,by=c("donor"="donor")) -->
<!-- regression_data_pre_menop_women <- inner_join(changedata,reg_data$womenpre,by=c("donor"="donor")) -->


<!-- #Later in this script we find that smoking behaviour can not be fitted so they are dropped now not to loose data over them.  -->
<!-- regression_data_men <- regression_data_men %>%  dplyr::select(-smoking_behavior,-health_outcome_neg) -->
<!-- regression_data_post_menop_women <- regression_data_post_menop_women %>% dplyr::select(-smoking_behavior,-health_outcome_neg) -->
<!-- regression_data_pre_menop_women <- regression_data_pre_menop_women %>%  dplyr::select(-smoking_behavior,-health_outcome_neg) -->
<!-- #Bringing in extra data will bring some NAs. -->
<!-- regression_data_men <-  droplevels(na.omit(regression_data_men)) -->
<!-- regression_data_post_menop_women <- droplevels(na.omit(regression_data_post_menop_women))  -->
<!-- regression_data_pre_menop_women <- droplevels(na.omit(regression_data_pre_menop_women)) -->

<!-- #Sleep night needs to be scaled -->
<!-- #Later in this script we find that sleep_night_d can not be fitted -->
<!-- # regression_data_men <- regression_data_men %>% mutate(sleep_night_d = scale(sleep_night_d, scale = FALSE)[,1])  -->
<!-- # regression_data_post_menop_women <- regression_data_post_menop_women %>% mutate(sleep_night_d = scale(sleep_night_d, scale = FALSE)[,1]) -->
<!-- # regression_data_pre_menop_women <- regression_data_pre_menop_women %>% mutate(sleep_night_d = scale(sleep_night_d, scale = FALSE)[,1]) -->


<!-- #Fuse back together for later use "Pre_menopause_women", "Post_menopause_women","Men" and join with unscaled data -->
<!-- m<- regression_data_men %>% mutate(group = "Men") -->
<!-- e<- regression_data_post_menop_women %>% mutate(group = "Post_menopause_women") -->
<!-- o <- regression_data_pre_menop_women %>% mutate(group = "Pre_menopause_women") -->
<!-- temp <- bind_rows(m,e,o) %>%  dplyr::select(-age,-BMI_diff,-donation_frequency,-Hb_beginning,-log_Ferritin_beginning,-starting_weight) -->
<!-- file <- "../results/hypregdata_unscaled.rdata" -->
<!-- load(file=file) -->
<!-- regression_data <- regression_data %>% dplyr::select(age,BMI_diff,donation_frequency,Hb_beginning,log_Ferritin_beginning, Ferritin_beginning,starting_weight,donor) %>% inner_join(temp,by=c("donor"="donor")) #%>% select(-anemia_hist_d,-high_hb_hist_d,-education,-smoking_behavior) -->
<!-- regression_data %>% group_by(group) %>% count() -->

<!-- ``` -->

<!-- ```{r} -->
<!-- regression_data %>% ungroup() %>% count() -->
<!-- ``` -->

<!-- # Table 1 type summary -->
<!-- ```{r} -->
<!-- #exclude variables that are not used in modelling -->
<!-- table1data <- regression_data %>% ungroup() %>% dplyr::select(#-age, -->
<!--   #-starting_weight, -->
<!--   #-BMI_diff,  -->
<!--   #-log_Ferritin_beginning, -->
<!--   #-Hb_beginning, -->
<!--   #-donation_frequency, -->
<!--   -health_very_good, -->
<!--   -health_excellent, -->
<!--   -high_hb_hist_d, -->
<!--   -health_outcome, -->
<!--   -donor, -->
<!--   #-high_hb_hist_d, -->
<!--   #-sleep_night_d, -->
<!--   #-vita_iron_d, -->
<!--   #-ment_inf_d, -->
<!--   #-education, -->
<!--   #-smoking_behavior -->
<!--   -health_poor, -->
<!--   -health_satisfactory -->
<!-- ) %>% mutate( -->
<!--   starting_weight=starting_weight*5, -->
<!--   log_Ferritin_beginning= exp((log(log_Ferritin_beginning)/log(2)) * log(2)), -->
<!--   Hb_beginning = Hb_beginning *5 -->
<!-- ) %>%  -->
<!--   rename( -->
<!--     Age = age, -->
<!--     "Full time employed" = "employment", -->
<!--     "Inital weight (kg)" =starting_weight, -->
<!--     "BMI difference (second - first)"=BMI_diff, -->
<!--     "Initial ferritin (ug/l)"=Ferritin_beginning, -->
<!--     "Initial hemoglobin (g/l)" =Hb_beginning, -->
<!--     "Donation frequency (yearly)"=donation_frequency, -->
<!--     "Anemia detected"=anemia_hist_d, -->
<!--     "Elevated cholesterol detected" =  "cholesterol_d", -->
<!--     "Elevated blood sugar detected"="blood_sugar_d" ,  -->
<!--     "High blood pressure detected"="HBP_d", -->
<!--     "Low haemoglobin detected"="low_hb_hist_d", -->
<!--     "FRCBS iron supplementation"="iron_supp_d",  -->
<!--     "Frequency multivitamine"="vita_multi_d", -->
<!--     "Frequency of anti-inflammatory"="anti_inf_d", -->
<!--     "Frequency of red meat"="freq_red_meat_d", -->
<!--     "Frequency of fruits"="freq_fruits_d", -->
<!--     "Frequency of juices"="freq_juices_d", -->
<!--     "Frequency of whole milk"="freq_wholem_d", -->
<!--     "Frequency of milk"="freq_milk_d", -->
<!--     "Frequency of coffee"="freq_cof_d", -->
<!--     "Frequency of tea"="freq_tea_d", -->
<!--     "Frequency of alcohol"="freq_alc_d", -->
<!--     "Physical symptoms inferred"="phys_inf_d", -->
<!--     "Physical condition" = "phys_cond_d", -->
<!--     "Ability for heavy activity" = "act_heavy_d", -->
<!--     "Typical day activity" = "day_act_d", -->
<!--     "Hours for light activity" = "h_act_light_d", -->
<!--     "Enough sleep" = "sleep_enough_d", -->
<!--     "Tired during day" = "sleep_tired_d", -->
<!--     "Frequency of feeling calm" = "freq_calm_d", -->
<!--     "Quality of life" = "life_qual" -->
<!--   ) -->

<!-- myVars<- colnames(table1data %>% dplyr::select(-group)) -->
<!-- non_normal_vars <- c("Initial ferritin (ug/l)") -->
<!-- summary_table <- CreateTableOne(data = table1data,vars=myVars, strata = "group",test = FALSE) -->

<!-- tab3Mat <- print(summary_table, nonnormal = non_normal_vars,vars=myVars, quote = FALSE, noSpaces = TRUE, printToggle = FALSE) -->

<!-- tab3Mat %>%  -->
<!--   kable() -->
<!--   write.csv(tab3Mat, file = paste0("../results/table_1_populationExp_",gsub("-","",as.Date(Sys.time())),".csv")) -->

<!-- #Does not work    -->
<!-- #openxlsx::write.xlsx(tab3Mat, file = "../results/table_1_populationExp_20200512.xlsx",rowNames =TRUE,borders="surrounding",keepNA=TRUE) -->
<!-- ``` -->

<!-- ## Percentage of iron depleted -->

<!-- ```{r} -->
<!-- regression_data %>%  dplyr::select(group, Ferritin_beginning) %>% mutate(Iron_depleted = Ferritin_beginning < 15) %>%  group_by(group,Iron_depleted) %>% count() %>% group_by(group)%>% summarise( -->
<!--   Iron_depleted = Iron_depleted, -->
<!--   n= n, -->
<!--   Percentage = round(n / sum(n),2) -->
<!-- ) -->
<!-- ``` -->


<!-- # Premenopausal women -->

<!-- Check that final data has some variation -->
<!-- ```{r} -->
<!-- summary(regression_data_pre_menop_women) -->
<!-- ``` -->

<!-- All explenatory variables selected in inital_data_analysis.Rmd were tried,  but could not be used because of problems in model fitting. These were then left out and also removed from the data assembly process in the inital_data_analysis.Rmd not to loose data because of these unusable variables. The variables left out are specified in code below for each group. -->

<!-- ```{r} -->
<!-- logit_pre_menop_women <- glm(health_outcome ~ age + starting_weight + BMI_diff +  -->
<!--                                log_Ferritin_beginning + Hb_beginning + -->
<!--                               donation_frequency + health_very_good + health_excellent + -->
<!--                               cholesterol_d +  -->
<!--                               blood_sugar_d +  -->
<!--                                HBP_d +  -->
<!--                                 low_hb_hist_d +  -->
<!--                                anemia_hist_d +  -->
<!-- #                               high_hb_hist_d + # The following terms are causing separation among the sample points: high_hb_hist_dTRUE -->
<!--                                iron_supp_d +  -->
<!--                                #sleep_night_d + # #fails in bootstrapping -->
<!--                                freq_calm_d +  -->
<!--                                freq_red_meat_d +  -->
<!--                                freq_fruits_d +  -->
<!--                                freq_juices_d +  -->
<!--                                freq_wholem_d +  -->
<!--                                freq_milk_d +  -->
<!--                                freq_cof_d +  -->
<!--                                freq_tea_d +  -->
<!--                                phys_cond_d +  -->
<!--                                act_heavy_d +  -->
<!--                                sleep_enough_d +  -->
<!--                                day_act_d +  -->
<!--                                h_act_light_d +  -->
<!--                                vita_multi_d +  -->
<!--                                #vita_iron_d +  #fails in bootstrapping -->
<!--                                sleep_tired_d +  -->
<!--                                anti_inf_d +  -->
<!--                                phys_inf_d +  -->
<!--                                #ment_inf_d +  #fails in bootstrapping -->
<!--                                freq_alc_d + -->
<!--                                life_qual + -->
<!--                                #education  +  #fails in bootstrapping -->
<!--                                employment   -->
<!--                              #smoking_behavior  #fails in bootstrapping -->
<!--                              , -->
<!--                              data=regression_data_pre_menop_women, -->
<!--                             family="binomial", separation = 'test') -->

<!--  summary(logit_pre_menop_women) -->
<!-- ``` -->

<!-- cholesterol_d + blood_sugar_d + HBP_d + low_hb_hist_d + anemia_hist_d + high_hb_hist_d + iron_supp_d + sleep_night_d + freq_calm_d + freq_red_meat_d + freq_fruits_d + freq_juices_d + freq_wholem_d + freq_milk_d + freq_cof_d + freq_tea_d + phys_cond_d + act_heavy_d + sleep_enough_d + day_act_d + h_act_light_d + vita_multi_d + vita_iron_d + sleep_tired_d + anti_inf_d + phys_inf_d + ment_inf_d + freq_alc_d  -->

<!-- ```{r} -->
<!-- plot(logit_pre_menop_women) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- vif(logit_pre_menop_women) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- logistic.display(logit_pre_menop_women) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- tidy_logit_pre_menop_women <- -->
<!--   logit_pre_menop_women %>%  -->
<!--   tidy %>%  -->
<!--   mutate(OR = exp(estimate), -->
<!--          group = "Pre_menopause_women") %>% mutate(p.value = p.value * 2) -->
<!-- #p-values are multiplied as we analyse the data twice -->
<!-- ``` -->


<!-- # Post menopausal women -->

<!-- ```{r} -->
<!-- summary(regression_data_post_menop_women) -->
<!-- ``` -->

<!-- There are no cases who aquired high Hb during the research period. high_hb_hist_d should be removed. -->

<!-- ```{r} -->
<!-- logit_post_menop_women <- glm(health_outcome ~ age + starting_weight + BMI_diff +  -->
<!--                                log_Ferritin_beginning + Hb_beginning + -->
<!--                               donation_frequency + health_very_good + health_excellent + -->
<!--                                cholesterol_d +  -->
<!--                               # blood_sugar_d + # Separation exists among the sample points. -->
<!--                                HBP_d +  -->
<!--                                low_hb_hist_d +  -->
<!--                               anemia_hist_d +  -->
<!--                                # high_hb_hist_d + # no postive cases  -->
<!--                                iron_supp_d +  -->
<!--                                #sleep_night_d + #fails in bootstrapping -->
<!--                                freq_calm_d +  -->
<!--                                freq_red_meat_d +  -->
<!--                                freq_fruits_d +  -->
<!--                                freq_juices_d +  -->
<!--                                freq_wholem_d +  -->
<!--                                freq_milk_d +  -->
<!--                                freq_cof_d +  -->
<!--                                freq_tea_d +  -->
<!--                                phys_cond_d +  -->
<!--                                act_heavy_d +  -->
<!--                                sleep_enough_d +  -->
<!--                                day_act_d +  -->
<!--                                h_act_light_d +  -->
<!--                                vita_multi_d +  -->
<!--                                #vita_iron_d + #eparation exists among the sample points. -->
<!--                                sleep_tired_d +  -->
<!--                                anti_inf_d +  -->
<!--                                phys_inf_d +  -->
<!--                                #ment_inf_d + #fails in bootstrapping -->
<!--                                freq_alc_d + -->
<!--                                life_qual + -->
<!--                                #education + #fails in bootstrapping -->
<!--                                employment  -->
<!--                                #smoking_behavior #fails in bootstrapping -->
<!--                              , -->
<!--                              data=regression_data_post_menop_women, -->
<!--                             family="binomial", separation = 'test') -->

<!--  summary(logit_post_menop_women) -->
<!-- ``` -->

<!-- cholesterol_d + blood_sugar_d + HBP_d + low_hb_hist_d + anemia_hist_d + high_hb_hist_d + iron_supp_d + sleep_night_d + freq_calm_d + freq_red_meat_d + freq_fruits_d + freq_juices_d + freq_wholem_d + freq_milk_d + freq_cof_d + freq_tea_d + phys_cond_d + act_heavy_d + sleep_enough_d + day_act_d + h_act_light_d + vita_multi_d + vita_iron_d + sleep_tired_d + anti_inf_d + phys_inf_d + ment_inf_d + freq_alc_d  -->

<!-- ```{r} -->
<!-- plot(logit_post_menop_women) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- vif(logit_post_menop_women) -->
<!-- ``` -->
<!-- High VIF values indicate that for post menopausal women the analysis is not reliable as there is much colinearity. -->

<!-- ```{r} -->
<!-- logistic.display(logit_post_menop_women) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- tidy_logit_post_menop_women <- -->
<!--   logit_post_menop_women %>%  -->
<!--   tidy %>%  -->
<!--   mutate(OR = exp(estimate), -->
<!--          group = "Post_menopause_women") %>% mutate(p.value = p.value * 2) -->
<!-- #p-values are multiplied as we analyse the data twice``` -->



<!-- # Men -->

<!-- ```{r} -->
<!-- summary(regression_data_men) -->
<!-- ``` -->

<!-- There are only 2 cases who aquired high Hb during the research period and it was not used in models with women. high_hb_hist_d should be removed. -->

<!-- ```{r} -->
<!-- logit_men <- glm(health_outcome ~ age + starting_weight + BMI_diff +  -->
<!--                                log_Ferritin_beginning + Hb_beginning + -->
<!--                               donation_frequency + health_very_good + health_excellent + -->
<!--                              cholesterol_d +  -->
<!--                                blood_sugar_d +  -->
<!--                                HBP_d +  -->
<!--                                low_hb_hist_d +  -->
<!--                               # anemia_hist_d + # eparation exists among the sample points. -->
<!--                               # high_hb_hist_d +  -->
<!--                                iron_supp_d +  -->
<!--                               # sleep_night_d + #fails in bootstrapping -->
<!--                                freq_calm_d +  -->
<!--                                freq_red_meat_d +  -->
<!--                                freq_fruits_d +  -->
<!--                                freq_juices_d +  -->
<!--                                freq_wholem_d +  -->
<!--                                freq_milk_d +  -->
<!--                                freq_cof_d +  -->
<!--                                freq_tea_d +  -->
<!--                                phys_cond_d +  -->
<!--                                act_heavy_d +  -->
<!--                                sleep_enough_d +  -->
<!--                                day_act_d +  -->
<!--                                h_act_light_d +  -->
<!--                                vita_multi_d +  -->
<!--                                #vita_iron_d +  -->
<!--                                sleep_tired_d +  -->
<!--                                anti_inf_d +  -->
<!--                                phys_inf_d +  -->
<!--                                #ment_inf_d + #fails in bootstrapping  -->
<!--                                freq_alc_d + -->
<!--                                life_qual + -->
<!--                               # education + #fails in bootstrapping  -->
<!--                                employment  -->
<!--                               # smoking_behavior #fails in bootstrapping -->
<!--                              , -->
<!--                              data=regression_data_men, -->
<!--                           #    data=test_data_std, -->
<!--                             family="binomial", separation = 'test') -->

<!--  summary(logit_men) -->


<!-- ``` -->

<!-- cholesterol_d + blood_sugar_d + HBP_d + low_hb_hist_d + anemia_hist_d + high_hb_hist_d + iron_supp_d + sleep_night_d + freq_calm_d + freq_red_meat_d + freq_fruits_d + freq_juices_d + freq_wholem_d + freq_milk_d + freq_cof_d + freq_tea_d + phys_cond_d + act_heavy_d + sleep_enough_d + day_act_d + h_act_light_d + vita_multi_d + vita_iron_d + sleep_tired_d + anti_inf_d + phys_inf_d + ment_inf_d + freq_alc_d  -->

<!-- ```{r} -->
<!-- plot(logit_men) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- vif(logit_men) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- logistic.display(logit_men) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- tidy_logit_men <- -->
<!--   logit_men %>%  -->
<!--   tidy %>%  -->
<!--   mutate(OR = exp(estimate), -->
<!--          group = "Men") %>% mutate(p.value = p.value * 2) -->
<!-- #p-values are multiplied as we analyse the data twice -->
<!-- ``` -->



################ END OF FITTING FOR LOGISTICAL REGRESSION






################ THIS PART WE WILL USE, START
  
```{r}
test_data_std <- test_data_std %>% 
  mutate(irondef = #ferritin variable name < 15
           ) #makes a variable of ID
logit_pre_menopausal <- glm(
  irondef ~ BMI +
    age,
    data=test_data_std,
    family="binomial", separation = 'test')
 
 summary(logit_pre_menopausal)


#https://stats.stackexchange.com/questions/11109/how-to-deal-with-perfect-separation-in-logistic-regression
#https://stats.stackexchange.com/questions/45803/logistic-regression-in-r-resulted-in-perfect-separation-hauck-donner-phenomenon
```
  

```{r}
plot(logit_pre_menopausal)
```

```{r}
vif(logit_pre_menopausal)
```

```{r}
logistic.display(logit_pre_menopausal)
```
  
############### END OF PART WE WILL USE 






  
# Bootstraps
https://data.princeton.edu/wws509/r/c3s1
https://stats.stackexchange.com/questions/304833/how-to-calculate-odds-ratio-and-95-confidence-interval-for-logistic-regression
https://www.andrewheiss.com/blog/2016/04/25/convert-logistic-regression-standard-errors-to-odds-ratios-with-r/

```{r}
detach("package:safeBinaryRegression", unload=TRUE)
#safeBinaryRegression messes up booting
```

## Define auxiliarry functions for bootstrapping
```{r}

# https://www.painblogr.org/2017-10-18-purrring-through-bootstraps.html


get_coefficients <- function(data, boot_ind){
  fit <- glm (health_outcome~.,
              data[boot_ind,],
              family = "binomial")
#  print(names(coef(fit)))
  return(coef(fit))
}



compute_Bca_CI <-function(fit_boot,conf = 0.95){
  cat("compute_Bca_CI with conf: ",conf,"\n")
  Bca_inf = rep(0, length(fit_boot$t0))
  Bca_sup = rep(0, length(fit_boot$t0))
  for (i_regressor in 1:length(fit_boot$t0)){
    CI <- boot.ci(fit_boot, type = "bca", index=i_regressor,conf = conf)
    Bca_inf[i_regressor] <- CI$bca[4]
    Bca_sup[i_regressor] <- CI$bca[5]
  }
  return(tibble(Bca_inf,Bca_sup, regressor = names(fit_boot$t0)) %>% 
           filter(regressor != "(Intercept)"))
}


get_bootstrap <- function(data, nb_boot){
 cat(nrow(data),"get_bootstrap 1\n")
 

    fit_boot <- boot(data, statistic= get_coefficients, R = nb_boot)

 cat(nrow(data),"get_bootstrap 2\n")
  fit_boot_distrib <- as.tibble(fit_boot$t)  
  names(fit_boot_distrib) <- names(fit_boot$t0)

  fit_boot_Bca <- compute_Bca_CI(fit_boot)
  fit_boot_Bca_2 <- compute_Bca_CI(fit_boot,conf=0.80)
  cat(nrow(data),"get_bootstrap 3\n")
  return(list(fit_boot_distrib,fit_boot_Bca,fit_boot_Bca_2))
}


```

## Run bootstrapping
```{r}
# Error in t.star[r, ] <- res[[r]] : 
#   number of items to replace is not a multiple of replacement length
#can sometimes be fixed by just adding bootstrap, but sometimes requires dropping out a variable


get_bootstrap_coeffs <- function(regression_data, current_group, nb_boot )
{

  ## preprocess and standardize data
  
    test_data_std <- regression_data %>% 
      filter(group == current_group) %>% 
      dplyr::select(donor, age, starting_weight, BMI_diff,
                    log_Ferritin_beginning, Hb_beginning, 
                    donation_frequency) %>% 
      gather(key = key, value = value, -donor) %>% 
      group_by(key) %>% 
      mutate(value_scale = scale(value, scale = FALSE)[,1]) %>% 
      dplyr::select(-value) %>% 
      spread(key = key, value = value_scale) 
     
    if (current_group == "Men") {
      test_data_std <- inner_join(test_data_std,regression_data %>% 
                                    dplyr::select(donor, health_outcome, health_very_good, health_excellent,
                                                #  sleep_night_d, ##Error in t.star[r, ] <- res[[r]] :
                                                  cholesterol_d,
                                                  blood_sugar_d,
                                                  HBP_d,
                                                  low_hb_hist_d,
                                                  iron_supp_d,
                                                  freq_calm_d,
                                                  freq_red_meat_d,
                                                  freq_fruits_d,
                                                  freq_juices_d,
                                                  freq_wholem_d,
                                                  freq_milk_d,
                                                  freq_cof_d,
                                                  freq_tea_d,
                                                  phys_cond_d,
                                                  act_heavy_d,
                                                  sleep_enough_d,
                                                  day_act_d,
                                                  h_act_light_d,
                                                  vita_multi_d,
                                                  #vita_iron_d, #Error in t.star[r, ] <- res[[r]]
                                                  sleep_tired_d,
                                                  anti_inf_d,
                                                  phys_inf_d,
                                                  #  ment_inf_d,
                                                  freq_alc_d,
                                                  life_qual,
                                                  employment
                                                #  education # Error in t.star[r, ] <- res[[r]] : 
                                                # smoking_behavior #Error in t.star[r, ] <- res[[r]] :
                                                  
                                    ),
                                  by = "donor") %>% 
        dplyr::select(-donor) 
    }
    
    if (current_group == "Pre_menopause_women") {
      test_data_std <- inner_join(test_data_std,regression_data %>% 
                                    dplyr::select(donor, health_outcome, health_very_good, health_excellent,
                                                #  sleep_night_d, #Error in t.star[r, ] <- res[[r]] :
                                                  cholesterol_d,
                                                  blood_sugar_d,
                                                  HBP_d,
                                                  low_hb_hist_d,
                                                  iron_supp_d,
                                                  freq_calm_d,
                                                  freq_red_meat_d,
                                                  freq_fruits_d,
                                                  freq_juices_d,
                                                  freq_wholem_d,
                                                  freq_milk_d,
                                                  freq_cof_d,
                                                  freq_tea_d,
                                                  phys_cond_d,
                                                  act_heavy_d,
                                                  sleep_enough_d,
                                                  day_act_d,
                                                  h_act_light_d,
                                                  vita_multi_d,
                                                  #vita_iron_d, #Error in t.star[r, ] <- res[[r]]
                                                  sleep_tired_d,
                                                  anti_inf_d,
                                                  phys_inf_d,
                                                  # ment_inf_d,
                                                  freq_alc_d,
                                                  life_qual,
                                                  employment
                                                #  education #Error in t.star[r, ] <- res[[r]] : 
                                               #   smoking_behavior # Error in t.star[r, ] <- res[[r]] : 
                                                  
                                    ),
                                  by = "donor") %>% 
        dplyr::select(-donor) 
    }
    
      if (current_group == "Post_menopause_women") {
      test_data_std <- inner_join(test_data_std,regression_data %>% 
                                    dplyr::select(donor, health_outcome, health_very_good, health_excellent,
                                                #  sleep_night_d, #Error in t.star[r, ] <- res[[r]] :
                                                 cholesterol_d,
                                                 blood_sugar_d,
                                                  HBP_d,
                                                  low_hb_hist_d,
                                                  iron_supp_d,
                                                  freq_calm_d,
                                                  freq_red_meat_d,
                                                  freq_fruits_d,
                                                  freq_juices_d,
                                                  freq_wholem_d,
                                                  freq_milk_d,
                                                  freq_cof_d,
                                                  freq_tea_d,
                                                  phys_cond_d,
                                                  act_heavy_d,
                                                  sleep_enough_d,
                                                  day_act_d,
                                                  h_act_light_d,
                                                  vita_multi_d,
                                                  #vita_iron_d #Error in t.star[r, ] <- res[[r]]
                                                  sleep_tired_d,
                                                  anti_inf_d,
                                                  phys_inf_d,
                                                  # ment_inf_d # Error in t.star[r, ] <- res[[r]]
                                                  freq_alc_d,
                                                  life_qual, # Error in bca.ci(boot.out, conf, index[1L], L = L, t = t.o, t0 = t0.o,  : estimated adjustment 'w' is infinite -> works if 10k boots.
                                                  employment
                                                #  education # Error in bca.ci(boot.out, conf, index[1L], L = L, t = t.o, t0 = t0.o,  :
                                                #  smoking_behavior # Error in t.star[r, ] <- res[[r]]
                                                  
                                    ),
                                  by = "donor") %>% 
        dplyr::select(-donor) 
    }
  cat(nrow(test_data_std),current_group,"get_bootstrap_coeffs 1\n")
  result <- test_data_std %>% 
  get_bootstrap(nb_boot = nb_boot)

  cat(nrow(test_data_std),current_group,"get_bootstrap_coeffs 2\n")
  bootstrap_distrib <- result[[1]] %>% 
    gather(key = regressor, value = coefficient) %>% 
    filter(regressor != "(Intercept)") %>% 
    mutate(group = current_group)
      cat(nrow(test_data_std),current_group,"get_bootstrap_coeffs 3\n")
  Bca_CI  <- result[[2]] %>% mutate(group = current_group)
  Bca_CI_2  <- result[[3]] %>% mutate(group = current_group)
  cat(nrow(test_data_std),current_group,"get_bootstrap_coeffs 4\n")
  return(list(bootstrap_distrib,Bca_CI,Bca_CI_2))
}




#get_from_file_ferr = TRUE



set.seed(125)
#group_str = "Pre_menopause_women"

for(group_str in c("Pre_menopause_women","Men","Post_menopause_women")){
#  for(group_str in c("Pre_menopause_women")) {
#for(group_str in c("Post_menopause_women","Pre_menopause_women")){
  output_file_distrib = paste0("../results/bootstraps/exp_coeff_boot_distrib_seed_125_ctr_strata_final_",group_str,".rdata")
  output_file_Bca = paste0("../results/bootstraps/exp_coeff_boot_Bca_seed_125_ctr_strata_final_",group_str,".rdata")
  output_file_Bca_2 = paste0("../results/bootstraps/exp_coeff_boot_Bca_2_seed_125_ctr_strata_final_",group_str,".rdata")
  get_from_file_ferr <- all(file.exists(output_file_distrib),file.exists(output_file_Bca),file.exists(output_file_Bca_2) )
  if (get_from_file_ferr){
    cat("Loading boots from file for group ",group_str,"\n")
    stuff <- list()
    load(file=output_file_distrib)
    load(file=output_file_Bca)
    load(file=output_file_Bca_2)
    stuff[[1]] <- temp1  
    stuff[[2]] <- temp2
    stuff[[3]] <- temp3
  }else {
    stuff <- get_bootstrap_coeffs(regression_data, group_str, nb_boot = 10000)
    #         stuff <- get_bootstrap_coeffs(regression_data, group_str, nb_boot = 1000)
    temp1 <- stuff[[1]] 
    temp2 <- stuff[[2]] 
    temp3 <- stuff[[3]] 
    save(temp1,file=output_file_distrib) 
    save(temp2,file=output_file_Bca)
    save(temp3,file=output_file_Bca_2)
  }
  
  if(!exists("bootstrap_distrib")){
    cat("initializing bootstrap_distrib, bootstrap_Bca_CI_2 & bootstrap_Bca_CI\n")

    bootstrap_distrib <- stuff[[1]]
    cat("nrow(bootstrap_distrib)",nrow(bootstrap_distrib),"\n")
    bootstrap_Bca_CI <- stuff[[2]]
    cat("nrow(bootstrap_Bca_CI)",nrow(bootstrap_Bca_CI),"\n")
    bootstrap_Bca_CI_2 <- stuff[[3]]
    cat("nrow(bootstrap_Bca_CI_2)",nrow(bootstrap_Bca_CI_2),"\n")
  }else
  {
    cat("bootstrap_distrib & bootstrap_Bca_CI already exist\n")
    bootstrap_distrib<-bind_rows(bootstrap_distrib, stuff[[1]])
    bootstrap_Bca_CI<-bind_rows(bootstrap_Bca_CI, stuff[[2]])
    bootstrap_Bca_CI_2<-bind_rows(bootstrap_Bca_CI_2, stuff[[3]])
  }
  
}


```

```{r}
bootstrap_Bca_CI %>%  group_by(group) %>% count()

```

# Forest plots

```{r}
# "\U0394 BMI"

regressor_values <- c(
 "age" = "Age",
   "BMI_diff" = "BMI difference",
   "donation_frequency" = "Donation frequency",
   "Hb_beginning" = "Initial Hemoglobin",
   "log_Ferritin_beginning" ="Initial Ferritin",
   "starting_weight" ="Initial weight",
   "health_very_good" ="Initial health rating \n Very good vs Satisfactory/Good",
    "health_excellent" = "Initial health rating \n Excellent vs Satisfactory/Good",
 
 
 "cholesterol_dTRUE" = "Elevated cholesterol detected",
"blood_sugar_dTRUE" = "Elevated blood sugar detected", 
"HBP_dTRUE" = "High blood pressure detected",
"low_hb_hist_dTRUE" = "Low haemoglobin detected",
"iron_supp_dStable" = "FRCBS iron supplementation stable", 
"iron_supp_dIncreased" = "FRCBS iron supplementation increased",
"vita_multi_dStable" = "Frequency multivitamine stable",
"vita_multi_dIncreased" = "Frequency multivitamine increased",
#"vita_iron_dStable" = "Frequency iron supplements stable",
#"vita_iron_dIncreased" = "Frequency iron supplements increased",
"anti_inf_dStable" = "Frequency of anti-inflammatory stable",
"anti_inf_dIncreased" = "Frequency of anti-inflammatory increased",
"freq_red_meat_dStable" = "Frequency of red meat stable",
"freq_red_meat_dIncreased" = "Frequency of red meat increased",
"freq_fruits_dStable" = "Frequency of fruits stable",
"freq_fruits_dIncreased" = "Frequency of fruits increased",
"freq_juices_dStable" = "Frequency of juices stable",
"freq_juices_dIncreased" = "Frequency of juices increased",
"freq_wholem_dStable" = "Frequency of whole milk stable",
"freq_wholem_dIncreased" = "Frequency of whole milk increased",
"freq_milk_dStable" = "Frequency of milk stable",
"freq_milk_dIncreased" = "Frequency of milk increased",
"freq_cof_dStable" = "Frequency of coffee stable",
"freq_cof_dIncreased"  = "Frequency of coffee increased",
"freq_tea_dStable" = "Frequency of tea stable",
"freq_tea_dIncreased" = "Frequency of tea increased",
"freq_alc_dStable" = "Frequency of alcohol stable",
"freq_alc_dIncreased" = "Frequency of alcohol increased",
"phys_inf_dStable" = "Physical symptoms inferred stable",
"phys_inf_dIncreased" = "Physical symptoms inferred increased",
"phys_cond_dStable" = "Physical condition stable",
"phys_cond_dIncreased" = "Physical condition increased",
"act_heavy_dStable" = "Ability for heavy activity stable",
"act_heavy_dIncreased" = "Ability for heavy activity increased",
"day_act_dStable"  = "Typical day activity stable",
"day_act_dIncreased" = "Typical day activity increased",
"h_act_light_dStable"  = "Hours for light activity stable",
"h_act_light_dIncreased"  = "Hours for light activity increased",
"sleep_enough_dStable" = "Enough sleep stable",
"sleep_enough_dIncreased"  = "Enough sleep increased",
"sleep_tired_dStable" = "Tired during day stable",
"sleep_tired_dIncreased" = "Tired during day increased",
"freq_calm_dStable" = "Frequency of feeling calm stable",
"freq_calm_dIncreased" =  "Frequency of feeling calm increase",
"life_qualgood" = "Quality of life good",
"life_qualvery_good" = "Quality of life very good",
"employmentTRUE" = "Full time employed"
)
```



```{r}
temp <-
  tidy_logit_men %>% 
  bind_rows(tidy_logit_post_menop_women) %>% 
  bind_rows(tidy_logit_pre_menop_women) %>% 
  dplyr::select(group, term, OR, p.value) %>% 
  rename(regressor = term) %>% 
  inner_join(bootstrap_Bca_CI,
             by =  c("regressor", "group")) %>% 
  filter(regressor != "(Intercept)") %>% 
  mutate(Bca_inf = exp(Bca_inf),
         Bca_sup = exp(Bca_sup),
         is_sig = p.value < 0.05) %>%
  mutate(
    regressor = plyr::revalue(regressor, regressor_values),
    regressor = ordered(regressor,
                        levels =  as.character(regressor_values)),
    regressor = fct_rev(regressor),
    group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women",
                      group == "Post_menopause_women" ~ "Post-menopausal women",
                      TRUE ~ "Men"),
    group = ordered(group, levels = c(  "Pre-menopausal women",  "Post-menopausal women", "Men"  )
    )
  )

temp2 <-
  tidy_logit_men %>% 
  bind_rows(tidy_logit_post_menop_women) %>% 
  bind_rows(tidy_logit_pre_menop_women) %>% 
  dplyr::select(group, term, OR, p.value) %>% 
  rename(regressor = term) %>% 
  inner_join(bootstrap_Bca_CI_2,
             by =  c("regressor", "group")) %>% 
  filter(regressor != "(Intercept)") %>% 
  mutate(Bca_inf = exp(Bca_inf),
         Bca_sup = exp(Bca_sup),
         is_sig = p.value < 0.05) %>%
  mutate(
    regressor = plyr::revalue(regressor, regressor_values),
    regressor = ordered(regressor,
                        levels =  as.character(regressor_values)),
    regressor = fct_rev(regressor),
    group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women",
                      group == "Post_menopause_women" ~ "Post-menopausal women",
                      TRUE ~ "Men"),
    group = ordered(group, levels = c(  "Pre-menopausal women",  "Post-menopausal women", "Men"  )
    )
  )
    

```

```{r}
bootstrap_Bca_CI <- temp
plotdata <- bootstrap_distrib %>% 
  mutate(regressor = plyr::revalue(regressor, regressor_values),
  regressor = ordered(regressor,
                  levels = as.character(regressor_values)),
  regressor = fct_rev(regressor),
  group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women",
    group == "Post_menopause_women" ~ "Post-menopausal women",
    TRUE ~ "Men"),
  group = ordered(group, levels = c( "Men",  "Post-menopausal women", "Pre-menopausal women" ) )) %>% 
  mutate(OR = exp(coefficient))

bootstrap_Bca_CI_2 <- temp2

file <- "../results/exp_bootstrap_Bca_CI.csv"
temp <- bootstrap_Bca_CI %>% mutate_if(is.numeric, round, digits = 3) %>% dplyr::select(regressor,group,p.value,OR,Bca_inf,Bca_sup) %>% arrange(desc(regressor),group)
write.csv(temp ,file = file,row.names = FALSE)

```

#Small plot of filtered variables for main paper
```{r fig1, fig.height = 10, fig.width = 15}
#set the forest plot OR scale lims
orlimR <- 1000
orlimL <- 1/orlimR
pointrange2size <- 1.4
#Filter for those that are sig in some:
bootstrap_Bca_CI_filtered <- bootstrap_Bca_CI %>% group_by(regressor) %>% mutate(some_sig= any(is_sig)) %>% filter(some_sig == TRUE)
#Set values on axes limits to facilitate plotting
bootstrap_Bca_CI_filtered$Bca_sup[bootstrap_Bca_CI_filtered$Bca_sup > orlimR] <- orlimR
bootstrap_Bca_CI_filtered$Bca_sup[bootstrap_Bca_CI_filtered$Bca_sup == 0] <- orlimR
bootstrap_Bca_CI_filtered$Bca_inf[bootstrap_Bca_CI_filtered$Bca_inf < orlimL] <- orlimL

#Filter for those that are sig in some:
bootstrap_Bca_CI_filtered_2 <- bootstrap_Bca_CI_2 %>% group_by(regressor) %>% mutate(some_sig= any(is_sig)) %>% filter(some_sig == TRUE)
#Set values on axes limits to facilitate plotting
bootstrap_Bca_CI_filtered_2$Bca_sup[bootstrap_Bca_CI_filtered_2$Bca_sup > orlimR] <- orlimR
bootstrap_Bca_CI_filtered_2$Bca_sup[bootstrap_Bca_CI_filtered_2$Bca_sup == 0] <- orlimR
bootstrap_Bca_CI_filtered_2$Bca_inf[bootstrap_Bca_CI_filtered_2$Bca_inf < orlimL] <- orlimL


plotdata_filtered <- plotdata %>% filter(plotdata$regressor %in% bootstrap_Bca_CI_filtered$regressor) 


p  <- ggplot(aes(x = regressor, y = OR, color = group, fill = group),data = plotdata_filtered) +
  geom_hline(yintercept = 1 , color = "grey", linetype = "dotted",
             size = 1) +
  # geom_violin(alpha = 0.15, trim = TRUE, position = position_dodge(width = 0.8), size = 0.1) +
  geom_pointrange(data = bootstrap_Bca_CI_filtered,
                  aes(ymin=Bca_inf, ymax = Bca_sup, shape = is_sig, color = group, fill = group),
                  position = position_dodge(width = 0.7), size = 1, linetype = 1) +
  # geom_pointrange(data = bootstrap_Bca_CI_filtered_2,
  #                 aes(ymin=Bca_inf, ymax = Bca_sup, shape = is_sig, color = group, fill = group),
  #                 position = position_dodge(width = 0.7), size = pointrange2size) +
  # # geom_point(data = bootstrap_Bca_CI, 
  #            aes(shape = is_sig, fill = group),
  #            position = position_dodge(width = 0.8), size = 2) +
    scale_color_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                     limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
     scale_fill_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                     limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
 scale_shape_manual(values = c(1,16)) +
  scale_y_log10(limits=c(orlimL,orlimR)) +
  # scale_x_discrete(labels=parse(text=unique(bootstrap_Bca_CI$regressor)))
  # ylim(1/35, 35) +
  guides(shape = FALSE) +
  theme_classic() +
  # theme(legend.position = c(0.8,0.9),
  #       legend.title = element_blank(),
  #       legend.box = "vertical",
  #       legend.text = element_text(size = 20),
  #       plot.title =  element_text(size = 24,
  #                                 hjust = 0),
  #       panel.grid.minor.y = element_blank(),
  #       axis.line = element_line(colour="black"),
  #       axis.title.y = element_blank(),
  #       axis.title.x = element_text(size = 24),
  #       axis.text = element_text(size = 22),
  #       panel.grid.major.y = element_blank()) +
  coord_flip() +
  xlab("Robust standardized coefficient") +
  labs(title = "PREDICTORS OF WORSENED SELF-RATED HEALTH \nBETWEEN QUESTIONNAIRES",
       subtitle = "") +
   theme(axis.title = element_text(size = 14),
        strip.text =  element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.background = element_rect(fill = "white"),
        strip.background.x = element_blank(),
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(size = 14, angle = 30,hjust = 0.8),
        plot.title =  element_text(size = 16),
        plot.subtitle = element_text(size = 16,
                                     hjust = 0.5))

p

```

```{r}

p <- p +   labs(title = NULL,subtitle = NULL)
ggplot2::ggsave(filename = "../results/figures/exp_small_log_reg_results_b.pdf",
                plot= p,
# do not change these rather just scale the ready pdf later if you need.
       width = 35,
       height = 25,
       dpi = 600,
       units = "cm")

```


#Big plot of all variables for supplement
```{r fig2, fig.height = 40, fig.width = 15}

bootstrap_Bca_CI$Bca_sup[bootstrap_Bca_CI$Bca_sup > orlimR] <- orlimR
bootstrap_Bca_CI$Bca_sup[bootstrap_Bca_CI$Bca_sup == 0] <- orlimR
bootstrap_Bca_CI$Bca_inf[bootstrap_Bca_CI$Bca_inf < orlimL] <- orlimL

#Set values on axes limits to facilitate plotting
bootstrap_Bca_CI_2$Bca_sup[bootstrap_Bca_CI_2$Bca_sup > orlimR] <- orlimR
bootstrap_Bca_CI_2$Bca_sup[bootstrap_Bca_CI_2$Bca_sup == 0] <- orlimR
bootstrap_Bca_CI_2$Bca_inf[bootstrap_Bca_CI_2$Bca_inf < orlimL] <- orlimL


p2  <- ggplot(aes(x = regressor, y = OR, color = group, fill = group),data = plotdata) +
  geom_hline(yintercept = 1 , color = "grey", linetype = "dotted",
             size = 1) +
  # geom_violin(alpha = 0.15, trim = TRUE, position = position_dodge(width = 0.8), size = 0.1) +
  geom_pointrange(data = bootstrap_Bca_CI,
                  aes(ymin=Bca_inf, ymax = Bca_sup, shape = is_sig, color = group, fill = group),
                  position = position_dodge(width = 0.7), size = 1, linetype = 2) +
  geom_pointrange(data = bootstrap_Bca_CI_2,
                  aes(ymin=Bca_inf, ymax = Bca_sup, shape = is_sig, color = group, fill = group),
                  position = position_dodge(width = 0.7), size = pointrange2size) +
  # # geom_point(data = bootstrap_Bca_CI, 
  #            aes(shape = is_sig, fill = group),
  #            position = position_dodge(width = 0.8), size = 2) +
    scale_color_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                     limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
     scale_fill_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                     limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
 scale_shape_manual(values = c(1,16)) +
  scale_y_log10(limits=c(orlimL,orlimR)) +
  # scale_x_discrete(labels=parse(text=unique(bootstrap_Bca_CI$regressor)))
  # ylim(1/35, 35) +
  guides(shape = FALSE) +
  theme_classic() +
  # theme(legend.position = c(0.8,0.9),
  #       legend.title = element_blank(),
  #       legend.box = "vertical",
  #       legend.text = element_text(size = 20),
  #       plot.title =  element_text(size = 24,
  #                                 hjust = 0),
  #       panel.grid.minor.y = element_blank(),
  #       axis.line = element_line(colour="black"),
  #       axis.title.y = element_blank(),
  #       axis.title.x = element_text(size = 24),
  #       axis.text = element_text(size = 22),
  #       panel.grid.major.y = element_blank()) +
  coord_flip() +
  xlab("Robust standardized coefficient") +
  labs(title = "PREDICTORS OF WORSENED SELF-RATED HEALTH \nBETWEEN QUESTIONNAIRES",
       subtitle = "") +
   theme(axis.title = element_text(size = 14),
        strip.text =  element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.background = element_rect(fill = "white"),
        strip.background.x = element_blank(),
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(size = 14, angle = 30,hjust = 0.8),
        plot.title =  element_text(size = 16),
        plot.subtitle = element_text(size = 16,
                                     hjust = 0.5))

p2
```



```{r}
p2 <- p2 +   labs(title = NULL,subtitle = NULL)
ggplot2::ggsave(filename = "../results/figures/exp_big_log_reg_results_b.pdf",
                plot= p2,
       width = 35,
       height = 55,
       dpi = 600,
       units = "cm")

```


#Separate plots for individual variables

```{r}
fig.width=15
fig.height=13
#Load questionnaire data
load(file = "../data/final_data.rdata") 
final_data <- data

```

## Change in physical condition color
```{r}

pdata <- final_data %>% 
  dplyr::select(donor, group, health, questionnaire) %>% 
  drop_na(group) %>% 
  filter( group != "Women_pre_menop_no_mens") %>% 
  spread(key = questionnaire, value = health) %>% 
  filter(First != "NA" & Second != "NA") %>%  
  mutate(Difference = case_when(First < Second ~ "Improved",
                                First == Second ~ "Stable",
                                TRUE ~ "Worsened")) %>% 
  inner_join(regression_data, by = c("donor", "group")) %>% 
  mutate(group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal \n women",
                           group == "Post_menopause_women" ~ "Post-menopausal \n women",
                           TRUE ~ "Men"),
         group = ordered(group, levels = c(  "Men", "Post-menopausal \n women", "Pre-menopausal \n women") ))
 
p <- ggplot(aes(x= phys_cond_d , fill = group), data=pdata) +
  # geom_hline(yintercept = 15, color = "red", linetype = 3)
  geom_bar()+
  #geom_boxplot(outlier.alpha = 0) +
  #geom_beeswarm(alpha = 0.5, shape = 1) +
  # scale_y_log10() +
  facet_grid(group ~ Difference,scales="free_y") +
  scale_fill_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                    limits = c("Pre-menopausal \n women", "Post-menopausal \n women","Men" )) +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("Evolution of self-rated health") +
  xlab("Change in physical condition") +
  ylab("Count") +
  geom_text(stat='count', aes(label=..count..),color="black", vjust=1)  
#scale_y_log10() 



ggplot2::ggsave(plot=p,filename = "../results/figures/exp_phys_cond_c.pdf",
       width = fig.width,
       height = fig.height,
       dpi = 600,
       units = "cm")

p
```




## Quality of life
```{r}

p <- ggplot(aes(x= life_qual), data=pdata) +
  # geom_hline(yintercept = 15, color = "red", linetype = 3)
  geom_bar(aes(fill=group))+
  #geom_boxplot(outlier.alpha = 0) +
  #geom_beeswarm(alpha = 0.5, shape = 1) +
  # scale_y_log10() +
  facet_grid(group ~ Difference,scales="free_y") +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("Evolution of self-rated health")+
  xlab("Quality of life") +
  ylab("Count") +
  geom_text(stat='count', aes(label=..count..),color="black", vjust=1) +
  scale_fill_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c("Pre-menopausal \n women", "Post-menopausal \n women","Men" ))

ggplot2::ggsave(plot=p,filename = "../results/figures/exp_life_qual_c.pdf",
       width = fig.width,
       height = fig.height,
       dpi = 600,
       units = "cm")

p
```


