---
title: "Low ferritin model"
author: "Mikko Arvas"
date: "15.11.2021"
output: pdf_document

---


# Summary

Version 5: only include variables available for Dutch cohorts, BLOOD DONATION VARIABLES INCLUDED.

```{r setup, , message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
# echo "rmarkdown::render('low_ferritin_model_bayes_version2.Rmd', clean=TRUE,output_format='pdf_document',output_file='../results/low_ferritin_model_bayes/low_ferritin_model_bayes_version2.pdf')" | R --slave
#library(tidyverse)
library(brms)
library(bayesplot)
library(dplyr)
library(ggplot2)
```

PART 2: BAYESIAN LOGISTIC REGRESSION

# Data loading and preparation 

Load data

```{r}
load("../data/ID_data_regression_cohorts.rdata")
all_cohorts <- regression_cohorts

# all_cohorts <- all_cohorts %>% 
#   filter(!is.na(CRP)) %>% 
#   filter(! CRP >= 30)
```

# All cohorts together 

## Data
```{r}
# men
## menstruation and childbirth variables removed 
regression_data_men <- all_cohorts %>% 
  filter(Group == "Men") %>% 
  dplyr::select(Weight, ID, log_ferritin, Age, donation_count, log_last_donation, Cohort, iron_deficiency, Smoking, Group, blood_vol) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         blood_vol = scale(blood_vol, scale = FALSE)[,1],
         Weight = scale(Weight, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# premenopausal women
regression_data_pre <- all_cohorts %>% 
  filter(Group == "Pre_menopause_women") %>% 
  dplyr::select(Weight, ID, log_ferritin, Age, donation_count, log_last_donation, Cohort, iron_deficiency, Smoking, Group, blood_vol) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         blood_vol = scale(blood_vol, scale = FALSE)[,1],
         Weight = scale(Weight, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# postmenopausal womem 
## menstruation variable removed 
regression_data_post <- all_cohorts %>% 
  filter(Group == "Post_menopause_women") %>% 
  dplyr::select(Weight, ID, log_ferritin, Age, donation_count, log_last_donation, Cohort, iron_deficiency, Smoking, Group, blood_vol) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         blood_vol = scale(blood_vol, scale = FALSE)[,1],
         Weight = scale(Weight, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 


```

## Models 


```{r men}
#men
file <- "../results/bmodels/version4/logit_men_b"

logistic_men_b <- brm(iron_deficiency ~ 
                        Age + 
                        #BMI + 
                        #Smoking + 
                        #iron_complience + 
                        #donation_count_2 + 
                        #donation_count +
                        #log_last_donation + 
                        #log_CRP +
                        #Region + 
                        Weight +
                        #RedMeat_n +
                        blood_vol +
                        Cohort, 
                      data = regression_data_men, 
                      family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(logistic_men_b)

```

```{r  pre}
# premenopausal women
file <- "../results/bmodels/version4/logit_pre_b"

logistic_pre_b <- brm(iron_deficiency ~ 
                        Age + 
                        #BMI + 
                        #Smoking + 
                        #iron_complience + 
                        #donation_count_2 + 
                        #donation_count +
                        #log_last_donation + 
                        #log_CRP +
                        #Region + 
                        Weight +
                        #RedMeat_n +
                        #PreviousChildbirth +
                        #Menstruation +
                        blood_vol +
                        Cohort, 
                      data = regression_data_pre,
                      family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(logistic_pre_b)


```

```{r  post}
# postmenopausal women
file <- "../results/bmodels/version4/logit_post_b"

logistic_post_b <- brm(iron_deficiency ~
                       Age + 
                        #BMI + 
                        #Smoking + 
                        #iron_complience + 
                        #donation_count_2 + 
                        #donation_count +
                        #log_last_donation + 
                        #log_CRP +
                        #Region + 
                        Weight +
                        #RedMeat_n +
                        #PreviousChildbirth +
                        #Menstruation +   # will not work as all observations are no_period
                        blood_vol + 
                        Cohort, 
                     data = regression_data_post, 
                     family = bernoulli(),
                     file=file, 
                     cores = 4,
                     iter= 10000
)

summary(logistic_post_b)
```

## Diagnostics

### Trace

```{r , message=FALSE}
mcmc_plot(logistic_men_b, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(logistic_pre_b, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(logistic_post_b, type = "trace")
```

Chains do not seem to get stuck.

### ACF bars

```{r}
mcmc_plot(logistic_men_b, type = "acf_bar")
```


```{r}
mcmc_plot(logistic_pre_b, type = "acf_bar")
```


```{r}
mcmc_plot(logistic_post_b, type = "acf_bar")
```

Autocorrelation drops nicely.

### Rhat

```{r}
mcmc_plot(logistic_men_b, type = "rhat")
```

```{r}
mcmc_plot(logistic_pre_b, type = "rhat")
```

```{r}
mcmc_plot(logistic_post_b, type = "rhat")
```
https://mc-stan.org/rstan/reference/Rhat.html


"The Rhat function produces R-hat convergence diagnostic, which compares the between- and within-chain estimates for model parameters and other univariate quantities of interest. If chains have not mixed well (ie, the between- and within-chain estimates don't agree), R-hat is larger than 1. We recommend running at least four chains by default and only using the sample if R-hat is less than 1.05."

## Forest plots

```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(logistic_men_b, 
         type = "areas",
         prob = 0.95,
         transformations = "exp") +
  geom_vline(xintercept = 1, color = "grey")

p <- p + scale_x_log10()
#Ferritin needs to be removed
# plot.OR<-  tidy_logit_pre_menop_women %>%filter(!term=="log_Ferritin_beginning") %>%  dplyr::select(OR) %>% mutate(y=rev(2:(nrow(tidy_logit_pre_menop_women)))) %>% bind_rows(tibble(OR=NA,y=1 ))
# p <- p + geom_point(aes(y=y,x=OR),data=plot.OR, col="red")
p
```


```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(logistic_pre_b, 
         type = "areas",
         prob=0.95,
         prob_est="median",
         transformations = "exp"
        ) +
  geom_vline(xintercept = 1, color = "grey")

p <- p + scale_x_log10()
p
```


```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(logistic_post_b, 
         type = "areas",
         prob = 0.95,
         transformations = "exp") +
  geom_vline(xintercept = 1, color = "grey")

p <- p + scale_x_log10()
#Ferritin needs to be removed
# plot.OR<-  tidy_logit_pre_menop_women %>%filter(!term=="log_Ferritin_beginning") %>%  dplyr::select(OR) %>% mutate(y=rev(2:(nrow(tidy_logit_pre_menop_women)))) %>% bind_rows(tibble(OR=NA,y=1 ))
# p <- p + geom_point(aes(y=y,x=OR),data=plot.OR, col="red")
p
```

### All groups together

```{r}
regressor_values <- c(
 "Age" = "Age (5 years)",
   "log_CRP" = "CRP (log transformed)",
   "BMI" = "BMI",
   "log_last_donation" = "Days since last donation (log, /log(2))",
   "Smokingyes" ="Smoker",
    "iron_complience" = "Iron supplementation",
    "donation_count_2" = "Donations in last 2 years (quadratic)",
    "Menstruationregular_period" = "Regular periods",
    "Weight" = "Weight (10 kg)",
    "PreviousChildbirth" = "Given birth in the past",
    "donation_count" = "Donations in last 2 years",
    "CohortHEALTH00" = "Cohort (Health 2000)",
    "CohortFINRISK97" = "Cohort (FinRisk 1997)",
    "blood_vol" = "Blood volume (Nadler's equation)"
)
```




```{r }

intervals_men <- mcmc_intervals_data(
  logistic_men_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median",
  transformations = "exp"
  ) %>% mutate(Group='Men')

intervals_pre <- mcmc_intervals_data(
  logistic_pre_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median",
  transformations = "exp"
  ) %>% mutate(Group='Pre-menopausal women') %>% 
  arrange(m) # how to order the variables?

intervals_post <- mcmc_intervals_data(
  logistic_post_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median",
  transformations = "exp"
  ) %>% mutate(Group='Post-menopausal women')

intervals_combined <- rbind(intervals_pre,intervals_post,intervals_men) %>% 
  filter(! (parameter == 'exp(b_Intercept)' | parameter == 'exp(lp__)' | parameter == 'exp(lprior)')) %>% 
  mutate(parameter = gsub("\\)","",gsub("exp\\(b_","",parameter))) %>% 
  mutate(regressor = plyr::revalue(parameter, regressor_values)) %>% 
  select(-parameter)
  #arrange(m)


```


```{r , fig1, fig.height = 10, fig.width = 15 ,warning=FALSE}

orlower <- 1/10
orupper <- 10

# Truncate ORs for plotting
plotdata <- intervals_combined %>% 
  mutate(
    ll = case_when(
      ll < orlower ~ orlower,
      TRUE ~ ll
    ),
    hh = case_when(
      hh > orupper ~ orupper,
      TRUE ~ hh
    )
  )

#theme_set(bayesplot::theme_default())
pos <- position_nudge(y=as.numeric(as.factor(plotdata$Group))/5 -0.1) 

p <- ggplot(plotdata, aes(x = m, y = regressor, color = Group)) + 
  geom_point(position = pos,
             size=4) +
  geom_linerange(aes(xmin = ll, xmax = hh), 
                 position = pos,
                 size = 1, 
                 linetype = 1
                 ) +
  scale_color_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                     limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
  scale_fill_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                    limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
  scale_shape_manual(values = c(1,16)) +
  scale_x_log10(limits=c(orlower,orupper))  + #how to get the lines that drop out of the visible?  
  geom_vline(xintercept = 1 , color = "grey", linetype = "dotted",
             size = 1) +
  theme(axis.title = element_text(size = 14),
        strip.text =  element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.background = element_rect(fill = "white"),
        strip.background.x = element_blank(),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14)
  ) +
  guides(shape = "none") +
  theme_classic() +
  ylab("Coefficient") 

p


```

```{r}
ggplot2::ggsave(filename = "../results/figures/low_ferritin_model/version4_forest_all_cohorts_bayes.pdf",
                plot = p,
                width = 35,
                height = 25,
                dpi = 600,
                units = "cm")

ggplot2::ggsave(filename = "../results/figures/low_ferritin_model/version4_forest_all_cohorts_bayes.png",
                plot = p,
                width = 35,
                height = 25,
                dpi = 300,
                units = "cm")


# save ORs for comparison with other models 
OR_logistic_bayes <- plotdata
save(OR_logistic_bayes, file = "../data/OR_low_ferritin_model_bayes_version4.rdata")
```

```{r , fig2, fig.height = 6, fig.width = 13 ,warning=FALSE}
# smaller picture for publication/presentation

p_pub <- p + theme(axis.title = element_text(size = 14),
      strip.text =  element_text(size = 14),
      legend.text = element_text(size = 12),
      strip.background = element_rect(fill = "white"),
      strip.background.x = element_blank(),
      axis.text.y = element_text(size = 14),
      axis.text.x = element_text(size = 14, angle = 30,hjust = 0.8),
      plot.title =  element_text(size = 16),
      plot.subtitle = element_text(size = 16,
                                   hjust = 0.5))

p_pub
```

```{r}
ggplot2::ggsave(filename = "../results/figures/low_ferritin_model/pub_version4_forest_all_cohorts_bayes.pdf",
                plot = p_pub,
                width = 35,
                height = 25,
                dpi = 600,
                units = "cm")

ggplot2::ggsave(filename = "../results/figures/low_ferritin_model/pub_version4_forest_all_cohorts_bayes.png",
                plot = p_pub,
                width = 35,
                height = 25,
                dpi = 300,
                units = "cm")
```

<!-- ## Compare to bootstrapped -->



<!-- ```{r  fig2, fig.height = 15, fig.width = 10 ,warning=FALSE} -->
<!-- file <- "../results/bootstraps/plotdata_filtered_version1.rdata" #Version 1 of the bootstrap model corresponds to version 3 of Bayes -->
<!-- if(file.exists(file)) { -->
<!--   load(file) -->
<!--   temp <- plotdata_filtered %>% select(-p.value,-is_sig) %>%  -->
<!--     rename(ll = Bca_inf, -->
<!--            hh = Bca_sup, -->
<!--            m = OR -->
<!--     ) %>%  mutate( -->
<!--       Method = "Bayes" -->
<!--     ) %>%  rename(Group=group) -->

<!--   temp <- plotdata %>%  select(regressor,ll,hh,m,Group) %>%  -->
<!--     mutate(Method="Boot") %>%  -->
<!--   bind_rows(temp) -->

<!--   temp <- temp %>%  -->
<!--     mutate( -->
<!--       ll = case_when( -->
<!--         ll < orlower ~ orlower, -->
<!--         TRUE ~ ll -->
<!--       ), -->
<!--       hh = case_when( -->
<!--         hh > orupper ~ orupper, -->
<!--         TRUE ~ hh -->
<!--       ) -->
<!--     ) -->

<!--   pos <- position_nudge(y=as.numeric(as.factor(temp$Method))/5 -0.1)  -->

<!--   p <- ggplot(temp, aes(x = m, y = regressor, color = Method)) +  -->
<!--      geom_point(position = pos, -->
<!--                 size=4) + -->
<!--     geom_linerange(aes(xmin = ll, xmax = hh),  -->
<!--                    position = pos, -->
<!--                    size = 1  -->
<!--     ) + -->
<!--     # scale_color_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ), -->
<!--     #                    limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) + -->
<!--     # scale_fill_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ), -->
<!--     #                   limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) + -->
<!--     #scale_shape_manual(values = c(1,16)) + -->
<!--     scale_x_log10(limits=c(orlower,orupper))  + #how to get the lines that drop out of the visible?   -->
<!--     geom_vline(xintercept = 1 , color = "grey", linetype = "dotted", -->
<!--                size = 1) + -->
<!--     theme(axis.title = element_text(size = 14), -->
<!--           strip.text =  element_text(size = 14), -->
<!--           legend.text = element_text(size = 12), -->
<!--           strip.background = element_rect(fill = "white"), -->
<!--           strip.background.x = element_blank(), -->
<!--           axis.text.x = element_text(size = 14), -->
<!--           axis.text.y = element_text(size = 14) -->
<!--     ) + -->
<!--     guides(shape = "none") + -->
<!--     theme_classic() + -->
<!--     facet_grid(Group~.) + -->
<!--     ylab("Coefficient")  -->

<!--   p -->


<!-- } -->
<!-- ``` -->


<!-- ```{r} -->
<!-- if(file.exists(file)) { -->

<!-- ggplot2::ggsave(filename = "../results/figures/forest_all_cohorts_bayes3_vs_boot1.pdf", -->
<!--                 plot = p, -->
<!--                 width = 35, -->
<!--                 height = 30, -->
<!--                 dpi = 600, -->
<!--                 units = "cm") -->
<!-- } -->

<!-- ``` -->




<!-- #Separate plots for individual variables -->


<!-- ```{r} -->
<!-- file <- "../results/summary_all_cohorts.rdata" #Note: this is temporary data and will be changed for every version, don't load it for previous versions -->
<!-- load(file) -->

<!-- summary_all_cohorts <- summary_all_cohorts %>%  -->
<!--   rename(donation_count = TwoYearsFromStartCount_FB, -->
<!--          last_donation = DaysToPreviousFB, -->
<!--          iron_complience = IronComplienceNumeric) %>%  -->
<!--     filter(!is.na(CRP)) %>%  # Ploting the same persons that we used in the model -->
<!--     filter(!CRP >= 30) -->

<!-- summary(summary_all_cohorts) -->
<!-- ``` -->

<!-- ## Ferritin and menstruation -->

<!-- ```{r , message=FALSE} -->

<!-- temp <- summary_all_cohorts %>%  -->
<!--   filter(Group == 'Pre_menopause_women') %>% -->
<!--   mutate( iron_deficiency = Ferritin < 15) %>%  -->
<!--   group_by(iron_deficiency,Menstruation,Cohort) %>%  -->
<!--   summarise(n=n() -->
<!--             ) %>%  -->
<!--   group_by(Cohort) %>%  -->
<!--   summarise( -->
<!--     iron_deficiency = iron_deficiency, -->
<!--     Menstruation = Menstruation, -->
<!--     Freq = paste(round(n / sum(n),2) *100, "%"), -->
<!--     N = n -->
<!--             ) %>%  -->
<!--   mutate( -->
<!--     x = c(150,5)[as.numeric(as.factor(iron_deficiency))], -->
<!--     y =  c(20,100)[as.numeric(as.factor(Menstruation))] -->
<!--   ) -->

<!-- p <- ggplot(summary_all_cohorts %>% filter(Group == 'Pre_menopause_women'), -->
<!--             aes(x=Ferritin)) +  -->
<!--   geom_histogram(position = "dodge") + -->
<!--   scale_x_log10() +  -->
<!--   geom_text(aes(x=x,y=y,label=Freq),data=temp) +  -->
<!--   facet_grid(Menstruation~Cohort,scales = "free_y") +  -->
<!--   geom_vline(aes(xintercept=15)) + -->
<!--   ylab("Count of persons") -->
<!-- p -->
<!-- ggsave(filename = "../results/figures/hist_ferritin_by_menstruation.png", -->
<!--         plot = p,  -->
<!--       width = 10, -->
<!--       height = 15, -->
<!--       dpi = 300, -->
<!--       units = "cm") -->


<!-- ``` -->

<!-- ## Ferritin and CRP -->

<!-- ```{r} -->

<!-- p <- ggplot(aes(y=Ferritin, x=CRP),data=summary_all_cohorts) -->
<!-- p <- p+ geom_point() -->
<!-- p <- p +  geom_hline(aes(yintercept=15)) -->
<!-- p <- p + facet_grid(Group~Cohort) -->
<!-- p <- p + scale_x_log10() + scale_y_log10() -->
<!-- p <- p + geom_smooth(method='gam') -->
<!-- p -->

<!-- ``` -->

<!-- ## Ferritin and CRP, include linear model -->

<!-- ```{r} -->
<!-- p <- ggplot(aes(y=Ferritin, x=CRP),data=summary_all_cohorts) -->
<!-- p <- p+ geom_point() -->
<!-- p <- p +  geom_hline(aes(yintercept=15)) -->
<!-- p <- p + facet_grid(Group~Cohort) -->
<!-- p <- p + scale_x_log10() + scale_y_log10() -->
<!-- p <- p + geom_smooth(method='gam') -->
<!-- p <- p + geom_smooth(method="lm", color="red") -->
<!-- p -->
<!-- ``` -->

