---
title: "Low ferritin model"
author: "Sofie Ekroos"
date: "19.7.2021"
output: github_document
bibliography: index.bib
---

Based on the original code written by Muriel Lobier and Mikko Arvas (https://github.com/FRCBS/iron_levels_of_blood_donors/blob/master/src/index.Rmd and https://github.com/FRCBS/changes_in_donor_health/blob/master/src/explorative_analysis.Rmd)

# Summary

This document includes all codes necessary to run the analysis of and produce the figures for the three Finnish Cohorts (FinDonor, FINRISK97, Health2000). The code allows the user to:

- describe the cohorts and compute the prevalence of iron deficiency
- identify co-variants of iron deficiency using multivariable linear regression 

```{r setup, , message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# library(janitor)
library(tableone)
library(stargazer)
library(gridExtra)
library(cowplot)
library(broom)
library(GGally)
library(ggfortify)
# library(sjPlot)
library(knitr)
library(lubridate)
#library(HDInterval) !! package ‘HDIinterval’ is not available (for R version 3.6.3)!!!
library(car)
#library(relaimpo)
library(sfsmisc)
library(MASS)
# library(kableExtra)
#library(brms)
#library(rstan)
#library(rms)
#library(modelr)
#library(tidybayes)
library(ordinal)
library(sjmisc)
#library(sjlabelled)
# library(lazerhawk)
library(epiDisplay)
library(safeBinaryRegression)
library(epitools)
library(ez)
library(ggbeeswarm)
library(tidymodels)
```

# Data loading and preparation

## Data loading

We have one cohort of blood donors, FinDonor, and two general population cohorts, FinRisk97 and Health2000. 

```{r , warning=FALSE}

# load FinDonor biomarker data
load("../data/r02.fd.bd.all.rdata")
indiv_donations_data <- output
rm(output)

# load FinDonor demographic data
load("../data/r02ds.donorData.rdata")
donor_demo <- output
rm(output)

# load THL data
load("../data/thldata.rdata")

# thldalta.rdata contains all five THL cohorts, extract FINRISK97 and Health2000 from the others 
fr1997 <- thldata$fr1997
h2000 <- thldata$h2000
rm(thldata)
```

## Data preparation

First we assign study participants to their cohorts.

```{r}
indiv_donations_data$Cohort <- c("FINDONOR")
fr1997$Cohort <- c("FINRISK97")
h2000$Cohort <- c("HEALTH00")
```

There are `r cohort_name %>% distinct(ID_variable) %>% nrow()` individials enrolled in the studies. 

  * FinDonor: 2584   `r indiv_donations_data %>% distinct(donor) %>% nrow()`
  * FinRisk97: 7943  `r fr1997 %>% distinct(RELEASE_ID) %>% nrow()`
  * Health2000: 6264 `r h2000 %>% distinct(RELEASE_ID) %>% nrow()` 
        
Once we remove participants that have no ferritin or no Hb measurement data (and for blood donors for any donation event) there are `r cohort_name %>%  filter(!is.na(Ferritin) & !is.na(Hb_v) ) %>% distinct(ID_variable) %>% nrow()` participants left  

  * FinDonor: 2580 `r indiv_donations_data %>%  filter(!is.na(Ferritin) & !is.na(Hb_v) ) %>% distinct(donor) %>% nrow()`
  
  * FinRisk97: 462 `r fr1997 %>%  filter(!is.na(FERRITIN) & !is.na(HGB) ) %>% distinct(RELEASE_ID) %>% nrow()`
           
            * 7380 `r fr1997 %>% filter(!is.na(FERRITIN)) %>% distinct(RELEASE_ID) %>% nrow()`
            * 491 `r fr1997 %>%  filter(!is.na(HGB)) %>% distinct(RELEASE_ID) %>% nrow()`
              
  * Health2000: 5250 `r h2000 %>%  filter(!is.na(FERRITIINI) & !is.na(B_Hb) ) %>% distinct(RELEASE_ID) %>% nrow()`

At this stage we noticed we only had hemoglobin data from 491 FinRisk97 participants. Using the table function (`r table(fr1997$HGB, fr1997$SUKUP)`, SUKUP = sukupuoli = sex, 1=male, 2=female) we found out that all these participants are male. As such, we didn't remove NA:s for Hb and will not use Hb in the model for this specific cohort.  

```{r , warning=FALSE}

#FINDONOR

## Get sex for each blood donor
blood_data_summary <- indiv_donations_data %>% 
  group_by(donor) %>% 
  summarise(Sex=first(gender)) 

## Get values for first study donation with the required measurements donation (regardless of donation type)
## We remove events with no Ferritin and hb_v to be inline with from when the nb of previous donations were counted.
blood_values_init <- indiv_donations_data %>% 
  filter(!is.na(Ferritin) & !is.na(Hb_v) ) %>% 
  group_by(donor) %>% 
  filter(date == min(date)) %>% 
  dplyr::select(age, Ferritin, Hb_v, CRP, DaysToPreviousFB, donor, Cohort) %>% 
  ungroup()

blood_data_summary <- blood_data_summary %>% 
    full_join(blood_values_init, by = "donor") 

blood_data_summary <- donor_demo %>%
  inner_join(blood_data_summary, by = "donor") %>%
  mutate(all_study_FB_donation_count = NonFinnDonorDonationCount_FB + YesFinnDonorDonationCount_FB) 

#FINRISK97 

## Change sex from binary 1,2 to men, women
fr1997$SUKUP <- gsub("1", "Men", fr1997$SUKUP)
fr1997$SUKUP <- gsub("2", "Women", fr1997$SUKUP)

fr1997_summary <- fr1997 %>%
  filter(!is.na(FERRITIN)) %>%
  group_by(RELEASE_ID) %>%
  dplyr::select(IKA, SUKUP, FERRITIN, HGB, CRP, RELEASE_ID, K129, BMI, KY100_22, Q69, KY163, GRAVID, Cohort) %>%
  ungroup()

#HEALTH2000
## Change sex from binary 1,2 to men, women
h2000$SP2 <- gsub("1", "Men", h2000$SP2)
h2000$SP2 <- gsub("2", "Women", h2000$SP2)  
  
 h2000_summary <- h2000 %>%
    filter(!is.na(FERRITIINI) & !is.na(B_Hb)) %>%
    group_by(RELEASE_ID) %>%
    dplyr::select(IKA2, SP2, FERRITIINI, B_Hb, CRP, RELEASE_ID, BD03, BMII_BMI.x, FB05, synnytys, BD07, Cohort, MENOP, BD06) %>%
    ungroup()  
  
```

Next we rename the variables. 

```{r}
## Participant ID 
names(blood_data_summary)[names(blood_data_summary) == "donor"] <- "ID"
names(fr1997_summary)[names(fr1997_summary) == "RELEASE_ID"] <- "ID"
names(h2000_summary)[names(h2000_summary) == "RELEASE_ID"] <- "ID"

## Sex
names(fr1997_summary)[names(fr1997_summary) == "SUKUP"] <- "Sex"
names(h2000_summary)[names(h2000_summary) == "SP2"] <- "Sex"

## Age
names(fr1997_summary)[names(fr1997_summary) == "IKA"] <- "Age"
names(h2000_summary)[names(h2000_summary) == "IKA2"] <- "Age"

## Hemoglobin
names(blood_data_summary)[names(blood_data_summary) == "Hb_v"] <- "Hb"
names(fr1997_summary)[names(fr1997_summary) == "HGB"] <- "Hb"
names(h2000_summary)[names(h2000_summary) == "B_Hb"] <- "Hb"

## Ferritin
names(fr1997_summary)[names(fr1997_summary) == "FERRITIN"] <- "Ferritin"
names(h2000_summary)[names(h2000_summary) == "FERRITIINI"] <- "Ferritin"

## Menstruation
names(blood_data_summary)[names(blood_data_summary) == "QR79"] <- "Menstruation"
names(fr1997_summary)[names(fr1997_summary) == "K129"] <- "Menstruation"
names(h2000_summary)[names(h2000_summary) == "BD03"] <- "Menstruation"

## BMI
names(h2000_summary)[names(h2000_summary) == "BMII_BMI.x"] <- "BMI"

## Smoking
names(blood_data_summary)[names(blood_data_summary) == "QR54"] <- "Smoking"
names(fr1997_summary)[names(fr1997_summary) == "Q69"] <- "Smoking"
names(h2000_summary)[names(h2000_summary) == "FB05"] <- "Smoking"

## Red meat
names(blood_data_summary)[names(blood_data_summary) == "QR40"] <- "RedMeat"
names(fr1997_summary)[names(fr1997_summary) == "KY100_22"] <- "RedMeat"

## Iron supplements
names(blood_data_summary)[names(blood_data_summary) == "vita_iron"] <- "HistoryOfIronSupplements"
names(blood_data_summary)[names(blood_data_summary) == "iron_supp"] <- "GivenIronSupplements"
names(blood_data_summary)[names(blood_data_summary) == "iron_comp"] <- "IronComplience"
names(blood_data_summary)[names(blood_data_summary) == "iron_comp_c"] <- "IronComplienceNumeric"

## History of childbirth
names(blood_data_summary)[names(blood_data_summary) == "QR83"] <- "PreviousChildbirth"
names(fr1997_summary)[names(fr1997_summary) == "KY163"] <- "PreviousChildbirth"
names(h2000_summary)[names(h2000_summary) == "synnytys"] <- "PreviousChildbirth"

## Current pregnancy
names(fr1997_summary)[names(fr1997_summary) == "GRAVID"] <- "CurrentPregnancy"
names(h2000_summary)[names(h2000_summary) == "BD07"] <- "CurrentPregnancy"

## THESE ARE ONLY NEEDED FOR h2000 to figure out the n/a:s related to menstruation
names(h2000_summary)[names(h2000_summary) == "BD06"] <- "WhyEnd"
#hysterectomy?

```

## Demographic group specification and assignment (coninued after "figuring out why so many Health2000 participants are removed" section)

In this section we assign women to period/no period groups. The menstruation question was similar across all three cohorts:

  * FinDonor: Do you still have regular periods? (1=reg., 2=irreg., 3=no)
  * FinRisk97: Do you still menstruate? (1=reg., 2=irreg., 3=no)
  * Health2000: BD03 = do you have periods nowadays? (1=reg., 2=irreg., 3=no)

There are `r cohort_summary_name %>% filter(Sex == "Women" & is.na(Menstruation)) %>%  nrow()` women with no  answer to the question regarding their menstrual status. 

  * FinDonor: 17 `r blood_data_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>%  nrow()`
  * FinRisk97: 41 `r fr1997_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>%  nrow()`
  * Health00: 1029 `r h2000_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>%  nrow()`
  
Evidently we are missing menstruation data from >37 % of female Health2000 participants.

### Figuring out why so many Health2000 participants are removed (skip this part unless there are problems with the cohort)

Menstruation was defined as: 
    1 = regular menstruation
    2 = irregular menstruation
    3 = no menstruation
The cohort also contains a variable for reproductive stages, MENOP:
    1 = postmenopause 
    2 = perimenopause
    3 = premenopause
    
Removing n/a:s for the MENOP variable `r h2000_summary %>% filter(Sex == "Women" & is.na(MENOP)) %>%  nrow()` would remove only 48 participants. In the next code chunk we investigate the difference between these variables in order to figure out which we can/should use.

In the home interview instruction (https://thl.fi/documents/189940/4108213/T2001_eng.pdf/cd17a5fe-ddf3-4649-9ddd-a282b1809de9, page 41) we find out that only women under the age of 55 yrs were asked about menstruation. This amounts to:
  * 180 NAs in the <55 yrs group `r h2000_summary %>% filter(Sex == "Women" & Age < 55 & is.na(Menstruation)) %>%  nrow()`
  * 849 NAs in the >=55 yrs group `r h2000_summary %>% filter(Sex == "Women" & Age >= 55 & is.na(Menstruation)) %>%  nrow()`

Our original plan was to place participants in pre- and postmenopausal groups based on menstrual status. Removing n/a:s based on the MENOP variable would remove fewer participants, however we are missing this data in the other cohorts. Next we try to find out how menstruation, reproductive stage and age align in the Health2000 cohort: 

```{r}
h2000_mens_missing <- h2000_summary 

# make new variables based on age (<45 yrs, >=45 yrs) and reproductive stage
h2000_mens_missing <- h2000_mens_missing %>% 
  mutate(AgeGroup = case_when(
    Sex == "Women" & (Age >= 45) ~ "older",
    Sex == "Women" & (Age < 45) ~ "younger",
    )) %>% 
  mutate(Stage = case_when(
    MENOP == "1" ~ "POSTmenopausal",
    MENOP == "2" ~ "PERImenopausal",
    MENOP == "3" ~ "PREmenopausal"
  )) %>% 
  mutate(MENSTRUATION = case_when(
    Menstruation == "1" ~ "regular",
    Menstruation == "2" ~ "irregular",
    Menstruation == "3" ~ "no"
  )) %>% 
  filter(Sex == "Women") 


## first we check  what data on reproductive state looks like 
h2000_mens_missing %>% 
  filter(is.na(Stage) ) %>% 
  nrow() 
# 48 participants missing data on reproductive stage 

h2000_missing_stage <- h2000_mens_missing %>% 
  filter(!is.na(Stage)) 

# check to see if recode worked
table(factor(h2000_missing_stage$Stage, levels = c("PREmenopausal", "PERImenopausal", "POSTmenopausal")), h2000_missing_stage$MENOP)

 #                   1    2    3
 # PREmenopausal     0    0 1516
 # PERImenopausal    0  103    0
 # POSTmenopausal 1092    0    0       -> yes, they were assigned to the right groups


# compare reproductive stage with reported bleeding in the two age groups (<45, >45)
table(factor(h2000_mens_missing$MENSTRUATION, levels = c("regular", "irregular", "no")), factor(h2000_mens_missing$Stage, levels = c("PREmenopausal", "PERImenopausal", "POSTmenopausal")), factor(h2000_mens_missing$AgeGroup, levels = c("younger", "older")))

                                  # YOUNGER (<45 yrs):
                                  #            
                                  #             PREmenopausal PERImenopausal POSTmenopausal
                                  #   regular             858              0              0
                                  #   irregular           119              0              0
                                  #   no                   11              6             57
                                  # 
                                  # OLDER (>= 45 yrs):
                                  #            
                                  #             PREmenopausal PERImenopausal POSTmenopausal
                                  #   regular             316              0              0
                                  #   irregular           135              0              0
                                  #   no                   16             38            160
   

# compare reported bleeding in the two age groups 
table(factor(h2000_mens_missing$MENSTRUATION, levels = c("regular", "irregular", "no")), factor(h2000_mens_missing$AgeGroup, levels = c("younger", "older")))


                                                #           younger older
                                                # regular       858   316
                                                # irregular     121   141
                                                # no             78   216

# compare reported reproductive stage in the two age groups 
table(factor(h2000_mens_missing$Stage, levels = c("PREmenopausal", "PERImenopausal", "POSTmenopausal")), factor(h2000_mens_missing$AgeGroup, levels = c("younger", "older")))


                                                #                younger older
                                                # PREmenopausal      995   521
                                                # PERImenopausal      11    92
                                                # POSTmenopausal      78  1014


## next we investigate how removing n/a:s only from women under 55 yrs compares to the above data (women older that 55 are concidered postmenopausal)

women_older_than_55 <- h2000_mens_missing %>% 
  filter(Age >= 55) %>% nrow() # 849 women 

h2000_mens_missing %>% 
  filter(Age < 55 & is.na(MENSTRUATION)) %>% 
  nrow() 
# we are missing data from 180 women < 55 years 

h2000_mens_missing %>% 
  filter(Age < 45 & is.na(MENSTRUATION)) %>% 
  nrow() 
# 43 women with missing data are < 45 years 


women_older_than_55 <- h2000_mens_missing %>% 
  filter(Age >= 55) %>% nrow() # 849 women 

women_older_than_55 <- filter(h2000_mens_missing,
                              Age >= 55 & is.na(Menstruation))$ID

test <- h2000_mens_missing %>% 
  mutate(Menstruation = ifelse(ID %in% women_older_than_55, "no", as.character(Menstruation))) 


```

For now we will remove  n/a:s from women under 55 and use the same strategy for assigning reproductive groups as we do for the other cohorts. Once we recieve an explanation for the reasoning behind the MENOP-varible we may reconcider. 


## Demographic group specification and assignment continued

Out of the women with no answer to the menstruation question, `r cohort_summary_name %>% filter(Sex == "Women" & is.na(Menstruation) & age >= 55) %>%  nrow()` are older than 55. We impute these to no period so they can be included in the postmenopausal women. 

  * FinDonor: 3 `r blood_data_summary %>% filter(Sex == "Women" & is.na(Menstruation) & Age >= 55) %>%  nrow()`
  * FinRisk97: 32 `r fr1997_summary %>% filter(Sex == "Women" & is.na(Menstruation) & Age >= 55) %>%  nrow()`
  * Health2000: 849 `r h2000_summary %>% filter(Sex == "Women" & is.na(Menstruation) & Age >= 55) %>%  nrow()`

```{r}
# FINDONOR
older_findonor_donors <- filter(blood_data_summary, 
                       Sex == "Women" & is.na(Menstruation) & Age >= 55)$ID

blood_data_summary <- blood_data_summary %>% 
  mutate(Menstruation = ifelse(ID %in% older_findonor_donors, "no_period", as.character(Menstruation))) 

# FINRISK97
older_fr1997 <- filter(fr1997_summary, 
                       Sex == "Women" & is.na(Menstruation) & Age >= 55)$ID

fr1997_summary <- fr1997_summary %>% 
  mutate(Menstruation = ifelse(ID %in% older_fr1997, "no_period", as.character(Menstruation)))

# HEALTH2000
older_h2000 <- filter(h2000_summary,
                      Sex == "Women" & is.na(Menstruation) & Age >= 55)$ID

h2000_summary <- h2000_summary %>%
  mutate(Menstruation = ifelse(ID %in% older_h2000, "no_period", as.character(Menstruation)))

# h2000_mens_missing <- h2000_summary 
# 
# # renaming the menstruation variable
# h2000_mens_missing <- h2000_mens_missing %>% 
#   mutate(Menstruation = case_when(
#     Menstruation == "1" ~ "regular",
#     Menstruation == "2" ~ "irregular",
#     Menstruation == "3" ~ "no"
#   )) %>% 
#   filter(Sex == "Women") 
# 
# women_older_than_55 <- filter(h2000_mens_missing,
#                               Age >= 55 & is.na(Menstruation))$ID
# 
# women_older_than <- h2000_mens_missing %>% 
#   mutate(MENSTRUATION = ifelse(ID %in% women_older_than_55, "no", as.character(MENSTRUATION))) 
# 
# older_h2000 <- filter(h2000_summary,
#                       Sex == "Women" & is.na(MENOP) & Age > 55)$ID
# h2000_summary <- h2000_summary %>% 
#   mutate(Menstruation = ifelse(ID %in% older_h2000, "no_period", as.character(MENOP)))

```


Female study participants with no menstruation response after imputation are removed:  `r cohort_name %>% filter(sex == "Women" & is.na(Menstruation)) %>% nrow()` donors.)

  * FinDonor: 14 `r blood_data_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>%  nrow()`
  * FinRisk97: 9 `r fr1997_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>%  nrow()`
  * Health2000: 180 `r h2000_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>%  nrow()`

```{r}
#FINDONOR
findonor_nb_women_no_menstruation_response <- blood_data_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>% nrow()

## remove female donors with no menstruation response
blood_data_summary_final <- blood_data_summary %>% 
  mutate(mens_ok_blood = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & !is.na(Menstruation)  ~ "Women",
    TRUE ~ "NA")) %>% 
  filter(mens_ok_blood != "NA") 

## We define the women's groups: 
findonor_n_women_removed <- blood_data_summary_final %>% 
  mutate(Group = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & (Menstruation == "no_period" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "no_period" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "irregular_period" | Menstruation == "regular_period") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>% 
  group_by(Group) %>% 
  filter(Group ==  "Women_pre_menop_no_mens") %>% 
  nrow()


#FINRISK97

finrisk_nb_women_no_menstruation_response <- fr1997_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>% nrow()


## remove female FinRisk97 participants with no menstruation response
fr1997_summary_final <- fr1997_summary %>% 
  mutate(mens_ok_fr1997 = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & !is.na(Menstruation)  ~ "Women",
    TRUE ~ "NA")) %>% 
  filter(mens_ok_fr1997 != "NA") 

## We define the women's groups:
fr1997_n_women_removed <- fr1997_summary_final %>%
  mutate(Group = case_when(
    Sex == "Men" ~ "Men",
    Sex == "Women" & (Menstruation == "3" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "3" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "1" | Menstruation == "2") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>%
  group_by(Group) %>%
  filter(Group ==  "Women_pre_menop_no_mens") %>%
  nrow()

#HEALTH2000

h2000_summary <- h2000_summary %>%
  mutate(Menstruation = ifelse(ID %in% older_h2000, "no_period", as.character(Menstruation)))



h2000_nb_women_no_menstruation_response <- h2000_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>% nrow()


## remove female Health2000 participants with no menstruation response
h2000_summary_final <- h2000_summary %>%
  mutate(mens_ok_h2000 = case_when(
    Sex == "Men" ~ "Men",
    Sex == "Women" & !is.na(Menstruation)  ~ "Women",
    TRUE ~ "NA")) %>%
  filter(mens_ok_h2000 != "NA")

## We define the women's groups:
h2000_n_women_removed <- h2000_summary_final %>%
  mutate(Group = case_when(
    Sex == "Men" ~ "Men",
    Sex == "Women" & (Menstruation == "3" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "3" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "1" | Menstruation == "2") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>%
  group_by(Group) %>%
  filter(Group ==  "Women_pre_menop_no_mens") %>%
  nrow()
# 
# 
# h2000_nb_women_no_menstruation_response <- h2000_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>% nrow()
# h2000_nb_women_no_menstruation_response <- h2000_summary %>% filter(Sex == "Women" & is.na(MENOP)) %>% nrow()
# h2000_summary %>% filter(MENOP == "2") %>% nrow() # 103 perimenopausal women 
# 
# 
# ## remove female Health2000 participants with no MENOP response and perimenopausal women
# h2000_summary_final <- h2000_summary %>% 
#   mutate(mens_ok_h2000 = case_when(
#     Sex == "Men" ~ "Men", 
#     Sex == "Women" & !is.na(MENOP)  ~ "Women",
#     TRUE ~ "NA")) %>% 
#   filter(mens_ok_h2000 != "NA") 
#   #%>% filter("Women" & MENOP != "2")
# 
# ## We define the women's groups:
# h2000_n_women_removed <- h2000_summary_final %>%
#   mutate(Group = case_when(
#     Sex == "Men" ~ "Men",
#     Sex == "Women" & (MENOP == "1") ~ "Post_menopause_women",
#     Sex == "Women" & (MENOP == "3") ~ "Pre_menopause_women",
#     TRUE ~ "NA")) %>%
#   group_by(Group) %>%
#   filter(Group ==  "Women_pre_menop_no_mens") %>%
#   nrow()

```

We now define the following groups:

+ premenopausal: regular or irregular menstruation reported
+ postmenopausal: no menstruation reported and age equal to > 45
+ donors younger than 45 and no reported menstruation are excluded (`r cohort_n_women_removed`)
  * 77 FinFonor `r findonor_n_women_removed`
  * 81 FinRisk `r fr1997_n_women_removed`
  * 83 FinFonor `r h2000_n_women_removed`

```{r}
#FINDONOR:
## We define the women's groups and drop n/a:s: 
blood_data_summary_final <- blood_data_summary_final %>% 
  mutate(Group = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & (Menstruation == "no_period" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "no_period" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "irregular_period" | Menstruation == "regular_period") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>% 
  filter(Group != "Women_pre_menop_no_mens" & Group != "NA") %>% 
  droplevels() %>% 
  mutate(Group = ordered(Group, levels =  c("Pre_menopause_women", "Post_menopause_women", "Men"))) 


#FINRISK97:
## We define the women's groups and drop n/a:s: 
fr1997_summary_final <- fr1997_summary_final %>% 
  mutate(Group = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & (Menstruation == "3" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "3" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "1" | Menstruation == "2") ~ "Pre_menopause_women",
    Sex == "Women" & (Menstruation == "no_period") ~ "Post_menopause_women",
    TRUE ~ "NA")) %>% 
  filter(Group != "Women_pre_menop_no_mens" & Group != "NA") %>% 
  droplevels() %>% 
  mutate(Group = ordered(Group, levels =  c("Pre_menopause_women", "Post_menopause_women", "Men"))) 
 
#HEALTH2000:
# We define the women's groups and drop n/a:s:

# h2000_summary_final_test1 <- h2000_summary_final %>%
#   mutate(Group = case_when(
#     Sex == "Men" ~ "Men",
#     Sex == "Women" & (Menstruation == "3" & Age > 45) ~ "Post_menopause_women",
#     Sex == "Women" & (Menstruation == "3" & Age <= 45) ~ "Women_pre_menop_no_mens",
#     Sex == "Women" & (Menstruation == "1" | Menstruation == "2") ~ "Pre_menopause_women",
#     TRUE ~ "NA")) %>%
#   filter(Group != "Women_pre_menop_no_mens" & Group != "NA") %>%
#   droplevels() %>%
#   mutate(Group = ordered(Group, levels =  c("Pre_menopause_women", "Post_menopause_women", "Men")))
 #no_period puuttuu, tässäkö ongelma?

h2000_summary_final <- h2000_summary_final %>%
  mutate(Group = case_when(
    Sex == "Men" ~ "Men",
    Sex == "Women" & (Menstruation == "3" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "3" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "1" | Menstruation == "2") ~ "Pre_menopause_women", 
    Sex == "Women" & (Menstruation == "no_period") ~ "Post_menopause_women",
    TRUE ~ "NA")) %>%
  filter(Group != "Women_pre_menop_no_mens" & Group != "NA") %>%
  droplevels() %>%
  mutate(Group = ordered(Group, levels =  c("Pre_menopause_women", "Post_menopause_women", "Men")))
# ilmeisesti joo...

# h2000_summary_final <- h2000_summary_final %>% 
#   mutate(Group = case_when(
#     Sex == "Men" ~ "Men",
#     Sex == "Women" & (MENOP == "1") ~ "Post_menopause_women",
#     Sex == "Women" & (MENOP == "3") ~ "Pre_menopause_women",
#     TRUE ~ "NA")) %>%
#   filter(Group != "Women_pre_menop_no_mens" & Group != "NA") %>% 
#   droplevels() %>% 
#   mutate(Group = ordered(Group, levels =  c("Pre_menopause_women", "Post_menopause_women", "Men"))) 



```

In order to later merge the three cohorts into one table, we need to make colums for the variables missing variables in each cohort and impute values for them. 

The general population cohorts don't contain any information on blood donor activity. As seen in a previous FinDonor study the effect of full blood donation on ferritin levels is marginal after 1000 days (see Lobier et al. 2019, https://pubmed.ncbi.nlm.nih.gov/31408501/). After referring to these results we decided to assign a randomized donation interval between three to six years for the general population cohorts.  

```{r}
# Add new columns and impute values for them

blood_data_summary_final$CurrentPregnancy <- 0

fr1997_summary_final$TwoYearsFromStartCount_FB <- 0
fr1997_summary_final$HistoryOfIronSupplements <- NA
fr1997_summary_final$GivenIronSupplements <- 0
fr1997_summary_final$IronComplience <- 0
fr1997_summary_final$IronComplienceNumeric <- 0

h2000_summary_final$TwoYearsFromStartCount_FB <- 0
h2000_summary_final$HistoryOfIronSupplements <- NA
h2000_summary_final$GivenIronSupplements <- 0
h2000_summary_final$IronComplience <- 0
h2000_summary_final$IronComplienceNumeric <- 0
h2000_summary_final$RedMeat <- NA

# We then  assign a randomized donation interval for 3-6 years for the general population cohorts 
fr1997_summary_final <- fr1997_summary_final %>% mutate(DaysToPreviousFB = round(runif(nrow(fr1997_summary_final), min = 3*365, max = 6*365) ,0))
h2000_summary_final <- h2000_summary_final %>% mutate(DaysToPreviousFB = round(runif(nrow(h2000_summary_final), min = 3*365, max = 6*365) ,0))
```

We also need to rename observations for the menstruation variable in general population participants in order to match blood donors.

```{r}
fr1997_summary_final <- fr1997_summary_final %>% 
  mutate(Menstruation = case_when(
    Sex == "Women" & (Menstruation == "3") ~ "no_period",
    Sex == "Women" & (Menstruation == "2") ~ "irregular_period",
    Sex == "Women" & (Menstruation == "1") ~ "regular_period",
    TRUE ~ "NA")) %>% 
  mutate(Menstruation = ordered(Menstruation, levels =  c("regular_period", "irregular_period", "no_period"))) 

h2000_summary_final <- h2000_summary_final %>% 
  mutate(Menstruation = case_when(
    Sex == "Women" & (Menstruation == "3") ~ "no_period",
    Sex == "Women" & (Menstruation == "2") ~ "irregular_period",
    Sex == "Women" & (Menstruation == "1") ~ "regular_period",
    Sex == "Women" & (Age >= 55 & Group == "Post_menopause_women") ~ "no_period",
    TRUE ~ "NA")) %>% 
  mutate(Menstruation = ordered(Menstruation, levels =  c("regular_period", "irregular_period", "no_period"))) 

```

## Remove cohort participants with missing data

### Donation history

We remove `r blood_data_summary_final %>% filter(is.na(DaysToPreviousFB)) %>% nrow()` = 42 blood donors who have not donated previously (they are missing the number of days since last donation variable). 

```{r}
new_donors_data <- blood_data_summary_final %>% 
  filter(is.na(DaysToPreviousFB)) %>% 
    mutate(Group = case_when(Group == "Pre_menopause_women" ~ "Pre-menopausal women",
                   Group == "Post_menopause_women" ~ "Post-menopausal women",
                   Group == "Men" ~ "Men" ,
                   TRUE~ "NA"),
         Group = ordered(Group)) 

blood_data_summary_final <- blood_data_summary_final %>% 
  drop_na(DaysToPreviousFB)
```

### BMI

We remove `r cohort_name %>% filter(is.na(BMI)) %>% nrow()` participants for whom we do not have the BMI data:
 * FinDonor: 21 `r blood_data_summary_final %>% filter(is.na(BMI)) %>% nrow()`
 * FinRisk97: 8 `r fr1997_summary_final %>% filter(is.na(BMI)) %>% nrow()`
 * Health2000: 1 `r h2000_summary_final %>% filter(is.na(BMI)) %>% nrow()`

```{r}
#### FinDonor
nb_removed_blood <- findonor_nb_women_no_menstruation_response +
                            blood_data_summary_final %>% filter(is.na(BMI)) %>% nrow() 

blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!is.na(BMI))

#### FinRisk97
nb_removed_fr97 <- finrisk_nb_women_no_menstruation_response +
                            fr1997_summary_final %>% filter(is.na(BMI)) %>% nrow() 

fr1997_summary_final <- fr1997_summary_final %>% 
  filter(!is.na(BMI))

#### Health2000
nb_removed_h2000 <- h2000_nb_women_no_menstruation_response +
                            h2000_summary_final %>% filter(is.na(BMI)) %>% nrow() 

h2000_summary_final <- h2000_summary_final %>% 
  filter(!is.na(BMI))
```

### Smoking

We remove `r cohort_name %>% filter(is.na(BMI)) %>% nrow()` participants for whom we do not have smoking data:
  * FinDonor: 1      `r blood_data_summary_final %>% filter(is.na(Smoking)) %>% nrow()`
  * FinRisk97: 3111  `r fr1997_summary_final  %>% filter(is.na(Smoking)) %>% nrow()`
           * `r fr1997 %>%  filter(is.na(Q69)) %>%  nrow()` = 3377/7943 -> 43 % missing
  * Health2000: 2153 `r h2000_summary_final %>% filter(is.na(Smoking)) %>% nrow()` 
           * `r h2000 %>%  filter(is.na(FB05)) %>%  nrow()` = 2743/6264 -> 44 % missing
    
As removing n/a:s for smoking removes >5000 general population participants as opposed to only one in the blood donor cohort, we suspect there being something wrong with the original data. Looking at the original rdata makes it clear this is not a coding error. At this point we impute the n/a:s for the general population cohorts as non-smokers, while we wait for THL to explain the missing data. 

```{r}
# FinDonor
nb_removed_blood <- nb_removed_blood +
                            blood_data_summary_final %>% filter(is.na(Smoking)) %>% nrow() 

blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!is.na(Smoking))

# FinRisk97
fr1997_summary_final$Smoking[is.na(fr1997_summary_final$Smoking)]  <- 3

nb_removed_fr97 <- nb_removed_fr97 +
                            fr1997_summary_final %>% filter(is.na(Smoking)) %>% nrow()

fr1997_summary_final <- fr1997_summary_final %>%
  filter(!is.na(Smoking))

# Health2000
h2000_summary_final$Smoking[is.na(h2000_summary_final$Smoking)]  <- 3

nb_removed_h2000 <- nb_removed_h2000 +
                            h2000_summary_final %>% filter(is.na(Smoking)) %>% nrow()

h2000_summary_final <- h2000_summary_final %>%
  filter(!is.na(Smoking))

```

### Previous childbirth

We remove `r cohort_name %>% filter((group != "Men" & is.na(PreviousChildbirth)))%>% nrow()` female donors who did not answer the question on childbirth: 
  * FinDonor: 3 `r blood_data_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth)))%>% nrow()`
  * FinRisk97: 16 `r fr1997_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth)))%>% nrow()`
  * Health2000: 6 `r h2000_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth)))%>% nrow()`

```{r}
#### FinDonor
nb_removed_blood <- nb_removed_blood +
                            blood_data_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth))) %>% nrow 

blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!(Group != "Men" & is.na(PreviousChildbirth)))

#### FinRisk97
nb_removed_fr97 <- nb_removed_fr97 +
                            fr1997_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth))) %>% nrow 

fr1997_summary_final <- fr1997_summary_final %>% 
  filter(!(Group != "Men" & is.na(PreviousChildbirth)))

#### Health2000
nb_removed_h2000 <- nb_removed_h2000 +
                            h2000_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth))) %>% nrow 

h2000_summary_final <- h2000_summary_final %>% 
  filter(!(Group != "Men" & is.na(PreviousChildbirth)))

```

### Current pregnancy

Only women <45 yrs were asked whether they are currently pregnant and if so, how many weeks (source: https://www.julkari.fi/bitstream/handle/10024/78181/2005b6.pdf?sequence=1&isAllowed=y, page 56). This means we have a total of 1246 n/a:s for this specific question. We added a filter for age <45 years old, which leaves us with 3 missing answers. NA:s in >45-year-oldes are imputed as not pregnant. Although pregnancy is possible after this age, chances are lower. We used this age as our cut-off for menopause. 

We will now remove `r cohort_name %>% filter((Group != "Men" & is.na(CurrentPregnancy)))%>% nrow()` females who did not answer the pregnancy question: 
  * FinRisk97: 9 `r fr1997_summary_final %>% filter((Group != "Men" & is.na(CurrentPregnancy)))%>% nrow()`
  * Health00: 3 `r h2000_summary_final %>% filter((Group != "Men" & Age <=45 & is.na(CurrentPregnancy)))%>% nrow()`

```{r}
# FinRisk97
nb_removed_fr97 <- nb_removed_fr97 +
                            fr1997_summary_final %>% filter((Group != "Men" & is.na(CurrentPregnancy))) %>% nrow 

fr1997_summary_final <- fr1997_summary_final %>% 
  filter(!(Group != "Men" & is.na(CurrentPregnancy)))

# Health2000

# check age variance 
table(h2000_summary_final$CurrentPregnancy, h2000_summary_final$Age)

nb_removed_h2000 <- nb_removed_h2000 +
                            h2000_summary_final %>% filter((Group != "Men" & Age <=45 & is.na(CurrentPregnancy))) %>% nrow 

h2000_summary_final <- h2000_summary_final %>% 
  filter(!(Group != "Men" & Age <=45 & is.na(CurrentPregnancy)))

h2000_summary_final$CurrentPregnancy[is.na(h2000_summary_final$CurrentPregnancy)]  <- 0
```


### Iron supplementation 

We remove `r blood_data_summary_final %>% mutate(IronComplienceNumeric = ifelse(GivenIronSupplements == FALSE, 0, IronComplienceNumeric )) %>%  filter(is.na(IronComplienceNumeric)) %>% nrow()` = 77 donors that did not answer the two questions (did they receive iron supplements at during their last donation and supplement complience)

For the modelling, we impute a 0 (no supplementation) to iron_comp_c when the donor reports not being offered iron supplementation.

This is not done for the general population cohorts, as they were not asked about iron supplementation and as such all observation are input as 0 (these questions specifically concern iron supplements provided in conjuction with blood donation). 

```{r}
nb_removed_blood <- nb_removed_blood +
                            blood_data_summary_final %>% 
                              mutate(IronComplienceNumeric = ifelse(GivenIronSupplements == FALSE, 0, IronComplienceNumeric)) %>% 
                              filter(is.na(IronComplienceNumeric)) %>% 
                              nrow() 

blood_data_summary_final <- blood_data_summary_final %>% 
  mutate(IronComplienceNumeric = ifelse(GivenIronSupplements == FALSE, 0, IronComplienceNumeric )) %>% 
  filter(!is.na(IronComplienceNumeric))
```

### Red meat intake 

We remove `r cohort_name %>% filter(is.na(RedMeat)) %>% nrow()` participants for whom we do not have Smoking data:
  * Findonor: 10 `r blood_data_summary_final %>% filter(is.na(RedMeat)) %>% nrow()` 
  * FinRisk97: 30 `r fr1997_summary_final %>% filter(is.na(RedMeat)) %>% nrow()`  

Health2000 participants were not asked about red meat intake.

```{r message=FALSE, warning=FALSE}
# FinDonor
donors_to_remove <- blood_data_summary_final %>%
  dplyr::select(ID, Group, RedMeat) %>%
  gather(key= question, value = answer, -ID,-Group) %>%
  filter(is.na(answer)) %>%
  dplyr::select(Group, ID) %>%
  distinct(ID, Group)

blood_data_summary_final <- blood_data_summary_final %>%
  filter(!ID %in% donors_to_remove$ID)

nb_removed_blood <- nb_removed_blood +
                            donors_to_remove %>% nrow()

# FinRisk97
finrisk97_to_remove <- fr1997_summary_final %>%
  dplyr::select(ID, Group, RedMeat) %>%
  gather(key= question, value = answer, -ID,-Group) %>%
  filter(is.na(answer)) %>%
  dplyr::select(Group, ID) %>%
  distinct(ID, Group)

fr1997_summary_final <- fr1997_summary_final %>%
  filter(!ID %in% finrisk97_to_remove$ID)

nb_removed_fr97 <- nb_removed_fr97 +
                            finrisk97_to_remove %>% nrow()
```

### Age

We remove `r cohort_name %>% filter(is.na(Age)) %>% nrow()` participants with unknown age:
 * FinDonor: 0 `r blood_data_summary_final %>% filter(is.na(Age)) %>% nrow()`
 * FinRisk97: 0 `r fr1997_summary_final %>% filter(is.na(Age)) %>% nrow()`
 * Health2000: 0 `r h2000_summary_final %>% filter(is.na(Age)) %>% nrow()`

```{r}
#### FinDonor
nb_removed_blood <- findonor_nb_women_no_menstruation_response +
                            blood_data_summary_final %>% filter(is.na(Age)) %>% nrow() 

blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!is.na(Age))

#### FinRisk97
nb_removed_fr97 <- finrisk_nb_women_no_menstruation_response +
                            fr1997_summary_final %>% filter(is.na(Age)) %>% nrow() 

fr1997_summary_final <- fr1997_summary_final %>% 
  filter(!is.na(Age))

#### Health2000
nb_removed_h2000 <- h2000_nb_women_no_menstruation_response +
                            h2000_summary_final %>% filter(is.na(Age)) %>% nrow() 

h2000_summary_final <- h2000_summary_final %>% 
  filter(!is.na(Age))
```

The total number removed because of missing questionnaire data is:
`r nb_removed_blood` = 14
`r nb_removed_fr97` = 9
`r nb_removed_h2000` = 180

## Remove cohort participants who are currently pregnant

As current pregnancy was an exclusion criteria for the NL general population cohort and pregnant women are not allowed to donate blood, we will remove women who are currently pregnant from the FIN general population cohorts. The number of removed participants is 
`r fr1997_summary_final %>% filter(CurrentPregnancy == 2) %>% nrow()`
  * FinRisk97: 69 `r fr1997_summary_final %>% filter(CurrentPregnancy == 2) %>% nrow()`
  * Health00: 25 `r h2000_summary_final %>% filter(CurrentPregnancy == 1) %>% nrow()`

FinRisk97: 1=no, 2=yes
Health2000: 0=no, 1=yes

We will also recode the FinRisk97 cohort answer of 1=no to 0=no.  

```{r}
# FinRisk97
nb_removed_pregnant_fr1997 <- fr1997_summary_final %>% 
   filter(CurrentPregnancy == 2) %>% 
  nrow()

fr1997_summary_final <- fr1997_summary_final %>% 
   filter(!(Group != "Men" & CurrentPregnancy == 2))


fr1997_summary_final$CurrentPregnancy[fr1997_summary_final$CurrentPregnancy == "1"] <- 0

#Health2000
nb_removed_pregnant_h2000 <- h2000_summary_final %>% 
   filter((CurrentPregnancy == 1)) %>% 
  nrow()

h2000_summary_final <- h2000_summary_final %>% 
   filter(!(Group != "Men" & CurrentPregnancy == 1))
```

## Remove participants with extreme physiological measures

As decided previously, we remove data according the following criteria:

* BMI > 50
* CRP > 30
* Ferritin > 400 

This amounts to `r cohort_name %>% filter(BMI >= 50 | CRP >= 30 | Ferritin >= 400) %>% nrow()` participants that are removed. 
  * FinDonor: 12 `r blood_data_summary_final %>% filter(BMI >= 50 | CRP >= 30 | Ferritin >= 400) %>% nrow()`
  * FinRisk97: 216 `r fr1997_summary_final %>% filter(BMI >= 50 | CRP >= 30 | Ferritin >= 400) %>% nrow()`
  * Health00: 138 `r h2000_summary_final %>% filter(BMI >= 50 | CRP >= 30 | Ferritin >= 400) %>% nrow()`


```{r}
# FinDonor
blood_data_summary_final <- blood_data_summary_final %>% 
   filter(BMI < 50 & CRP < 30 & Ferritin < 400)

# FinRisk97
fr1997_summary_final <- fr1997_summary_final %>% 
   filter(BMI < 50 & CRP < 30 & Ferritin < 400)

#Health2000
h2000_summary_final <- h2000_summary_final %>% 
   filter(BMI < 50 & CRP < 30 & Ferritin < 400)

```

## Final group N for the three cohorts

```{r}
# FinDonor
blood_data_summary_final %>% 
   mutate(Group = dplyr::recode(Group, Pre_menopause_women = "Pre-menopausal women",
         Post_menopause_women = "Post-menopausal women")) %>% 
  group_by(Group) %>% 
  summarise ( N = n()) %>% 
  kable() 

# FinRisk97
fr1997_summary_final %>% 
   mutate(Group = dplyr::recode(Group, Pre_menopause_women = "Pre-menopausal women",
         Post_menopause_women = "Post-menopausal women")) %>% 
  group_by(Group) %>% 
  summarise ( N = n()) %>% 
  kable() 

# Health2000
h2000_summary_final %>% 
   mutate(Group = dplyr::recode(Group, Pre_menopause_women = "Pre-menopausal women",
         Post_menopause_women = "Post-menopausal women")) %>% 
  group_by(Group) %>% 
  summarise ( N = n()) %>% 
  kable() 

 
```

Next we recode red meat consumption into a linear variable. We use a linear scale from 1 to 4 for and mapped the different responses to correspond to the four options for the FinDonor and FinRisk97 cohorts. Data on red meat consumption was not avaliable in the Health2000 cohort.  

* red meat:
    * FinDonor
        * "never" ~ 1,
              # never
        * "less_than_once_weekly" ~ 2,
              # less than once a week
        * "1.3_week" or "4.6_week"~ 3,
              # 1-3 times a week
              # 4-6 times a week
        * "daily" or "several_daily" ~ 4,
              # daily
              # several times a day
    * FinRisk97
        * "1" ~ 1, 
              # 1 = more seldom than once a month or not at all
        * "2" ~ 2,
              # 2 = once or twice a month
        * "3" or "4" or "5" ~ 3,
              # 3 = once a week
              # 4 = twice a week or more often
              # 5 = almost daily
        * "5" or "several_daily" ~ 4,
              # 6 = daily
              # 7 = several times a day
              
This will create a new variable, RedMeat_n. As we don't have red meat data in the Health2000 cohort, we need to make a column, name it RedMeat_n and impute the observations as n/a.

```{r}
# FinDonor
blood_data_summary_final <- blood_data_summary_final %>% 
  mutate(RedMeat_n = case_when(
    RedMeat == "never" ~ 1,
    RedMeat == "less_than_once_weekly" ~ 2,
    RedMeat %in% c("1.3_week" , "4.6_week") ~ 3,
    TRUE   ~ 4 )) 

# FinRisk97
fr1997_summary_final <- fr1997_summary_final %>% 
  mutate(RedMeat_n = case_when(
    RedMeat == "1" ~ 1,
    RedMeat == "2" ~ 2,
    RedMeat %in% c("3" , "4", "5") ~ 3,
    TRUE   ~ 4 )) # "6" | "7"
  
# Health2000
h2000_summary_final$RedMeat_n <- NA


```

Due to difference in the CRP measurements (hs-CRP for the general population cohorts, CRP for the blood donor cohort) we decided to impute CRP <3 mg/l to 2.9 mg/l for the general population cohorts in order to align them with the blood donor data. CRP <3 mg/l was previously imputed as 2.9 mg/ml in the blood donor cohort.

```{r}
# FinRisk97
fr1997_summary_final$CRP[fr1997_summary_final$CRP < "3.0"] <- 2.9

# Health2000
h2000_summary_final$CRP[h2000_summary_final$CRP < "3.0"] <- 2.9
```

Previous childbirth is recoded 
* FinDonor: have you given birth? no=no, yes=yes
* FinRisk97: how many children have you given birth to? 1=none, 2=one, 3=two, 4=three or more
* Health2000: have you given birth? 0=no, 1=yes

Nulliparous women are coded as 0 (no) and reporting childbirth or giving birth to any number of children as 1 (yes). 

```{r}
# FinDonor
blood_data_summary_final <- blood_data_summary_final %>% 
  mutate(PreviousChildbirth = case_when(
    PreviousChildbirth == "no" ~ 0,
    PreviousChildbirth == "yes" ~ 1)) 

# FinRisk97
fr1997_summary_final <- fr1997_summary_final %>% 
  mutate(PreviousChildbirth = case_when(
    PreviousChildbirth == "1" ~ 0,
    PreviousChildbirth %in% c("2" , "3", "4") ~ 1)) 
    

```


## Merging cohorts into one table 

For the code to work all three cohorts need to be merged into one single table. Before this can be done variables the variables selected need to have the same name.  

```{r}
### mutate smoking and pregnancy 
blood_data_summary_final <- blood_data_summary_final %>% 
  dplyr::select(ID, Group, Sex, Age, TwoYearsFromStartCount_FB, DaysToPreviousFB,
                Ferritin, Hb, BMI, Menstruation, Smoking, RedMeat_n, IronComplienceNumeric, PreviousChildbirth,
                CRP, Cohort) %>% 
    mutate(Smoking = ifelse(Smoking == "daily", "yes", "no"),
         Smoking = factor(Smoking, levels = c( "no",  "yes")))

### mutate smoking and pregnancy 
fr1997_summary_final <- fr1997_summary_final %>% 
  dplyr::select(ID, Group, Sex, Age, TwoYearsFromStartCount_FB, DaysToPreviousFB,
                Ferritin, Hb, BMI, Menstruation, Smoking, RedMeat_n, IronComplienceNumeric, PreviousChildbirth,
                CRP, Cohort) %>% 
       mutate(Smoking = ifelse(Smoking == "1", "yes", "no"),
       Smoking = factor(Smoking, levels = c( "no",  "yes")))

### mutate smoking and pregnancy
h2000_summary_final <- h2000_summary_final %>% 
  dplyr::select(ID, Group, Sex, Age, TwoYearsFromStartCount_FB, DaysToPreviousFB,
                Ferritin, Hb, BMI, Menstruation, Smoking, RedMeat_n, IronComplienceNumeric, PreviousChildbirth,
                CRP, Cohort) %>% 
         mutate(Smoking = ifelse(Smoking == "1", "yes", "no"),
         Smoking = factor(Smoking, levels = c( "no",  "yes")))
         

# merge the cohorts

summary_all_cohorts <- bind_rows(blood_data_summary_final, fr1997_summary_final, h2000_summary_final)

```

## Final group N for the summarized table

```{r}
summary_all_cohorts %>% 
   mutate(Group = dplyr::recode(Group, Pre_menopause_women = "Pre-menopausal women",
         Post_menopause_women = "Post-menopausal women")) %>% 
  group_by(Group) %>% 
  summarise ( N = n()) %>% 
  kable() 
 
```


```{r}
ggplot(data = summary_all_cohorts,
       mapping = aes(x = Ferritin,
                     y = Group)) + 
  geom_point()


ggplot(summary_all_cohorts, aes(Ferritin, Group)) +
  geom_point()+
  geom_line()

summary_all_cohorts %>% 
  ggplot(aes(Ferritin, BMI,
             colour = Cohort))+
  geom_point(alpha = 0.5)+ 
  facet_wrap(~Group)

summary_all_cohorts %>% 
  ggplot(aes(Group, Ferritin))+
  geom_boxplot()+
  geom_point(size = 0.1)+
  facet_wrap(~Cohort)

summary_all_cohorts %>% 
  ggplot(aes(Ferritin, BMI,
             colour = Cohort))+
  geom_point(size = 0.1, 
             alpha = 0.5)+
  facet_wrap(~Group)
  
```



# Results - Regression analysis

## Regressions on ferritin  

We analyzed the data with multiple linear regression. We first ran Ordinary Least Square (OLS) regressions and their diagnostics. If the diagnostics identified problematic observations (e.g., outliers, influential observations), we also ran robust regression regressions [@yu2017robust]. Robust regression is based on less assumptions than OLS regression and is less sensitive to influential observations. It  gives more accurate estimates of regression coefficients in the presence of problematic observations. Using robust regression leads to a more principled approach to regression analysis and avoids potential confounds introduced by selective removal of influential observations. As the rlm command from the MASS package does not compute directly the *t* and *p* values, we used the f.robftest function from the sfsmisc package. Practical examples of using robust regression can be found in the literature [@davison2017platelet].  


In addition, we used relative importance analysis to estimate the average proportion of variance in the outcome variable explained by each co-variate [@Groemping2006]. Relative iportance was computed using the pmvd method which assigns 0 to the relative importance of a regressor if it is non significant in the complete model. .

The regressors were entered as follows in the regression models for ferritin:

* Continuous variables were entered as continuous predictors
    * age, number of donations in last two years, nb of days since last donation, CRP, BMI
  
* Binary variables were entered as categorical (but coded as numerical for relative importance):
    * Smoking, pregnancy
  
* Ordinal variables were coded on a linear scale and entered as continuous [(Reference)](https://www3.nd.edu/~rwilliam/stats3/OrdinalIndependent.pdf)
    * Iron supplementation 
    * Meat intake
  
We compute both coefficients and standardized coefficients (with all dependent and independent variables standardized) for robust regression and only coefficients for OLS regression. We present robust coefficients and their bootstrapped BCa 95% confidence intervals in Table 1 (ferritin) and  Table 2 (sTfR), robust standardized coefficients, their bootstrapped distribution and BCa 95% confidence intervals in Figure 2.A and non-standardized OLS regression coefficients in Supplementary Table 1 (Ferritin) and 2 (sTfR). 

## Pre-processing

### Rename and transform

```{r}
regression_cohorts <- summary_all_cohorts %>% 
    rename(donation_count = TwoYearsFromStartCount_FB,
            last_donation = DaysToPreviousFB,
            iron_complience = IronComplienceNumeric) %>% 
     mutate(Age = Age / 5, 
            donation_count_2 = donation_count^2,
            log_ferritin = log(Ferritin),
            log_last_donation = log(last_donation)/log(2),
            log_CRP = log(CRP)) 

```

Data transformations:

* Log transformed variables: 
    * Ferritin
    * CRP 
  
* Age is divided by 5 to simplify coefficient interpretation
* Number of days before donation is transformed as $log\_last\_donation = log(last\_donation)/log(2)$ to help with the interpretation of coefficients. 
    * An increase in one of the transformed variable is equivalent to a doubling of the number of days since last donation.

* Standardization:
    * Standardized coefficients: 
        * All dependent and independent variables were standardized
      
    * Coefficients:
        * All dependent variables entered as continuous varaibles are centered but not scaled. 

We center all variables entered as continuous.

## Premenopausal women


```{r}

test_data_pre_women <- regression_cohorts %>% 
  filter(Group == "Pre_menopause_women")

test_data_pre_women %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 


```

## Postmenopausal women

```{r}

test_data_post_women <- regression_cohorts %>% 
  filter(Group == "Post_menopause_women") 

test_data_post_women %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2)

```


## Men


```{r}

test_data_men <- regression_cohorts %>% 
  filter(Group == "Men") 

test_data_men %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 


```



### Mutiple regression - ferritin as outcome

#### Correlograms

```{r}
# Premenopausal women
ggpairs(test_data_pre_women, 
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)

# Postmenopausal women
ggpairs(test_data_post_women, 
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)

# Men
ggpairs(test_data_men, 
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)



```

```{r}
#same as above, cohorts separated 

# Premenopausal women
ggpairs(test_data_pre_women, ggplot2::aes(colour = Cohort, alpha = 0.5),
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE) + 
  ggtitle("Premenopausal women")

# Postmenopausal women
ggpairs(test_data_post_women, ggplot2::aes(colour = Cohort, alpha = 0.5),
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE) +
  ggtitle("Postmenopausal women")

# Men
ggpairs(test_data_men, ggplot2::aes(colour = Cohort, alpha = 0.5),
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE) + 
ggtitle("Men")

```


```{r}

ggpairs(test_data_pre_women, ggplot2::aes(colour = Cohort, alpha = 0.5),
        columns = c("log_ferritin", "PreviousChildbirth",  "RedMeat_n", "Smoking", "iron_complience", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE) + 
ggtitle("Premenopausal women")

ggpairs(test_data_post_women, ggplot2::aes(colour = Cohort, alpha = 0.5), 
        columns = c("log_ferritin", "PreviousChildbirth",  "RedMeat_n", "Smoking", "iron_complience", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)+ 
ggtitle("Postmenopausal women")

ggpairs(test_data_men, ggplot2::aes(colour = Cohort, alpha = 0.5),
        columns = c("log_ferritin",  "RedMeat_n", "Smoking", "iron_complience", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE) + 
ggtitle("Men")

```


#### Correlograms by cohort

##### Premenopausal women 

###### FinDonor

```{r}
test_data_pre_women_findonor <- regression_cohorts %>% 
  filter(Group == "Pre_menopause_women" & Cohort == "FINDONOR") 

test_data_pre_women_findonor %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2)

ggpairs(test_data_pre_women_findonor, 
        columns = c("log_ferritin", "PreviousChildbirth",  "RedMeat_n", "Smoking", "iron_complience", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)

ggpairs(test_data_pre_women_findonor, 
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)


```

###### FinRisk97

```{r}
test_data_pre_women_fr1997 <- regression_cohorts %>% 
  filter(Group == "Pre_menopause_women" & Cohort == "FINRISK97") 

test_data_pre_women_fr1997 %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2)

ggpairs(test_data_pre_women_fr1997, 
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)

ggpairs(test_data_pre_women_fr1997, 
        columns = c("log_ferritin", "PreviousChildbirth",  "RedMeat_n", "Smoking", "iron_complience", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)


```

###### Health2000

```{r}
test_data_pre_women_h2000 <- regression_cohorts %>% 
  filter(Group == "Pre_menopause_women" & Cohort == "HEALTH00") 

test_data_pre_women_h2000 %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2)

ggpairs(test_data_pre_women_h2000, 
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)

ggpairs(test_data_pre_women_h2000, 
        columns = c("log_ferritin", "PreviousChildbirth", "Smoking", "iron_complience", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)


```



##### Postmenopausal women 

###### FinDonor

```{r}
test_data_post_women_findonor <- regression_cohorts %>% 
  filter(Group == "Post_menopause_women" & Cohort == "FINDONOR") 

test_data_post_women_findonor %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2)

ggpairs(test_data_post_women_findonor, 
        columns = c("log_ferritin", "PreviousChildbirth",  "RedMeat_n", "Smoking", "iron_complience", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)

ggpairs(test_data_post_women_findonor, 
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)


```

###### FinRisk97

```{r}
test_data_post_women_fr1997 <- regression_cohorts %>% 
  filter(Group == "Post_menopause_women" & Cohort == "FINRISK97") 

test_data_post_women_fr1997 %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2)

ggpairs(test_data_post_women_fr1997, 
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)

ggpairs(test_data_post_women_fr1997, 
        columns = c("log_ferritin", "PreviousChildbirth",  "RedMeat_n", "Smoking", "iron_complience", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)


```

###### Health2000

```{r}
test_data_post_women_h2000 <- regression_cohorts %>% 
  filter(Group == "Post_menopause_women" & Cohort == "HEALTH00") 

test_data_post_women_h2000 %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2)

ggpairs(test_data_post_women_h2000, 
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)

ggpairs(test_data_post_women_h2000, 
        columns = c("log_ferritin", "PreviousChildbirth",  "RedMeat_n", "Smoking", "iron_complience", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)


```

##### Men 

###### FinDonor

```{r}
test_data_men_findonor <- regression_cohorts %>% 
  filter(Group == "Men" & Cohort == "FINDONOR") 

test_data_men_findonor %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2)

ggpairs(test_data_men_findonor, 
        columns = c("log_ferritin", "RedMeat_n", "Smoking", "iron_complience", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)

ggpairs(test_data_men_findonor, 
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)


```

###### FinRisk97

```{r}
test_data_men_fr1997 <- regression_cohorts %>% 
  filter(Group == "Men" & Cohort == "FINRISK97") 

test_data_men_fr1997 %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2)

ggpairs(test_data_men_fr1997, 
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)

ggpairs(test_data_men_fr1997, 
        columns = c("log_ferritin",  "RedMeat_n", "Smoking", "iron_complience", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)


```

###### Health2000

```{r}
test_data_men_h2000 <- regression_cohorts %>% 
  filter(Group == "Men" & Cohort == "HEALTH00") 

test_data_men_h2000 %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2)

ggpairs(test_data_men_h2000, 
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)

ggpairs(test_data_men_h2000, 
        columns = c("log_ferritin", "Smoking", "iron_complience", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)


```

```{r}
ggpairs(test_data_men, ggplot2::aes(colour = Cohort),
        columns = c("log_ferritin", "BMI"),
        lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

