---
title: "Low ferritin subgroup analysis"   
author: "Sofie Ekroos"
date: "21.11.2021"
output: pdf_document

---

# Summary

This code is based on Mikko Arvas' Bayesian model for iron deficiency, "Low ferritin model" (https://github.com/FRCBS/iron_measurement_comparisons/blob/master/src/low_ferritin_model_bayes.Rmd). In this analysis we analyse iron deficiency risk factors in pre-menopausal women.

```{r setup, , message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
# echo "rmarkdown::render('low_ferritin_model_bayes_subgroup_analysis.Rmd', clean=TRUE,output_format='pdf_document',output_file='../results/low_ferritin_model_bayes_subgroup_analysis.pdf')" | R --slave

library(tidyverse)
library(brms)
library(bayesplot)
```



PART 2: BAYESIAN LOGISTIC REGRESSION

# Data loading and preparation 

Load data

```{r}
load("../data/ID_data_regression_cohorts.rdata")
all_cohorts <- regression_cohorts %>% filter(Group == "Pre_menopause_women")

all_cohorts <- all_cohorts %>% 
  filter(!is.na(CRP)) %>% 
  filter(! CRP >= 30)
```



# Build subgroups of pre-menopausal women (three cohorts: FinDonor, FinRisk97, Health2000) 

## Data
```{r}
# FinDonor
regression_data_blood <- all_cohorts %>% 
  filter(Cohort == "FINDONOR") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, iron_deficiency, Smoking, iron_complience, Region, Weight, Menstruation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         Weight = scale(Weight, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# FinRisk97
regression_data_fr1997 <- all_cohorts %>% 
  filter(Cohort == "FINRISK97") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, iron_deficiency, Smoking, iron_complience, Region, Weight, Menstruation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         Weight = scale(Weight, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# Health2000 
regression_data_h2000 <- all_cohorts %>% 
  filter(Cohort == "HEALTH00") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, iron_deficiency, Smoking, iron_complience, Region, Weight, Menstruation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         Weight = scale(Weight, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 


```

## Models 

### FinDonor

The Region-variable only has one observation (HKI), since all blood donors are from Helsinki or Espoo. As the code will not run as long as Region is included, it was removed for the FinDonor cohort by imputing all observations to NA (we later find out just removing it won't allow the code to run when combining).

Even later it turns out that imputing NA:s won't work in this stage of building the model, so it was removed completely.

```{r}
#trying to see if we can get region to work, uncomment this if it doesn't work

#regression_data_blood <- regression_data_blood

#regression_data_blood$Region <- NA
```


```{r blood}
file <- "../results/bmodels_subgroup/logit_pre_blood_b"

logistic_pre_blood_b <- brm(iron_deficiency ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        iron_complience + 
                        donation_count_2 + 
                        log_last_donation + 
                        log_CRP +
                        #Region + 
                        Weight +
                        Menstruation, 
                      data = regression_data_blood, 
                      family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(logistic_pre_blood_b)

```

FinRisk97

```{r fr1997}
file <- "../results/bmodels_subgroup/logit_pre_fr1997_b"

logistic_pre_fr1997_b <- brm(iron_deficiency ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        #iron_complience + 
                        #donation_count_2 + 
                        log_last_donation + 
                        log_CRP +
                        Region + 
                        Weight +
                        Menstruation, 
                      data = regression_data_fr1997, 
                      family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(logistic_pre_fr1997_b)


```

Hmm, there are problems with this cohort. Warning message: 

There were 5357 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10. See
http://mc-stan.org/misc/warnings.html#maximum-treedepth-exceededExamine the pairs() plot to diagnose sampling problems
The largest R-hat is 1.92, indicating chains have not mixed.
Running the chains for more iterations may help. See
http://mc-stan.org/misc/warnings.html#r-hatBulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.
Running the chains for more iterations may help. See
http://mc-stan.org/misc/warnings.html#bulk-essTail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable.
Running the chains for more iterations may help. See
http://mc-stan.org/misc/warnings.html#tail-essParts of the model have not converged (some Rhats are > 1.05). Be careful when analysing the results! We recommend running more iterations and/or setting stronger priors. Family: bernoulli 

--> solved this by removing the variables iron_complience and donation_count_2 (found these to mess up the diagnostics in the later stages). Donation count were imputed as a randomized number between 3-6 years in the general population cohorts (see low_ferritin_data) as we didn't have this information. We also don't have data on iron supplementation in the general population cohorts so all observations were imputed as 0.

## Health00

```{r h2000}
file <- "../results/bmodels_subgroup/logit_pre_h2000_b"

logistic_pre_h2000_b <- brm(iron_deficiency ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        #iron_complience + 
                        #donation_count_2 + 
                        log_last_donation + 
                        log_CRP +
                        Region + 
                        Weight +
                        Menstruation, 
                      data = regression_data_h2000, 
                      family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(logistic_pre_h2000_b)
``` 

There are problems with this cohort as well, same warning message. Solved this as explained for FinRisk97.

## Diagnostics

### Trace

```{r , message=FALSE}
mcmc_plot(logistic_pre_blood_b, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(logistic_pre_fr1997_b, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(logistic_pre_h2000_b, type = "trace")
```

Chains do not seem to get stuck now that the variables iron_complience and donation_count_2 were removed in both general population cohorts. 

### ACF bars

```{r}
mcmc_plot(logistic_pre_blood_b, type = "acf_bar")
```


```{r}
mcmc_plot(logistic_pre_fr1997_b, type = "acf_bar")
```


```{r}
mcmc_plot(logistic_pre_h2000_b, type = "acf_bar")
```

Autocorrelation drops in a satisfactory way.

### Rhat

```{r}
mcmc_plot(logistic_pre_blood_b, type = "rhat")
```

```{r}
mcmc_plot(logistic_pre_fr1997_b, type = "rhat")
```

```{r}
mcmc_plot(logistic_pre_h2000_b, type = "rhat")
```
https://mc-stan.org/rstan/reference/Rhat.html


"The Rhat function produces R-hat convergence diagnostic, which compares the between- and within-chain estimates for model parameters and other univariate quantities of interest. If chains have not mixed well (ie, the between- and within-chain estimates don't agree), R-hat is larger than 1. We recommend running at least four chains by default and only using the sample if R-hat is less than 1.05."

Rhat looks good as well.

## Forest plots

```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(logistic_pre_blood_b, 
         type = "areas",
         prob = 0.95,
         transformations = "exp") +
  geom_vline(xintercept = 1, color = "grey")

p <- p + scale_x_log10()
#Ferritin needs to be removed
# plot.OR<-  tidy_logit_pre_menop_women %>%filter(!term=="log_Ferritin_beginning") %>%  dplyr::select(OR) %>% mutate(y=rev(2:(nrow(tidy_logit_pre_menop_women)))) %>% bind_rows(tibble(OR=NA,y=1 ))
# p <- p + geom_point(aes(y=y,x=OR),data=plot.OR, col="red")
p
```

```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(logistic_pre_fr1997_b, 
         type = "areas",
         prob=0.95,
         prob_est="median",
         transformations = "exp"
        ) +
  geom_vline(xintercept = 1, color = "grey")

p <- p + scale_x_log10()
p
```

```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(logistic_pre_h2000_b, 
         type = "areas",
         prob = 0.95,
         transformations = "exp") +
  geom_vline(xintercept = 1, color = "grey")

p <- p + scale_x_log10()
#Ferritin needs to be removed
# plot.OR<-  tidy_logit_pre_menop_women %>%filter(!term=="log_Ferritin_beginning") %>%  dplyr::select(OR) %>% mutate(y=rev(2:(nrow(tidy_logit_pre_menop_women)))) %>% bind_rows(tibble(OR=NA,y=1 ))
# p <- p + geom_point(aes(y=y,x=OR),data=plot.OR, col="red")
p
```


These all look strange compared to the ones in the main analysis (men and post-menopausal women included), not sure about the two colours, is it the CI?


### All groups together

```{r}
regressor_values <- c(
 "Age" = "Age",
   "log_CRP" = "log(CRP)",
   "BMI" = "BMI",
   "log_last_donation" = "log(days to last donation)",
   "Smokingyes" ="Smoker",
    "iron_complience" = "Iron compliance",
    "donation_count_2" = "Donations in last 2 years",
    "Menstruationregular_period" = "Regular periods",
    "Weight" = "Weight",
    "Region" = "Region"
 
)
```




```{r }

intervals_pre_blood <- mcmc_intervals_data(
  logistic_pre_blood_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median",
  transformations = "exp"
  ) %>% mutate(Cohort='FINDONOR')

intervals_pre_fr1997 <- mcmc_intervals_data(
  logistic_pre_fr1997_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median",
  transformations = "exp"
  ) %>% mutate(Cohort='FINRISK97') %>% 
  arrange(m) # how to order the variables?

intervals_pre_h2000 <- mcmc_intervals_data(
  logistic_pre_h2000_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median",
  transformations = "exp"
  ) %>% mutate(Cohort='HEALTH00')

intervals_combined <- rbind(intervals_pre_blood,intervals_pre_fr1997,intervals_pre_h2000) %>% 
  filter(! (parameter == 'exp(b_Intercept)' | parameter == 'exp(lp__)')) %>% 
  mutate(parameter = gsub("\\)","",gsub("exp\\(b_","",parameter))) %>% 
  mutate(regressor = plyr::revalue(parameter, regressor_values)) %>% 
  select(-parameter)
  #arrange(m)


```


```{r , fig1, fig.height = 10, fig.width = 15 ,warning=FALSE}

orlower <- 1/10
orupper <- 10

# Truncate ORs for plotting
plotdata <- intervals_combined %>% 
  mutate(
    ll = case_when(
      ll < orlower ~ orlower,
      TRUE ~ ll
    ),
    hh = case_when(
      hh > orupper ~ orupper,
      TRUE ~ hh
    )
  )

#theme_set(bayesplot::theme_default())
pos <- position_nudge(y=as.numeric(as.factor(plotdata$Cohort))/5 -0.1) 

p <- ggplot(plotdata, aes(x = m, y = regressor, color = Cohort)) + 
  geom_point(position = pos,
             size=4) +
  geom_linerange(aes(xmin = ll, xmax = hh), 
                 position = pos,
                 size = 1, 
                 linetype = 1
                 ) +
  scale_color_manual(values = c("#84b56f", "#88b5bc", "#a2c4c9"),
                     limits = c("FINDONOR",  "FINRISK97","HEALTH00")) +
  scale_fill_manual(values = c("#84b56f", "#88b5bc", "#a2c4c9"),
                    limits = c("FINDONOR",  "FINRISK97","HEALTH00")) +
  scale_shape_manual(values = c(1,16)) +
  scale_x_log10(limits=c(orlower,orupper))  + #how to get the lines that drop out of the visible?  
  geom_vline(xintercept = 1 , color = "grey", linetype = "dotted",
             size = 1) +
  theme(axis.title = element_text(size = 14),
        strip.text =  element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.background = element_rect(fill = "white"),
        strip.background.x = element_blank(),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14)
  ) +
  guides(shape = "none") +
  theme_classic() +
  ylab("Coefficient") 

p


```

```{r}


ggplot2::ggsave(filename = "../results/figures/subgroup_analysis/forest_pre_women_all_cohorts_bayes.pdf",
                plot = p,
                width = 35,
                height = 25,
                dpi = 600,
                units = "cm")


```


## Compare to bootstrapped

This need to be redone once we have done a subgroup analysis in the bootsrapped code, as of now we don't have one. (change group to cohort!)

```{r  fig2, fig.height = 15, fig.width = 10 ,warning=FALSE}
file <- "../results/bootstraps/plotdata_pre_women_filtered.rdata" #this I already changed, nothing else has been done
# if(file.exists(file)) {
#   load(file)
#   temp <- plotdata_filtered %>% select(-p.value,-is_sig) %>% 
#     rename(ll = Bca_inf,
#            hh = Bca_sup,
#            m = OR
#     ) %>%  mutate(
#       Method = "Bayes"
#     ) %>%  rename(Group=group)
#   
#   temp <- plotdata %>%  select(regressor,ll,hh,m,Group) %>% 
#     mutate(Method="Boot") %>% 
#   bind_rows(temp)
#   
#   temp <- temp %>% 
#     mutate(
#       ll = case_when(
#         ll < orlower ~ orlower,
#         TRUE ~ ll
#       ),
#       hh = case_when(
#         hh > orupper ~ orupper,
#         TRUE ~ hh
#       )
#     )
#   
#   pos <- position_nudge(y=as.numeric(as.factor(temp$Method))/5 -0.1) 
#   
#   p <- ggplot(temp, aes(x = m, y = regressor, color = Method)) + 
#      geom_point(position = pos,
#                 size=4) +
#     geom_linerange(aes(xmin = ll, xmax = hh), 
#                    position = pos,
#                    size = 1 
#     ) +
#     # scale_color_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
#     #                    limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
#     # scale_fill_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
#     #                   limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
#     #scale_shape_manual(values = c(1,16)) +
#     scale_x_log10(limits=c(orlower,orupper))  + #how to get the lines that drop out of the visible?  
#     geom_vline(xintercept = 1 , color = "grey", linetype = "dotted",
#                size = 1) +
#     theme(axis.title = element_text(size = 14),
#           strip.text =  element_text(size = 14),
#           legend.text = element_text(size = 12),
#           strip.background = element_rect(fill = "white"),
#           strip.background.x = element_blank(),
#           axis.text.x = element_text(size = 14),
#           axis.text.y = element_text(size = 14)
#     ) +
#     guides(shape = "none") +
#     theme_classic() +
#     facet_grid(Group~.) +
#     ylab("Coefficient") 
#   
#   p
#   
#   
# }
```


```{r}
# if(file.exists(file)) {
# 
# ggplot2::ggsave(filename = "../results/figures/forest_all_cohorts_bayes_vs_boot.pdf",
#                 plot = p,
#                 width = 35,
#                 height = 30,
#                 dpi = 600,
#                 units = "cm")
# }

```