---
title: "Low ferritin model"
author: "Sofie Ekroos"
date: "19.7.2021"
output: github_document
bibliography: index.bib
---

Based on the original code written by Muriel Lobier and Mikko Arvas (https://github.com/FRCBS/iron_levels_of_blood_donors/blob/master/src/index.Rmd and https://github.com/FRCBS/changes_in_donor_health/blob/master/src/explorative_analysis.Rmd)

# Summary

This document includes all codes necessary to run the analysis of and produce the figures for the three Finnish Cohorts (FinDonor, FINRISK97, Health2000). The code allows the user to:

- describe the cohorts and compute the prevalence of iron deficiency
- identify co-variants of iron deficiency using multivariable linear regression 

```{r setup, , message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# library(janitor)
library(tableone)
library(stargazer)
library(gridExtra)
library(cowplot)
library(broom)
library(GGally)
library(ggfortify)
# library(sjPlot)
library(knitr)
library(lubridate)
#library(HDInterval) !! package ‘HDIinterval’ is not available (for R version 3.6.3)!!!
library(car)
#library(relaimpo)
library(sfsmisc)
library(MASS)
# library(kableExtra)
#library(brms)
#library(rstan)
#library(rms)
#library(modelr)
#library(tidybayes)
library(ordinal)
library(sjmisc)
#library(sjlabelled)
# library(lazerhawk)
library(epiDisplay)
library(safeBinaryRegression)
library(epitools)
library(ez)
library(ggbeeswarm)
library(tidymodels)
```

# Data loading and preparation

## Data loading

```{r , warning=FALSE}

# load FinDonor biomarker data
load("../data/r02.fd.bd.all.rdata")
indiv_donations_data <- output
rm(output)

# load FinDonor demographic data
load("../data/r02ds.donorData.rdata")
donor_demo <- output
rm(output)

# load THL data
load("../data/thldata.rdata")

# thldalta.rdata contains all five THL cohorts, extract FINRISK97 and Health2000 from the others 
fr1997 <- thldata$fr1997
h2000 <- thldata$h2000
rm(thldata)
```

## Data preparation

There are `r cohort_name %>% distinct(ID_variable) %>% nrow()` individials enrolled in the studies. 

  * FinDonor: 2584   `r indiv_donations_data %>% distinct(donor) %>% nrow()`
  * FinRisk97: 7943  `r fr1997 %>% distinct(RELEASE_ID) %>% nrow()`
  * Health2000: 6264 `r h2000 %>% distinct(RELEASE_ID) %>% nrow()` 
        
There are `r cohort_name %>%  filter(!is.na(Ferritin) & !is.na(Hb_v) ) %>% distinct(ID_variable) %>% nrow()` left once we remove donors that have no Ferritin or no Hb measurement for any donation event. 

  * FinDonor: 2580 `r indiv_donations_data %>%  filter(!is.na(Ferritin) & !is.na(Hb_v) ) %>% distinct(donor) %>% nrow()`
  
  * FinRisk97: 462 `r fr1997 %>%  filter(!is.na(FERRITIN) & !is.na(HGB) ) %>% distinct(RELEASE_ID) %>% nrow()`
            * 7380 `r fr1997 %>% filter(!is.na(FERRITIN)) %>% distinct(RELEASE_ID) %>% nrow()`
            * 491 `r fr1997 %>%  filter(!is.na(HGB)) %>% distinct(RELEASE_ID) %>% nrow()`
              
  * Health2000: 5250 `r h2000 %>%  filter(!is.na(FERRITIINI) & !is.na(B_Hb) ) %>% distinct(RELEASE_ID) %>% nrow()`

At this stage we noticed we only had hemoglobin data from 491 FinRisk97 participants. Using the table function (`r table(fr1997$HGB, fr1997$SUKUP)`, SUKUP = sukupuoli = sex, 1=male, 2=female) we found out that all these participants are male. As such, we didn't remove N/A:s for Hb and will not use Hb in the model for this specific cohort.  

```{r , warning=FALSE}
#FINDONOR
## Assign all blood donors to FINDONOR cohort
indiv_donations_data$Cohort <- c("FINDONOR")

## Get sex for each blood donor
blood_data_summary <- indiv_donations_data %>% 
  group_by(donor) %>% 
  summarise(Sex=first(gender)) 

## Get values for first study donation with the required measurements donation (regardless of donation type)
## We remove events with no Ferritin and hb_v to be inline with from when the nb of previous donations were counted.
blood_values_init <- indiv_donations_data %>% 
  filter(!is.na(Ferritin) & !is.na(Hb_v) ) %>% 
  group_by(donor) %>% 
  filter(date == min(date)) %>% 
  dplyr::select(age, Ferritin, Hb_v, CRP, DaysToPreviousFB, donor, Cohort) %>% 
  ungroup()

blood_data_summary <- blood_data_summary %>% 
  full_join(blood_values_init, by = "donor") 

blood_data_summary <- donor_demo %>%
  inner_join(blood_data_summary, by = "donor") %>%
  mutate(all_study_FB_donation_count = NonFinnDonorDonationCount_FB + YesFinnDonorDonationCount_FB) 


#FINRISK97 
## Assign all FinRisk97 paprticipants to FINRISK97 cohort
fr1997$Cohort <- c("FINRISK97")

## Change term for sex from binary 1,2 to men, women
fr1997$SUKUP <- gsub("1", "Men", fr1997$SUKUP)
fr1997$SUKUP <- gsub("2", "Women", fr1997$SUKUP)

## Get sex for each participant
fr1997_summary <- fr1997 %>%
  filter(!is.na(FERRITIN)) %>%
  group_by(RELEASE_ID) %>%
  summarise(Sex=first(SUKUP))

## remove n/a:s for ferritin 
fr1997_remove_na <- fr1997 %>%
  filter(!is.na(FERRITIN)) %>%
  group_by(RELEASE_ID) %>%
  dplyr::select(IKA, FERRITIN, HGB, CRP, RELEASE_ID, K129, BMI, KY100_22, Q69, KY163, GRAVID, Cohort) %>%
  ungroup()

fr1997_summary <- fr1997_summary %>%
  full_join(fr1997_remove_na, by = "RELEASE_ID")


#HEALTH2000
## Assign all Health2000 participants to HEALTH2000 cohort

h2000$Cohort <- c("HEALTH00")

## Change term for sex from binary 1,2 to men, women
h2000$SP2 <- gsub("1", "Men", h2000$SP2)
h2000$SP2 <- gsub("2", "Women", h2000$SP2)  

## Get sex for each participant
h2000_summary <- h2000 %>%
  filter(!is.na(FERRITIINI)) %>%
  group_by(RELEASE_ID) %>%
  summarise(Sex=first(SP2))

## remove n/a:s for ferritin and Hb
  h2000_remove_na <- h2000 %>%
    filter(!is.na(FERRITIINI) & !is.na(B_Hb)) %>%
    group_by(RELEASE_ID) %>%
    dplyr::select(IKA2, FERRITIINI, B_Hb, CRP, RELEASE_ID, BD03, BMII_BMI.x, FB05, synnytys, BD07, Cohort) %>%
    ungroup()
  
  h2000_summary <- h2000_summary %>%
    full_join(h2000_remove_na, by = "RELEASE_ID")

  
# Rename variables to maintain cohesion between the cohorts
  
## Participant ID 
names(blood_data_summary)[names(blood_data_summary) == "donor"] <- "ID"
names(fr1997_summary)[names(fr1997_summary) == "RELEASE_ID"] <- "ID"
names(h2000_summary)[names(h2000_summary) == "RELEASE_ID"] <- "ID"

## Age
names(fr1997_summary)[names(fr1997_summary) == "IKA"] <- "Age"
names(h2000_summary)[names(h2000_summary) == "IKA2"] <- "Age"

## Hemoglobin
names(blood_data_summary)[names(blood_data_summary) == "Hb_v"] <- "Hb"
names(fr1997_summary)[names(fr1997_summary) == "HGB"] <- "Hb"
names(h2000_summary)[names(h2000_summary) == "B_Hb"] <- "Hb"

## Ferritin
names(fr1997_summary)[names(fr1997_summary) == "FERRITIN"] <- "Ferritin"
names(h2000_summary)[names(h2000_summary) == "FERRITIINI"] <- "Ferritin"

## Menstruation
names(blood_data_summary)[names(blood_data_summary) == "QR79"] <- "Menstruation"
names(fr1997_summary)[names(fr1997_summary) == "K129"] <- "Menstruation"
names(h2000_summary)[names(h2000_summary) == "BD03"] <- "Menstruation"

## BMI
names(h2000_summary)[names(h2000_summary) == "BMII_BMI.x"] <- "BMI"

## Smoking
names(blood_data_summary)[names(blood_data_summary) == "QR54"] <- "Smoking"
names(fr1997_summary)[names(fr1997_summary) == "Q69"] <- "Smoking"
names(h2000_summary)[names(h2000_summary) == "FB05"] <- "Smoking"

## Red meat
names(blood_data_summary)[names(blood_data_summary) == "QR40"] <- "RedMeat"
names(fr1997_summary)[names(fr1997_summary) == "KY100_22"] <- "RedMeat"

## Iron supplements
names(blood_data_summary)[names(blood_data_summary) == "vita_iron"] <- "HistoryOfIronSupplements"
names(blood_data_summary)[names(blood_data_summary) == "iron_supp"] <- "GivenIronSupplements"
names(blood_data_summary)[names(blood_data_summary) == "iron_comp"] <- "IronComplience"
names(blood_data_summary)[names(blood_data_summary) == "iron_comp_c"] <- "IronComplienceNumeric"

## History of childbirth
names(blood_data_summary)[names(blood_data_summary) == "QR83"] <- "PreviousChildbirth"
names(fr1997_summary)[names(fr1997_summary) == "KY163"] <- "PreviousChildbirth"
names(h2000_summary)[names(h2000_summary) == "synnytys"] <- "PreviousChildbirth"

## Current pregnancy
names(fr1997_summary)[names(fr1997_summary) == "GRAVID"] <- "CurrentPregnancy"
names(h2000_summary)[names(h2000_summary) == "BD07"] <- "CurrentPregnancy"
```

## Demographic group specification and assignment

In this section we assign women to period/no period groups

There are `r cohort_summary_name %>% filter(Sex == "Women" & is.na(Menstruation)) %>%  nrow()` women with no  answer to the question regarding their menstrual status. Of these, `r cohort_summary_name %>% filter(Sex == "Women" & is.na(Menstruation) & age > 55) %>%  nrow()` are older than 55. We impute these to no period so they can be included in the postmenopausal women. 

FinDonor: Do you still have regular periods? (1=reg, 2=irreg, 3=no)
FinRisk97: Do you still menstruate? (1=reg, 2=irreg, 3=no)
Health2000: BD03 = do you have periods nowadays? (1=reg, 2=irreg, 3=no)

```{r}
# FINDONOR
older_findonor_donors <- filter(blood_data_summary, 
                       Sex == "Women" & is.na(Menstruation) & Age > 55)$ID

blood_data_summary <- blood_data_summary %>% 
  mutate(Menstruation = ifelse(ID %in% older_findonor_donors, "no_period", as.character(Menstruation))) 

#FINRISK97
older_fr1997 <- filter(fr1997_summary, 
                       Sex == "Women" & is.na(Menstruation) & Age > 55)$ID

fr1997_summary <- fr1997_summary %>% 
  mutate(Menstruation = ifelse(ID %in% older_fr1997, "no_period", as.character(Menstruation)))

#HEALTH2000
older_h2000 <- filter(h2000_summary,
                      Sex == "Women" & is.na(Menstruation) & Age > 55)$ID

h2000_summary <- h2000_summary %>% 
  mutate(Menstruation = ifelse(ID %in% older_h2000, "no_period", as.character(Menstruation)))
```


Female study participants with no menstruation response after imputation are removed:  `r cohort_name %>% filter(sex == "Women" & is.na(Menstruation)) %>% nrow()` donors.)


```{r}
#FINDONOR
findonor_nb_women_no_menstruation_response <- blood_data_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>% nrow()

## remove female donors with no menstruation response
blood_data_summary_final <- blood_data_summary %>% 
  mutate(mens_ok_blood = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & !is.na(Menstruation)  ~ "Women",
    TRUE ~ "NA")) %>% 
  filter(mens_ok_blood != "NA") 

## We define the women's groups: 
findonor_n_women_removed <- blood_data_summary_final %>% 
  mutate(Group = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & (Menstruation == "no_period" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "no_period" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "irregular_period" | Menstruation == "regular_period") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>% 
  group_by(Group) %>% 
  filter(Group ==  "Women_pre_menop_no_mens") %>% 
  nrow()


#FINRISK97
finrisk_nb_women_no_menstruation_response <- fr1997_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>% nrow()


## remove female FinRisk97 participants with no menstruation response
fr1997_summary_final <- fr1997_summary %>% 
  mutate(mens_ok_fr1997 = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & !is.na(Menstruation)  ~ "Women",
    TRUE ~ "NA")) %>% 
  filter(mens_ok_fr1997 != "NA") 

## We define the women's groups:
fr1997_n_women_removed <- fr1997_summary_final %>%
  mutate(Group = case_when(
    Sex == "Men" ~ "Men",
    Sex == "Women" & (Menstruation == "3" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "3" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "1" | Menstruation == "2") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>%
  group_by(Group) %>%
  filter(Group ==  "Women_pre_menop_no_mens") %>%
  nrow()

#HEALTH2000
h2000_nb_women_no_menstruation_response <- h2000_summary %>% filter(Sex == "Women" & is.na(Menstruation)) %>% nrow()


## remove female Health2000 participants with no menstruation response
h2000_summary_final <- h2000_summary %>% 
  mutate(mens_ok_h2000 = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & !is.na(Menstruation)  ~ "Women",
    TRUE ~ "NA")) %>% 
  filter(mens_ok_h2000 != "NA") 

## We define the women's groups:
h2000_n_women_removed <- h2000_summary_final %>%
  mutate(Group = case_when(
    Sex == "Men" ~ "Men",
    Sex == "Women" & (Menstruation == "3" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "3" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "1" | Menstruation == "2") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>%
  group_by(Group) %>%
  filter(Group ==  "Women_pre_menop_no_mens") %>%
  nrow()

table(h2000$BD03, h2000$SP2)
table(h2000$MENOP, h2000$SP2)
table(h2000$MENOP, h2000$BD03)

#BD03 = reg, irreg, no
#MENOP = post, peri, pre 
```

We now define the following groups:

+ premenopausal: regular or irregular menstruation reported
+ postmenopausal: no menstruation reported and age equal to > 45
+ donors younger than 45 and no reported menstruation are excluded (`r n_women_removed` donors)



```{r}
#FINDONOR:
## We define the women's groups: 
blood_data_summary_final <- blood_data_summary_final %>% 
  mutate(Group = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & (Menstruation == "no_period" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "no_period" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "irregular_period" | Menstruation == "regular_period") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>% 
  filter(Group != "Women_pre_menop_no_mens" & Group != "NA") %>% 
  droplevels() %>% 
  mutate(Group = ordered(Group, levels =  c("Pre_menopause_women", "Post_menopause_women", "Men"))) 


#FINRISK97:
## We define the women's groups: 
fr1997_summary_final <- fr1997_summary_final %>% 
  mutate(Group = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & (Menstruation == "3" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "3" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "1" | Menstruation == "2") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>% 
  filter(Group != "Women_pre_menop_no_mens" & Group != "NA") %>% 
  droplevels() %>% 
  mutate(Group = ordered(Group, levels =  c("Pre_menopause_women", "Post_menopause_women", "Men"))) 

#HEALTH2000:
## We define the women's groups: 
h2000_summary_final <- h2000_summary_final %>% 
  mutate(Group = case_when(
    Sex == "Men" ~ "Men", 
    Sex == "Women" & (Menstruation == "3" & Age > 45) ~ "Post_menopause_women",
    Sex == "Women" & (Menstruation == "3" & Age <= 45) ~ "Women_pre_menop_no_mens",
    Sex == "Women" & (Menstruation == "1" | Menstruation == "2") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>% 
  filter(Group != "Women_pre_menop_no_mens" & Group != "NA") %>% 
  droplevels() %>% 
  mutate(Group = ordered(Group, levels =  c("Pre_menopause_women", "Post_menopause_women", "Men"))) 


```

In order to later merge the three cohorts into one table, we need to make colums for the variables the different cohorts don't take into consideration. 

```{r}
# We make new columns for donation data for the general population cohorts
fr1997_summary_final$TwoYearsFromStartCount_FB <- 0
h2000_summary_final$TwoYearsFromStartCount_FB <- 0

# We then  assign a randomized donation interval for 3-6 years for the general population cohorts (at this point ferritin levels out, see Lobier et al.)
fr1997_summary_final <- fr1997_summary_final %>% mutate(DaysToPreviousFB = round(runif(nrow(fr1997_summary_final), min = 3*365, max = 6*365) ,0))
h2000_summary_final <- h2000_summary_final %>% mutate(DaysToPreviousFB = round(runif(nrow(h2000_summary_final), min = 3*365, max = 6*365) ,0))

# Add columns for the variables that are missing from the cohorts (CurrentPregnancy for FinDonor, RedMeat for Health2000 and all iron supplement variables for both general population cohorts) and add values to them (no for pregnancy in FinDonor, n/a for the rest)
blood_data_summary_final$CurrentPregnancy <- 0
fr1997_summary_final$HistoryOfIronSupplements <- NA
fr1997_summary_final$GivenIronSupplements <- NA
fr1997_summary_final$IronComplience <- NA
fr1997_summary_final$IronComplienceNumeric <- NA
h2000_summary_final$HistoryOfIronSupplements <- NA
h2000_summary_final$GivenIronSupplements <- NA
h2000_summary_final$IronComplience <- NA
h2000_summary_final$IronComplienceNumeric <- NA
h2000_summary_final$RedMeat <- NA
```

## Remove cohort participants with missing data

### Donation history

We remove `r blood_data_summary_final %>% filter(is.na(DaysToPreviousFB)) %>% nrow()`, that is donors who have not donated previously (they are missing the nb of days since last FB donation variable). As there is no information on blood donation in the general population cohorts this is not done for them. 

```{r}
new_donors_data <- blood_data_summary_final %>% 
  filter(is.na(DaysToPreviousFB)) %>% 
    mutate(Group = case_when(Group == "Pre_menopause_women" ~ "Pre-menopausal women",
                   Group == "Post_menopause_women" ~ "Post-menopausal women",
                   Group == "Men" ~ "Men" ,
                   TRUE~ "NA"),
         Group = ordered(Group)) 

blood_data_summary_final <- blood_data_summary_final %>% 
  drop_na(DaysToPreviousFB)
```

### BMI

We remove `r cohort_name %>% filter(is.na(BMI)) %>% nrow()` participants for whom we do not have the BMI data:

```{r}
#### FinDonor
nb_removed_blood <- findonor_nb_women_no_menstruation_response +
                            blood_data_summary_final %>% filter(is.na(BMI)) %>% nrow() 

blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!is.na(BMI))

#### FinRisk97
nb_removed_fr97 <- finrisk_nb_women_no_menstruation_response +
                            fr1997_summary_final %>% filter(is.na(BMI)) %>% nrow() 

fr1997_summary_final <- fr1997_summary_final %>% 
  filter(!is.na(BMI))

#### Health2000
nb_removed_h2000 <- h2000_nb_women_no_menstruation_response +
                            h2000_summary_final %>% filter(is.na(BMI)) %>% nrow() 

h2000_summary_final <- h2000_summary_final %>% 
  filter(!is.na(BMI))
```

### Smoking

We remove `r cohort_name %>% filter(is.na(BMI)) %>% nrow()` participants for whom we do not have the BMI data:
  * FinDonor: 1      `r blood_data_summary_final %>% filter(is.na(Smoking)) %>% nrow()`
  * FinRisk97: 3086  `r fr1997_summary_final  %>% filter(is.na(Smoking)) %>% nrow()`
  * Health2000: 1560 `r h2000_summary_final %>% filter(is.na(Smoking)) %>% nrow()` 
  
As removing n/a:s for smoking removes >4600 general population participants as opposed to only one in the blood donor cohort, we suspect there being something wrong with the original data. At this point we impute the n/a:s for the general population cohorts, while we wait for THL to explain the missing data. 

```{r}
# FinDonor
nb_removed_blood <- nb_removed_blood +
                            blood_data_summary_final %>% filter(is.na(Smoking)) %>% nrow() 

blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!is.na(Smoking))

# FinRisk97
fr1997_summary_final$Smoking[is.na(fr1997_summary_final$Smoking)]  <- 3

# nb_removed_fr97 <- nb_removed_fr97 +
#                             fr1997_summary_final %>% filter(is.na(Smoking)) %>% nrow() 
# 
# fr1997_summary_final <- fr1997_summary_final %>% 
#   filter(!is.na(Smoking))

# Health2000
h2000_summary_final$Smoking[is.na(h2000_summary_final$Smoking)]  <- 3

# nb_removed_h2000 <- nb_removed_h2000 +
#                             h2000_summary_final %>% filter(is.na(Smoking)) %>% nrow() 
# 
# h2000_summary_final <- h2000_summary_final %>% 
#   filter(!is.na(Smoking))

```

### Previous childbirth

We remove `r cohort_name %>% filter((group != "Men" & is.na(QR83)))%>% nrow()` female donors who did not answer the question about childbirth: 

Women who answer no to the of previous childbirth = nulliparous women

```{r}
#### FinDonor
nb_removed_blood <- nb_removed_blood +
                            blood_data_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth))) %>% nrow 

blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!(Group != "Men" & is.na(PreviousChildbirth)))

#### FinRisk97
nb_removed_fr97 <- nb_removed_fr97 +
                            fr1997_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth))) %>% nrow 

fr1997_summary_final <- fr1997_summary_final %>% 
  filter(!(Group != "Men" & is.na(PreviousChildbirth)))

#### Health2000
nb_removed_h2000 <- nb_removed_h2000 +
                            h2000_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth))) %>% nrow 

h2000_summary_final <- h2000_summary_final %>% 
  filter(!(Group != "Men" & is.na(PreviousChildbirth)))

```

### Current pregnancy

Pregnant women are not allowed to donate blood in NL and FIN, and one of the exclusion criteria for the NL PREVEND cohort was current pregnancy. Therefore women who are currently pregnant will need to be removed from the FIN general population cohorts. We will now remove `r cohort_name %>% filter((group != "Men" & is.na(QR83)))%>% nrow()` females who did not answer the pregnancy question: 

```{r}
#### FinRisk97
nb_removed_fr97 <- nb_removed_fr97 +
                            fr1997_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth))) %>% nrow 

fr1997_summary_final <- fr1997_summary_final %>% 
  filter(!(Group != "Men" & is.na(PreviousChildbirth)))

#### Health2000
nb_removed_h2000 <- nb_removed_h2000 +
                            h2000_summary_final %>% filter((Group != "Men" & is.na(PreviousChildbirth))) %>% nrow 

h2000_summary_final <- h2000_summary_final %>% 
  filter(!(Group != "Men" & is.na(PreviousChildbirth)))
```


### Iron supplementation 

We remove `r blood_data_summary_final %>% mutate(IronComplienceNumeric = ifelse(GivenIronSupplements == FALSE, 0, IronComplienceNumeric )) %>%  filter(is.na(IronComplienceNumeric)) %>% nrow()` donors that did not answer the two questions (did they receive iron supplementation and complience to supplementation)

For the modelling, we impute a 0 (no supplementation) to iron_comp_c when the donor reports not being offered iron supplementation.

This is not done for the general population cohorts, as they were not asked about iron supplementation and as such all observation are input as n/a. 

```{r}
nb_removed_blood <- nb_removed_blood +
                            blood_data_summary_final %>% 
                              mutate(IronComplienceNumeric = ifelse(GivenIronSupplements == FALSE, 0, IronComplienceNumeric)) %>% 
                              filter(is.na(IronComplienceNumeric)) %>% 
                              nrow() 

blood_data_summary_final <- blood_data_summary_final %>% 
  mutate(IronComplienceNumeric = ifelse(GivenIronSupplements == FALSE, 0, IronComplienceNumeric )) %>% 
  filter(!is.na(IronComplienceNumeric))
```

### Red meat intake 
      
Health2000 participants were not asked about red meat intake, so we only remove FinDonor and FinRisk97 participants who did not answer the red meat question. 

```{r message=FALSE, warning=FALSE}
# FinDonor
donors_to_remove <- blood_data_summary_final %>%
  dplyr::select(ID, Group, RedMeat) %>%
  gather(key= question, value = answer, -ID,-Group) %>%
  filter(is.na(answer)) %>%
  dplyr::select(Group, ID) %>%
  distinct(ID, Group)

blood_data_summary_final <- blood_data_summary_final %>%
  filter(!ID %in% donors_to_remove$Donor)

nb_removed_blood <- nb_removed_blood +
                            donors_to_remove %>% nrow()

# FinRisk97
finrisk97_to_remove <- fr1997_summary_final %>%
  dplyr::select(ID, Group, RedMeat) %>%
  gather(key= question, value = answer, -ID,-Group) %>%
  filter(is.na(answer)) %>%
  dplyr::select(Group, ID) %>%
  distinct(ID, Group)

fr1997_summary_final <- fr1997_summary_final %>%
  filter(!ID %in% finrisk97_to_remove$Donor)

nb_removed_fr97 <- nb_removed_fr97 +
                            finrisk97_to_remove %>% nrow()


```

### Total number of donors removed for missing questionnaire answers

The total number of donors removed because of missing questionnaire data is:
`r nb_removed_blood`
`r nb_removed_fr97`
`r nb_removed_h2000`

## Remove cohort participants who are currently pregnant

As current pregnancy was an exclusion criteria for the NL general population cohort and pregnant women are not allowed to donate blood, we will remove women who are currently pregnant from the FIN general population cohorts. 

FinRisk97: 1=no, 2=yes
Health2000: 0=no, 1=yes

We will also need to recode the FinRisk97 cohort answer of 1=no to 0=no for cohesion.  

```{r}
# FinRisk97
nb_removed_pregnant_fr1997 <- fr1997_summary_final %>% 
   filter(CurrentPregnancy == 2) %>% 
  nrow()

fr1997_summary_final <- fr1997_summary_final %>% 
   filter(!(Group != "Men" & CurrentPregnancy == 2))

fr1997_summary_final$CurrentPregnancy[fr1997_summary_final$CurrentPregnancy == "1"] <- 0

#Health2000
nb_removed_pregnant_h2000 <- h2000_summary_final %>% 
   filter(!(CurrentPregnancy == 0)) %>% 
  nrow()

h2000_summary_final <- h2000_summary_final %>% 
   filter(!(Group != "Men" & CurrentPregnancy == 1))

```

## Remove donors with extreme physiological measures

As decided previously, we remove data according the following criteria:

* BMI > 50
* CRP > 30
* Ferritin > 400 

This amounts to `r cohort_name %>% filter(BMI >= 50 | CRP >= 30 | Ferritin >= 400) %>% nrow()` donors that are removed. 


```{r}
# FinDonor
blood_data_summary_final <- blood_data_summary_final %>% 
   filter(BMI < 50 & CRP < 30 & Ferritin < 400)

# FinRisk97
fr1997_summary_final <- fr1997_summary_final %>% 
   filter(BMI < 50 & CRP < 30 & Ferritin < 400)

#Health2000
h2000_summary_final <- h2000_summary_final %>% 
   filter(BMI < 50 & CRP < 30 & Ferritin < 400)

```

## Final group N for the three cohorts

```{r}
# FinDonor
blood_data_summary_final %>% 
   mutate(Group = dplyr::recode(Group, Pre_menopause_women = "Pre-menopausal women",
         Post_menopause_women = "Post-menopausal women")) %>% 
  group_by(Group) %>% 
  summarise ( N = n()) %>% 
  kable() 

# FinRisk97
fr1997_summary_final %>% 
   mutate(Group = dplyr::recode(Group, Pre_menopause_women = "Pre-menopausal women",
         Post_menopause_women = "Post-menopausal women")) %>% 
  group_by(Group) %>% 
  summarise ( N = n()) %>% 
  kable() 

# Health2000
h2000_summary_final %>% 
   mutate(Group = dplyr::recode(Group, Pre_menopause_women = "Pre-menopausal women",
         Post_menopause_women = "Post-menopausal women")) %>% 
  group_by(Group) %>% 
  summarise ( N = n()) %>% 
  kable() 

 
```

Next we recode red meat consumption into a linear variable. We use a linear scale from 1 to 4 for and mapped the different responses to correspond to the four options for the FinDonor and FinRisk97 cohorts. Data on red meat consumption was not avaliable in the Health2000 cohort.  

* red meat:
    * FinDonor
        * "never" ~ 1,
              # never
        * "less_than_once_weekly" ~ 2,
              # less than once a week
        * "1.3_week" or "4.6_week"~ 3,
              # 1-3 times a week
              # 4-6 times a week
        * "daily" or "several_daily" ~ 4,
              # daily
              # several times a day
    * FinRisk97
        * "1" ~ 1, 
              # 1 = more seldom than once a month or not at all
        * "2" ~ 2,
              # 2 = once or twice a month
        * "3" or "4" or "5" ~ 3,
              # 3 = once a week
              # 4 = twice a week or more often
              # 5 = almost daily
        * "5" or "several_daily" ~ 4,
              # 6 = daily
              # 7 = several times a day
              
This will create a new variable, RedMeat_n. As we don't have red meat data in the Health2000 cohort, we need to make a column, name it RedMeat_n and impute the observations as n/a.

```{r}
# FinDonor
blood_data_summary_final <- blood_data_summary_final %>% 
  mutate(RedMeat_n = case_when(
    RedMeat == "never" ~ 1,
    RedMeat == "less_than_once_weekly" ~ 2,
    RedMeat %in% c("1.3_week" , "4.6_week") ~ 3,
    TRUE   ~ 4 )) 

# FinRisk97
fr1997_summary_final <- fr1997_summary_final %>% 
  mutate(RedMeat_n = case_when(
    RedMeat == "1" ~ 1,
    RedMeat == "2" ~ 2,
    RedMeat %in% c("3" , "4", "5") ~ 3,
    TRUE   ~ 4 )) # "6" | "7"
  
# Health2000
h2000_summary_final$RedMeat_n <- NA


```

Due to difference in the CRP measurements (hs-CRP for the general population cohorts, CRP for the blood donor cohort) we decided to impute CRP <3 mg/l to 2.9 mg/l for the FinDonor cohort.  

```{r}
# blood_data_summary_final_CRP <- blood_data_summary_final %>% 
#   mutate(CRP = case_when(
#     CRP == "2.9" ~ 1,
#     TRUE ~ as.character())) 
```

Previous childbirth is recoded 
* FinDonor: have you given birth? no=no, yes=yes
* FinRisk97: how many children have you given birth to? 1=none, 2=one, 3=two, 4=three or more
* Health2000: have you given birth? 0=no, 1=yes

Nulliparous women are coded as 0 (no) and reporting childbirth or giving birth to any number of children to 1 (yes). 

```{r}
# FinDonor
blood_data_summary_final <- blood_data_summary_final %>% 
  mutate(PreviousChildbirth = case_when(
    PreviousChildbirth == "no" ~ 0,
    PreviousChildbirth == "yes" ~ 1)) 

# FinRisk97
fr1997_summary_final <- fr1997_summary_final %>% 
  mutate(PreviousChildbirth = case_when(
    PreviousChildbirth == "1" ~ 0,
    PreviousChildbirth %in% c("2" , "3", "4") ~ 1)) 


```


## Merging cohorts into one table 

For the code to work all three cohorts need to be merged into one single table. Before this can be done variables the variables selected need to have the same name.  

```{r}
### mutate smoking and pregnancy 
blood_data_summary_final <- blood_data_summary_final %>% 
  dplyr::select(ID, Group, Sex, Age, TwoYearsFromStartCount_FB, DaysToPreviousFB,
                Ferritin, Hb, BMI, Menstruation, Smoking, RedMeat_n, HistoryOfIronSupplements, IronComplienceNumeric, PreviousChildbirth,
                CRP, Cohort) %>% 
    mutate(Smoking = ifelse(Smoking == "daily", "yes", "no"),
         Smoking = factor(Smoking, levels = c( "no",  "yes")))

#HistoryOfIronSupplements!!!!!!!!

### mutate smoking and pregnancy 
fr1997_summary_final <- fr1997_summary_final %>% 
  dplyr::select(ID, Group, Sex, Age, TwoYearsFromStartCount_FB, DaysToPreviousFB,
                Ferritin, Hb, BMI, Menstruation, Smoking, RedMeat_n, HistoryOfIronSupplements, IronComplienceNumeric, PreviousChildbirth,
                CRP, Cohort) %>% 
       mutate(Smoking = ifelse(Smoking == "1", "yes", "no"),
       Smoking = factor(Smoking, levels = c( "no",  "yes")))

### mutate smoking and pregnancy
h2000_summary_final <- h2000_summary_final %>% 
  dplyr::select(ID, Group, Sex, Age, TwoYearsFromStartCount_FB, DaysToPreviousFB,
                Ferritin, Hb, BMI, Menstruation, Smoking, RedMeat_n, HistoryOfIronSupplements, IronComplienceNumeric, PreviousChildbirth,
                CRP, Cohort) %>% 
         mutate(Smoking = ifelse(Smoking == "1", "yes", "no"),
         Smoking = factor(Smoking, levels = c( "no",  "yes")))
         

# merge the cohorts

summary_all_cohorts <- bind_rows(blood_data_summary_final, fr1997_summary_final, h2000_summary_final)

```

## Final group N for the summarized table

```{r}
summary_all_cohorts %>% 
   mutate(Group = dplyr::recode(Group, Pre_menopause_women = "Pre-menopausal women",
         Post_menopause_women = "Post-menopausal women")) %>% 
  group_by(Group) %>% 
  summarise ( N = n()) %>% 
  kable() 
 
```

# Results - Regression analyses

## Regressions on sTfR and ferritin  

We analyzed the data with multiple linear regression. We first ran Ordinary Least Square (OLS) regressions and their diagnostics. If the diagnostics identified problematic observations (e.g., outliers, influential observations), we also ran robust regression regressions [@yu2017robust]. Robust regression is based on less assumptions than OLS regression and is less sensitive to influential observations. It  gives more accurate estimates of regression coefficients in the presence of problematic observations. Using robust regression leads to a more principled approach to regression analysis and avoids potential confounds introduced by selective removal of influential observations. As the rlm command from the MASS package does not compute directly the *t* and *p* values, we used the f.robftest function from the sfsmisc package. Practical examples of using robust regression can be found in the literature [@davison2017platelet].  


In addition, we used relative importance analysis to estimate the average proportion of variance in the outcome variable explained by each co-variate [@Groemping2006]. Relative iportance was computed using the pmvd method which assigns 0 to the relative importance of a regressor if it is non significant in the complete model. .

The regressors were entered as follows in the regression models for sTfR and ferritin:

* Continuous variables were entered as continuous predictors
    * age, number of donations in last two years, nb of days since last donation, CRP, BMI
  
* Binary variables were entered as categorical (but coded as numerical for relative importance):
    * Smoking, pregnancy
  
* Ordinal variables were coded on a linear scale and entered as continuous [(Reference)](https://www3.nd.edu/~rwilliam/stats3/OrdinalIndependent.pdf)
    * Iron supplementation 
    * All dietary intake
  
We compute both coefficients and standardized coefficients (with all dependent and independent variables standardized) for robust regression and only coefficients for OLS regression. We present robust coefficients and their bootstrapped BCa 95% confidence intervals in Table 1 (ferritin) and  Table 2 (sTfR), robust standardized coefficients, their bootstrapped distribution and BCa 95% confidence intervals in Figure 2.A and non-standardized OLS regression coefficients in Supplementary Table 1 (Ferritin) and 2 (sTfR). 

## Pre-processing

### Rename and transform

```{r}
regression_cohorts <- summary_all_cohorts %>% 
    rename(donation_count = TwoYearsFromStartCount_FB,
            last_donation = DaysToPreviousFB,
            iron_complience = IronComplienceNumeric) %>% 
     mutate(Age = Age / 5, 
            donation_count_2 = donation_count^2,
            log_ferritin = log(Ferritin),
            log_last_donation = log(last_donation)/log(2),
            log_CRP = log(CRP)) 

```

Data transformations:

* Log transformed variables: 
    * Ferritin
    * sTfR
    * CRP 
  
* Age is divided by 5 to simplify coefficient interpretation
* Number of days before donation is transformed as $log\_last\_donation = log(last\_donation)/log(2)$ to help with the interpretation of coefficients. 
    * An increase in one of the transformed variable is equivalent to a doubling of the number of days since last donation.

* Standardization:
    * Standardized coefficients: 
        * All dependent and independent variables were standardized
      
    * Coefficients:
        * All dependent variables entered as continuous varaibles are centered but not scaled. 

## Premenopausal women
### Data pre-processing

We center all variables entered as continuous.

```{r}
# Center variables

# test_data_pre_women <- regression_cohorts %>% 
#   filter(Group == "Pre_menopause_women") %>% 
#   dplyr::select(ID, RedMeat_n) %>% 
#   gather(key = key, value = value, -ID) %>% 
#   group_by(key) %>% 
#   mutate(value = scale(value, scale = FALSE)[,1]) %>% 
#   spread(key = key, value = value)
# ei ehkä tarvita tätä kohtaa 

test_data_pre_women <- regression_cohorts %>% 
  select(ID, #luokkamuuttujat mitä ei transformoida)

test_data_pre_women <- regression_cohorts %>% 
  filter(Group == "Pre_menopause_women") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) %>% 
  full_join(test_data_pre_women, by = "ID")  %>% 
  dplyr::select(-ID)

```


### Mutiple regression - ferritin as outcome

#### Correlograms

```{r}

ggpairs(test_data_pre_women, 
        columns = c("log_ferritin", "Age", "log_CRP",  "donation_count", "donation_count_2",
                    "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

```{r}

ggpairs(test_data_pre_women, 
        columns = c("log_ferritin", "PreviousChildbirth",  "RedMeat_n", "Smoking", "GivenIronSupplements", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```


# Correlogram for age

```{r}

ggpairs(test_data_pre_women, 
        columns = c("log_ferritin", "Age"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram log CRP 

```{r}

ggpairs(test_data_pre_women, 
        columns = c("log_ferritin", "log_CRP"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram donation count 

```{r}

ggpairs(test_data_pre_women, 
        columns = c("log_ferritin", "donation_count"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram donation count 2 

```{r}

ggpairs(test_data_pre_women, 
        columns = c("log_ferr", "donation_count_2"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram log last donation

```{r}

ggpairs(test_data_pre_women, 
        columns = c("log_ferritin", "log_last_donation"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram pregnancy (QR83=have you given birth? yes/no)

```{r}

ggpairs(test_data_pre_women, 
        columns = c("log_ferritin", "PreviousChildbirth"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram red meat

```{r}

ggpairs(test_data_pre_women, 
        columns = c("log_ferritin", "RedMeat_n"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram smoking 

```{r}

ggpairs(test_data_pre_women, 
        columns = c("log_ferritin", "Smoking"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram iron supplements 

```{r}

ggpairs(test_data_pre_women, 
        columns = c("log_ferritin", "GivenIronSupplements"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

# correlogram BMI

```{r}

ggpairs(test_data_pre_women, 
        columns = c("log_ferritin", "BMI"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```


## Postmenopausal women
### Data pre-processing



## Men
### Data pre-processing



