---
title: "Low ferritin model"
author: "Sofie Ekroos"
date: "19.7.2021"
output: github_document
bibliography: index.bib
---

Based on the original code written by Muriel Lobier and Mikko Arvas (https://github.com/FRCBS/iron_levels_of_blood_donors/blob/master/src/index.Rmd and https://github.com/FRCBS/changes_in_donor_health/blob/master/src/explorative_analysis.Rmd)

# Summary

This document includes all codes necessary to run the analysis of and produce the figures for the three Finnish Cohorts (FinDonor, FINRISK97, Health2000). The code is in two parts. Part 1 allows the user to identify co-variants of iron deficiency using logistic regression, part 2 to identify co-variants of ferritin concentration using multiple linear regression.

```{r setup, , message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(safeBinaryRegression)
library(tidymodels)
library(ggplot2)
library(car)
library(boot)
```



PART 1: LOGISTIC REGRESSION

# Data loading and preparation 

Load data

```{r}
load("../data/ID_data_regression_cohorts.rdata")
all_cohorts <- regression_cohorts
```

## Percentage of ID, all cohorts

Here we check the prevalence of iron deficiency in the cohorts 

```{r}
# Percentage of ID, all cohorts 
all_cohorts %>%
  dplyr::select(Group, iron_deficiency) %>% 
  group_by(Group, iron_deficiency) %>% 
  count() %>% 
  group_by(Group) %>% 
  summarise(
    ID = iron_deficiency,
    n=n,
    Percentage = round(n/sum(n),2),
    .groups='keep'
  )

```

```{r}
# Percentage of ID, FinDonor  
all_cohorts %>%
  filter(Cohort == "FINDONOR") %>% 
  dplyr::select(Group, iron_deficiency) %>% 
  group_by(Group, iron_deficiency) %>% 
  count() %>% 
  group_by(Group) %>% 
  summarise(
    ID = iron_deficiency,
    n=n,
    Percentage = round(n/sum(n),2),
    .groups='keep'
  )
```

```{r}
# percentage of ID, FinRisk97
all_cohorts %>%
  filter(Cohort == "FINRISK97") %>% 
  dplyr::select(Group, iron_deficiency) %>% 
  group_by(Group, iron_deficiency) %>% 
  count() %>% 
  group_by(Group) %>% 
  summarise(
    ID = iron_deficiency,
    n=n,
    Percentage = round(n/sum(n),2),
    .groups='keep'
  )
```

```{r}
# percentage of ID, Health2000
all_cohorts %>%
  filter(Cohort == "HEALTH00") %>% 
  dplyr::select(Group, iron_deficiency) %>% 
  group_by(Group, iron_deficiency) %>% 
  count() %>% 
  #group_by(Group) %>% 
  summarise(
    ID = iron_deficiency,
    n=n,
    Percentage = round(n/sum(n),2),
    .groups='keep'
  )
```

Data transformations:

* Log transformed variables: 
    * Ferritin
    * CRP 
  
* Age is divided by 5 to simplify coefficient interpretation
* Number of days before donation is transformed as $log\_last\_donation = log(last\_donation)/log(2)$ to help with the interpretation of coefficients. 
    * An increase in one of the transformed variable is equivalent to a doubling of the number of days since last donation.

* Standardization:
    * Standardized coefficients: 
        * All dependent and independent variables were standardized
      
    * Coefficients:
        * All dependent variables entered as continuous varaibles are centered but not scaled. 

We center all variables entered as continuous.



#######################################################################################################################################
** POISTA TÄMÄ KOHTA ENNEN VERTAILUA NL-KOHORTTIVERTAILUA, HEILTÄ PUUTTUU CRP DATA. SÄILYTÄ JOS SAADAAN CRP DATAA UUDESTA KOHORTISTA ** 

We are missing CRP data from `r all_cohorts %>% filter(is.na(CRP)) %>% nrow()`. These will need to be removed. 

```{r}
all_cohorts <- all_cohorts %>% 
  filter(!is.na(CRP))
```

*** "POISTA TÄMÄ KOHTA" LOPPUU ***
#######################################################################################################################################


## Data transformations

```{r}
# men
## menstruation variable removed 
regression_data_men <- all_cohorts %>% 
  filter(Group == "Men") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, Cohort, iron_deficiency, Smoking, iron_complience, Group) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# premenopausal women
regression_data_pre <- all_cohorts %>% 
  filter(Group == "Pre_menopause_women") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, Menstruation, Cohort, iron_deficiency, Smoking, iron_complience, Group) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# postmenopausal womem 
## menstruation variable removed 
regression_data_post <- all_cohorts %>% 
  filter(Group == "Post_menopause_women") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, Cohort, iron_deficiency, Smoking, iron_complience, Group) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 


```

%% Remove ordering from variables !!!!!!!!!!HUOM !!!!!!! :)

```{r}
all_cohorts <- all_cohorts %>% mutate_if(is.ordered, function(x){factor(x, ordered = FALSE)})
```

HUOM: lisäsin tämän kohdan (poistaa ordered-koodauksen Groupista ja Menstruationista, ilman tätä "relevel" simple model-kohdassa ei toimi)... En keksinyt fiksumpaa tapaa, mutta google kertoo että näin olisi ehkä pitänyt olla alunperinkin (?) ja olin turhaan koodannut sen ordered-kohdan low_ferritin_dataan. Ihan hyvää harjoittelua sinäänsä mutta nyt olen epävarma siitä että lisäänkö simple model-kohdan jälkeen uuden koodipätkän missä ordered:aan nämä uudestaan vai meneekö tämä tällaisenaan glm:iin? :D     

# Simple model 

Trying a simple model for ID and group (premenopausal, postmenopausal, men). Specifying wanting the binomial family of generalized models tells the glm function to do logistic regression as opposed to some other type of generalized linear model. The summary function is used to get details about the logistic regression. The Group-variable is releved to reflect that we want to compare ID rates to men (who have the lowest prevalence).

```{r}
all_cohorts <- within(all_cohorts, Group <- relevel(Group, ref = "Men"))

logistic_simple <- glm(iron_deficiency ~ Group, data = all_cohorts, family = "binomial")
summary(logistic_simple)
```

```{r}
rm(logistic_simple)
```

# Model using more than one variable

Next let's try creating a model that uses all of the variables to predict ID. 

`iron_deficiency ~ .` means that we want to model ID using all of the remaining variables. --> This did not work (why?), added variables one by one.

## Men

```{r}
#men
logistic_men <- glm(iron_deficiency ~ Age + 
                  BMI + 
                  Smoking + 
                  iron_complience + 
                  donation_count_2 + 
                  log_last_donation + 
                  log_CRP +
                  Cohort, 
                data = regression_data_men, 
                family = "binomial", 
                separation = "test")
summary(logistic_men)
```

## Pre

```{r}
# premenopausal women
logistic_pre <- glm(iron_deficiency ~ Age + 
                  BMI + 
                  Smoking + 
                  iron_complience + 
                  donation_count_2 + 
                  log_last_donation + 
                  log_CRP +
                  Cohort + 
                  Menstruation, 
                data = regression_data_pre, family = "binomial", separation = "test")
summary(logistic_pre)
```

## Post

```{r}
# postmenopausal women
logistic_post <- glm(iron_deficiency ~ Age + 
                  BMI + 
                  Smoking + 
                  iron_complience + 
                  donation_count_2 + 
                  log_last_donation + 
                  log_CRP +
                  Cohort, 
                data = regression_data_post, family = "binomial", separation = "test")
summary(logistic_post)
```

%% thoughts on the summary (CHANGE THIS PART TO REFLECT CURRENT MODEL, THIS IS OUT OF DATE):

1) Deviance residuals look good 
2) The coefficients look largly statistically significant, with some variance among the groups  
3) AIC and residual deviance are much smaller for this model than the more simple model in which only gender was used to predict ID, this is to be expected 
4) to calculate FcFadden's Pseudo R² the log-likelihood of the null model can be extracted out of the logistic variable by getting the value for the null deviance and dividing by -2:

```{r}
# men
ll.null_men <- logistic_men$null.deviance/-2
# premenopausal women
ll.null_pre <- logistic_pre$null.deviance/-2
# postmenopausal women
ll.null_post <- logistic_post$null.deviance/-2
```

the log-likelihood for the model can be extracted out out of the logistic variable by getting the value for the residual deviance and dividing by -2:

```{r}
# men
ll.proposed_men <- logistic_men$deviance/-2
# premenopausal women
ll.proposed_pre <- logistic_pre$deviance/-2
# postmenopausal women
ll.proposed_post <- logistic_post$deviance/-2
```

Pseudo R² can then be calculated out of these. These can be interpreted as the overall effect sizes in the three groups

```{r}
# men
(ll.null_men - ll.proposed_men) / ll.null_men
# premenopausal women
(ll.null_pre - ll.proposed_pre) / ll.null_pre
# postmenopausal women
(ll.null_post - ll.proposed_post) / ll.null_post
```

A p-value for R² can then be calculated (Chi-square distribution)

```{r}
# men
1 - pchisq(2*(ll.proposed_men - ll.null_men), df=(length(logistic_men$coefficients)-1))
# premenopausal women
1 - pchisq(2*(ll.proposed_pre - ll.null_pre), df=(length(logistic_pre$coefficients)-1))
# postmenopausal women
1 - pchisq(2*(ll.proposed_post - ll.null_post), df=(length(logistic_post$coefficients)-1))
```

p-values are close to 0, so R² is statistically significant

## Graph

Next draw a graph that shows the predicted probabilities that each individual has ID along with their actual ID status. First create a new data frame that contains the probabilities and actual statuses. 

```{r}
# men
predicted_ID_men <- data.frame(
  probability_of_ID_men = logistic_men$fitted.values,
  ID=regression_data_men$iron_deficiency)

# premenopausal women
predicted_ID_pre <- data.frame(
  probability_of_ID_pre = logistic_pre$fitted.values,
  ID=regression_data_pre$iron_deficiency)

# postmenopausal women
predicted_ID_post <- data.frame(
  probability_of_ID_post = logistic_post$fitted.values,
  ID=regression_data_post$iron_deficiency)
```

Then sort the data frame from low to high probabilities 

```{r}
# men
predicted_ID_men <- predicted_ID_men[
  order(predicted_ID_men$probability_of_ID_men, decreasing = FALSE),]

# premenopausal women 
predicted_ID_pre <- predicted_ID_pre[
  order(predicted_ID_pre$probability_of_ID_pre, decreasing = FALSE),]

# postmenopausal women
predicted_ID_post <- predicted_ID_post[
  order(predicted_ID_post$probability_of_ID_post, decreasing = FALSE),]
```

Next add a new column to the data frame that has the rank of each sample, from low to high probability

```{r}
# men 
predicted_ID_men$rank <- 1:nrow(predicted_ID_men)

# premenopausal women 
predicted_ID_pre$rank <- 1:nrow(predicted_ID_pre)

# postmenopausal women
predicted_ID_post$rank <- 1:nrow(predicted_ID_post)
```

Draw the data and save to pdf 

```{r , message=FALSE}
# men
ggplot(data = predicted_ID_men, aes(x=rank, y=probability_of_ID_men)) + 
  geom_point(aes(color=ID), alpha=1, shape=4, stroke=2) + 
  xlab("Index") + 
  ylab("Predicted probability of iron deficiency, men") 

ggsave("iron_deficiency_probabilities_men.pdf")
```

add below each plot some interpretation of what it might mean, is it good or bad?


```{r , message=FALSE}
# premenopausal women
ggplot(data = predicted_ID_pre, aes(x=rank, y=probability_of_ID_pre)) + 
  geom_point(aes(color=ID), alpha=1, shape=4, stroke=2) + 
  xlab("Index") + 
  ylab("Predicted probability of iron deficiency, premenopausal women") 

ggsave("iron_deficiency_probabilities_pre.pdf")
```

```{r , message=FALSE}
# postmenopausal women 
ggplot(data = predicted_ID_post, aes(x=rank, y=probability_of_ID_post)) + 
  geom_point(aes(color=ID), alpha=1, shape=4, stroke=2) + 
  xlab("Index") + 
  ylab("Predicted probability of iron deficiency, postmenopausal women") 

ggsave("iron_deficiency_probabilities_post.pdf")
```




## Diagnostics for model: 

### Plots

```{r}
# men
#plot(logistic_men,which = 5)
```


```{r}
# premenopausal women 
#plot(logistic_pre,which = 5)
```
```{r}
# postmenopausal women
#plot(logistic_post,which = 5)
```

Most likely these just should not be used https://stats.stackexchange.com/questions/121490/interpretation-of-plot-glm-model .

### Variance inflaction factor 

The more VIF increases, the less reliable regression results are going to be. A VIF > 10 indicates high correlation and is cause for concern.

```{r}
# men 
vif(logistic_men)
```
As the last_donation is simulated for general population cohorts this problem is not suprising. But does it matter?


```{r}
# premenopausal women
vif(logistic_pre)
```
```{r}
# postmenopausal women
vif(logistic_post)
```

Cohort-variables have very high VIFs...

```{r}
# merge the groups back into one table
regression_data <- bind_rows(regression_data_men, 
                             regression_data_post, 
                             regression_data_pre) 
  
```
  
# Bootstraps
https://data.princeton.edu/wws509/r/c3s1
https://stats.stackexchange.com/questions/304833/how-to-calculate-odds-ratio-and-95-confidence-interval-for-logistic-regression
https://www.andrewheiss.com/blog/2016/04/25/convert-logistic-regression-standard-errors-to-odds-ratios-with-r/


```{r}
detach("package:safeBinaryRegression", unload=TRUE)
#safeBinaryRegression messes up booting, this removes the package
```


## Define auxiliary functions for bootstrapping

In this section we build the mechanisms for the bootstrapping itself 

```{r}

# https://www.painblogr.org/2017-10-18-purrring-through-bootstraps.html


get_coefficients <- function(data, boot_ind){
  fit <- glm(iron_deficiency~.,
             #iron_deficiency~., i.e. model iron_deficiency as outcome against everything else in the table. 
              data[boot_ind,],
              family = "binomial")
#  print(names(coef(fit)))
  return(coef(fit))
}

# lasketaan luottamusvälejä
compute_Bca_CI <-function(fit_boot,conf = 0.95){
  cat("compute_Bca_CI with conf: ",conf,"\n")
  Bca_inf = rep(0, length(fit_boot$t0))
  Bca_sup = rep(0, length(fit_boot$t0))
  for (i_regressor in 1:length(fit_boot$t0)){
    CI <- boot.ci(fit_boot, type = "bca", index=i_regressor,conf = conf)
    Bca_inf[i_regressor] <- CI$bca[4]
    Bca_sup[i_regressor] <- CI$bca[5]
  }
  return(tibble(Bca_inf,Bca_sup, regressor = names(fit_boot$t0)) %>% 
           filter(regressor != "(Intercept)"))
}


get_bootstrap <- function(data, nb_boot){
 cat(nrow(data),"get_bootstrap 1\n")
 

    fit_boot <- boot(data, statistic= get_coefficients, R = nb_boot)

 cat(nrow(data),"get_bootstrap 2\n")
  fit_boot_distrib <- as_tibble(fit_boot$t)  
  names(fit_boot_distrib) <- names(fit_boot$t0)

  fit_boot_Bca <- compute_Bca_CI(fit_boot)
  fit_boot_Bca_2 <- compute_Bca_CI(fit_boot,conf=0.90)
  cat(nrow(data),"get_bootstrap 3\n")
  return(list(fit_boot_distrib,fit_boot_Bca,fit_boot_Bca_2))
}


```

## Run bootstrapping

Muuttujat regression_data-taulukossa (regression_data = CRP-NA:t poistettu):

ID
log_ferritin
Age
log_CRP
BMI
donation_count
log_last_donation
Cohort
iron_deficiency
Smoking
iron_complience
Group
donation_count_2
Menstruation

Näistä log_CRP ja iron_complience  puuttuvat NL-kohorteista, kiinnosti vaan nähdä miltä ne näyttää. Jos ne menee hankalaksi niitä ei siis tarvii saada niitä toimimaan :) 



```{r}
get_bootstrap_coeffs <- function(regression_data, current_group, #men, pre or post
                                 nb_boot ) #start at  100, then build up to around 1000
{
  
  
  if (current_group == "Men") {
    test_data_std <- regression_data %>% 
      filter(Group == "Men") %>% 
      select(
      #By commenting out variables from this list you change what is included in the bootsrapped models
      #ID,
      #log_ferritin,
      Age,
      log_CRP,
      BMI,
      #donation_count,
      log_last_donation,
      Cohort,           
      iron_deficiency,
      Smoking,
      iron_complience,
      #Group,
      donation_count_2#,
      #Menstruation
    )
  }
  
  if (current_group == "Pre_menopause_women") {
    test_data_std <- regression_data %>% 
      filter(Group == "Pre_menopause_women") %>% 
      select(
      #By commenting out variables from this list you change what is included in the bootsrapped models
      #ID,
      #log_ferritin,
      Age,
      log_CRP,
      BMI,
      #donation_count,
      log_last_donation,
      Cohort,           
      iron_deficiency,
      Smoking,
      iron_complience,
      #Group,
      donation_count_2,
      Menstruation
    )  
  }
  
  if (current_group == "Post_menopause_women") {
    test_data_std <- regression_data %>%
      filter(Group == "Post_menopause_women") %>%
      select(
      #By commenting out variables from this list you change what is included in the bootsrapped models
      #ID,
      #log_ferritin,
      Age,
      log_CRP,
      BMI,
      #donation_count,
      log_last_donation,
      Cohort,           
      iron_deficiency,
      Smoking,
      iron_complience,
      #Group,
      donation_count_2#,
      #Menstruation
    )
  }
  cat(nrow(test_data_std),current_group,"get_bootstrap_coeffs 1\n")
  
  # tässä itse bootstrapping
  result <- test_data_std %>% 
    get_bootstrap(nb_boot = nb_boot)
  
  cat(nrow(test_data_std),current_group,"get_bootstrap_coeffs 2\n")
  bootstrap_distrib <- result[[1]] %>% 
    gather(key = regressor, value = coefficient) %>% 
    filter(regressor != "(Intercept)") %>% 
    mutate(group = current_group)
  cat(nrow(test_data_std),current_group,"get_bootstrap_coeffs 3\n")
  Bca_CI  <- result[[2]] %>% mutate(group = current_group)
  Bca_CI_2  <- result[[3]] %>% mutate(group = current_group)
  cat(nrow(test_data_std),current_group,"get_bootstrap_coeffs 4\n")
  return(list(bootstrap_distrib,Bca_CI,Bca_CI_2))
}




#get_from_file_ferr = TRUE



set.seed(125)
#group_str = "Pre_menopause_women"

for(group_str in c("Pre_menopause_women","Men","Post_menopause_women")){
#  for(group_str in c("Pre_menopause_women")) {
#for(group_str in c("Post_menopause_women","Pre_menopause_women")){
  output_file_distrib = paste0("../results/bootstraps/exp_coeff_boot_distrib_seed_125_ctr_strata_final_",group_str,".rdata")
  output_file_Bca = paste0("../results/bootstraps/exp_coeff_boot_Bca_seed_125_ctr_strata_final_",group_str,".rdata")
  output_file_Bca_2 = paste0("../results/bootstraps/exp_coeff_boot_Bca_2_seed_125_ctr_strata_final_",group_str,".rdata")
  get_from_file_ferr <- all(file.exists(output_file_distrib),file.exists(output_file_Bca),file.exists(output_file_Bca_2) )
  if (get_from_file_ferr){
    cat("Loading boots from file for group ",group_str,"\n")
    stuff <- list()
    load(file=output_file_distrib)
    load(file=output_file_Bca)
    load(file=output_file_Bca_2)
    stuff[[1]] <- temp1  
    stuff[[2]] <- temp2
    stuff[[3]] <- temp3
  }else {
    stuff <- get_bootstrap_coeffs(regression_data, group_str, nb_boot = 10000)
    #         stuff <- get_bootstrap_coeffs(regression_data, group_str, nb_boot = 1000)
    temp1 <- stuff[[1]] 
    temp2 <- stuff[[2]] 
    temp3 <- stuff[[3]] 
    save(temp1,file=output_file_distrib) 
    save(temp2,file=output_file_Bca)
    save(temp3,file=output_file_Bca_2)
  }
  
  if(!exists("bootstrap_distrib")){
    cat("initializing bootstrap_distrib, bootstrap_Bca_CI_2 & bootstrap_Bca_CI\n")

    bootstrap_distrib <- stuff[[1]]
    cat("nrow(bootstrap_distrib)",nrow(bootstrap_distrib),"\n")
    bootstrap_Bca_CI <- stuff[[2]]
    cat("nrow(bootstrap_Bca_CI)",nrow(bootstrap_Bca_CI),"\n")
    bootstrap_Bca_CI_2 <- stuff[[3]]
    cat("nrow(bootstrap_Bca_CI_2)",nrow(bootstrap_Bca_CI_2),"\n")
  }else
  {
    cat("bootstrap_distrib & bootstrap_Bca_CI already exist\n")
    bootstrap_distrib<-bind_rows(bootstrap_distrib, stuff[[1]])
    bootstrap_Bca_CI<-bind_rows(bootstrap_Bca_CI, stuff[[2]])
    bootstrap_Bca_CI_2<-bind_rows(bootstrap_Bca_CI_2, stuff[[3]])
  }
  
}


#Typical errors:
# Error in bca.ci(boot.out, conf, index[1L], L = L, t = t.o, t0 = t0.o, : estimated adjustment 'a' is NA
#   https://stats.stackexchange.com/questions/37918/why-is-the-error-estimated-adjustment-a-is-na-generated-from-r-boot-package
#   try to increase nb_boot
# Error in t.star[r, ] <- res[[r]] : 
#   number of items to replace is not a multiple of replacement length
#   can sometimes be fixed by just adding bootstrap, but sometimes requires dropping out a variable

```


```{r}
bootstrap_Bca_CI %>%  group_by(group) %>% count()

```

# Forest plots

```{r}
# "\U0394 BMI"

regressor_values <- c(
 "age" = "Age",
   "BMI_diff" = "BMI difference",
   "donation_frequency" = "Donation frequency",
   "Hb_beginning" = "Initial Hemoglobin",
   "log_Ferritin_beginning" ="Initial Ferritin",
   "starting_weight" ="Initial weight",
   "health_very_good" ="Initial health rating \n Very good vs Satisfactory/Good",
    "health_excellent" = "Initial health rating \n Excellent vs Satisfactory/Good",
 
 
 "cholesterol_dTRUE" = "Elevated cholesterol detected",
"blood_sugar_dTRUE" = "Elevated blood sugar detected", 
"HBP_dTRUE" = "High blood pressure detected",
"low_hb_hist_dTRUE" = "Low haemoglobin detected",
"iron_supp_dStable" = "FRCBS iron supplementation stable", 
"iron_supp_dIncreased" = "FRCBS iron supplementation increased",
"vita_multi_dStable" = "Frequency multivitamine stable",
"vita_multi_dIncreased" = "Frequency multivitamine increased",
#"vita_iron_dStable" = "Frequency iron supplements stable",
#"vita_iron_dIncreased" = "Frequency iron supplements increased",
"anti_inf_dStable" = "Frequency of anti-inflammatory stable",
"anti_inf_dIncreased" = "Frequency of anti-inflammatory increased",
"freq_red_meat_dStable" = "Frequency of red meat stable",
"freq_red_meat_dIncreased" = "Frequency of red meat increased",
"freq_fruits_dStable" = "Frequency of fruits stable",
"freq_fruits_dIncreased" = "Frequency of fruits increased",
"freq_juices_dStable" = "Frequency of juices stable",
"freq_juices_dIncreased" = "Frequency of juices increased",
"freq_wholem_dStable" = "Frequency of whole milk stable",
"freq_wholem_dIncreased" = "Frequency of whole milk increased",
"freq_milk_dStable" = "Frequency of milk stable",
"freq_milk_dIncreased" = "Frequency of milk increased",
"freq_cof_dStable" = "Frequency of coffee stable",
"freq_cof_dIncreased"  = "Frequency of coffee increased",
"freq_tea_dStable" = "Frequency of tea stable",
"freq_tea_dIncreased" = "Frequency of tea increased",
"freq_alc_dStable" = "Frequency of alcohol stable",
"freq_alc_dIncreased" = "Frequency of alcohol increased",
"phys_inf_dStable" = "Physical symptoms inferred stable",
"phys_inf_dIncreased" = "Physical symptoms inferred increased",
"phys_cond_dStable" = "Physical condition stable",
"phys_cond_dIncreased" = "Physical condition increased",
"act_heavy_dStable" = "Ability for heavy activity stable",
"act_heavy_dIncreased" = "Ability for heavy activity increased",
"day_act_dStable"  = "Typical day activity stable",
"day_act_dIncreased" = "Typical day activity increased",
"h_act_light_dStable"  = "Hours for light activity stable",
"h_act_light_dIncreased"  = "Hours for light activity increased",
"sleep_enough_dStable" = "Enough sleep stable",
"sleep_enough_dIncreased"  = "Enough sleep increased",
"sleep_tired_dStable" = "Tired during day stable",
"sleep_tired_dIncreased" = "Tired during day increased",
"freq_calm_dStable" = "Frequency of feeling calm stable",
"freq_calm_dIncreased" =  "Frequency of feeling calm increase",
"life_qualgood" = "Quality of life good",
"life_qualvery_good" = "Quality of life very good",
"employmentTRUE" = "Full time employed"
)
```



```{r}
temp <-
  tidy_logit_men %>% 
  bind_rows(tidy_logit_post_menop_women) %>% 
  bind_rows(tidy_logit_pre_menop_women) %>% 
  dplyr::select(group, term, OR, p.value) %>% 
  rename(regressor = term) %>% 
  inner_join(bootstrap_Bca_CI,
             by =  c("regressor", "group")) %>% 
  filter(regressor != "(Intercept)") %>% 
  mutate(Bca_inf = exp(Bca_inf),
         Bca_sup = exp(Bca_sup),
         is_sig = p.value < 0.05) %>%
  mutate(
    regressor = plyr::revalue(regressor, regressor_values),
    regressor = ordered(regressor,
                        levels =  as.character(regressor_values)),
    regressor = fct_rev(regressor),
    group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women",
                      group == "Post_menopause_women" ~ "Post-menopausal women",
                      TRUE ~ "Men"),
    group = ordered(group, levels = c(  "Pre-menopausal women",  "Post-menopausal women", "Men"  )
    )
  )

temp2 <-
  tidy_logit_men %>% 
  bind_rows(tidy_logit_post_menop_women) %>% 
  bind_rows(tidy_logit_pre_menop_women) %>% 
  dplyr::select(group, term, OR, p.value) %>% 
  rename(regressor = term) %>% 
  inner_join(bootstrap_Bca_CI_2,
             by =  c("regressor", "group")) %>% 
  filter(regressor != "(Intercept)") %>% 
  mutate(Bca_inf = exp(Bca_inf),
         Bca_sup = exp(Bca_sup),
         is_sig = p.value < 0.05) %>%
  mutate(
    regressor = plyr::revalue(regressor, regressor_values),
    regressor = ordered(regressor,
                        levels =  as.character(regressor_values)),
    regressor = fct_rev(regressor),
    group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women",
                      group == "Post_menopause_women" ~ "Post-menopausal women",
                      TRUE ~ "Men"),
    group = ordered(group, levels = c(  "Pre-menopausal women",  "Post-menopausal women", "Men"  )
    )
  )
    

```

```{r}
bootstrap_Bca_CI <- temp
plotdata <- bootstrap_distrib %>% 
  mutate(regressor = plyr::revalue(regressor, regressor_values),
  regressor = ordered(regressor,
                  levels = as.character(regressor_values)),
  regressor = fct_rev(regressor),
  group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women",
    group == "Post_menopause_women" ~ "Post-menopausal women",
    TRUE ~ "Men"),
  group = ordered(group, levels = c( "Men",  "Post-menopausal women", "Pre-menopausal women" ) )) %>% 
  mutate(OR = exp(coefficient))

bootstrap_Bca_CI_2 <- temp2

file <- "../results/exp_bootstrap_Bca_CI.csv"
temp <- bootstrap_Bca_CI %>% mutate_if(is.numeric, round, digits = 3) %>% dplyr::select(regressor,group,p.value,OR,Bca_inf,Bca_sup) %>% arrange(desc(regressor),group)
write.csv(temp ,file = file,row.names = FALSE)

```

#Small plot of filtered variables for main paper
```{r fig1, fig.height = 10, fig.width = 15}
#set the forest plot OR scale lims
orlimR <- 1000
orlimL <- 1/orlimR
pointrange2size <- 1.4
#Filter for those that are sig in some:
bootstrap_Bca_CI_filtered <- bootstrap_Bca_CI %>% group_by(regressor) %>% mutate(some_sig= any(is_sig)) %>% filter(some_sig == TRUE)
#Set values on axes limits to facilitate plotting
bootstrap_Bca_CI_filtered$Bca_sup[bootstrap_Bca_CI_filtered$Bca_sup > orlimR] <- orlimR
bootstrap_Bca_CI_filtered$Bca_sup[bootstrap_Bca_CI_filtered$Bca_sup == 0] <- orlimR
bootstrap_Bca_CI_filtered$Bca_inf[bootstrap_Bca_CI_filtered$Bca_inf < orlimL] <- orlimL

#Filter for those that are sig in some:
bootstrap_Bca_CI_filtered_2 <- bootstrap_Bca_CI_2 %>% group_by(regressor) %>% mutate(some_sig= any(is_sig)) %>% filter(some_sig == TRUE)
#Set values on axes limits to facilitate plotting
bootstrap_Bca_CI_filtered_2$Bca_sup[bootstrap_Bca_CI_filtered_2$Bca_sup > orlimR] <- orlimR
bootstrap_Bca_CI_filtered_2$Bca_sup[bootstrap_Bca_CI_filtered_2$Bca_sup == 0] <- orlimR
bootstrap_Bca_CI_filtered_2$Bca_inf[bootstrap_Bca_CI_filtered_2$Bca_inf < orlimL] <- orlimL


plotdata_filtered <- plotdata %>% filter(plotdata$regressor %in% bootstrap_Bca_CI_filtered$regressor) 


p  <- ggplot(aes(x = regressor, y = OR, color = group, fill = group),data = plotdata_filtered) +
  geom_hline(yintercept = 1 , color = "grey", linetype = "dotted",
             size = 1) +
  # geom_violin(alpha = 0.15, trim = TRUE, position = position_dodge(width = 0.8), size = 0.1) +
  geom_pointrange(data = bootstrap_Bca_CI_filtered,
                  aes(ymin=Bca_inf, ymax = Bca_sup, shape = is_sig, color = group, fill = group),
                  position = position_dodge(width = 0.7), size = 1, linetype = 1) +
  # geom_pointrange(data = bootstrap_Bca_CI_filtered_2,
  #                 aes(ymin=Bca_inf, ymax = Bca_sup, shape = is_sig, color = group, fill = group),
  #                 position = position_dodge(width = 0.7), size = pointrange2size) +
  # # geom_point(data = bootstrap_Bca_CI, 
  #            aes(shape = is_sig, fill = group),
  #            position = position_dodge(width = 0.8), size = 2) +
    scale_color_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                     limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
     scale_fill_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                     limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
 scale_shape_manual(values = c(1,16)) +
  scale_y_log10(limits=c(orlimL,orlimR)) +
  # scale_x_discrete(labels=parse(text=unique(bootstrap_Bca_CI$regressor)))
  # ylim(1/35, 35) +
  guides(shape = FALSE) +
  theme_classic() +
  # theme(legend.position = c(0.8,0.9),
  #       legend.title = element_blank(),
  #       legend.box = "vertical",
  #       legend.text = element_text(size = 20),
  #       plot.title =  element_text(size = 24,
  #                                 hjust = 0),
  #       panel.grid.minor.y = element_blank(),
  #       axis.line = element_line(colour="black"),
  #       axis.title.y = element_blank(),
  #       axis.title.x = element_text(size = 24),
  #       axis.text = element_text(size = 22),
  #       panel.grid.major.y = element_blank()) +
  coord_flip() +
  xlab("Robust standardized coefficient") +
  labs(title = "PREDICTORS OF WORSENED SELF-RATED HEALTH \nBETWEEN QUESTIONNAIRES",
       subtitle = "") +
   theme(axis.title = element_text(size = 14),
        strip.text =  element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.background = element_rect(fill = "white"),
        strip.background.x = element_blank(),
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(size = 14, angle = 30,hjust = 0.8),
        plot.title =  element_text(size = 16),
        plot.subtitle = element_text(size = 16,
                                     hjust = 0.5))

p

```

```{r}

p <- p +   labs(title = NULL,subtitle = NULL)
ggplot2::ggsave(filename = "../results/figures/exp_small_log_reg_results_b.pdf",
                plot= p,
# do not change these rather just scale the ready pdf later if you need.
       width = 35,
       height = 25,
       dpi = 600,
       units = "cm")

```


#Big plot of all variables for supplement
```{r fig2, fig.height = 40, fig.width = 15}

bootstrap_Bca_CI$Bca_sup[bootstrap_Bca_CI$Bca_sup > orlimR] <- orlimR
bootstrap_Bca_CI$Bca_sup[bootstrap_Bca_CI$Bca_sup == 0] <- orlimR
bootstrap_Bca_CI$Bca_inf[bootstrap_Bca_CI$Bca_inf < orlimL] <- orlimL

#Set values on axes limits to facilitate plotting
bootstrap_Bca_CI_2$Bca_sup[bootstrap_Bca_CI_2$Bca_sup > orlimR] <- orlimR
bootstrap_Bca_CI_2$Bca_sup[bootstrap_Bca_CI_2$Bca_sup == 0] <- orlimR
bootstrap_Bca_CI_2$Bca_inf[bootstrap_Bca_CI_2$Bca_inf < orlimL] <- orlimL


p2  <- ggplot(aes(x = regressor, y = OR, color = group, fill = group),data = plotdata) +
  geom_hline(yintercept = 1 , color = "grey", linetype = "dotted",
             size = 1) +
  # geom_violin(alpha = 0.15, trim = TRUE, position = position_dodge(width = 0.8), size = 0.1) +
  geom_pointrange(data = bootstrap_Bca_CI,
                  aes(ymin=Bca_inf, ymax = Bca_sup, shape = is_sig, color = group, fill = group),
                  position = position_dodge(width = 0.7), size = 1, linetype = 2) +
  geom_pointrange(data = bootstrap_Bca_CI_2,
                  aes(ymin=Bca_inf, ymax = Bca_sup, shape = is_sig, color = group, fill = group),
                  position = position_dodge(width = 0.7), size = pointrange2size) +
  # # geom_point(data = bootstrap_Bca_CI, 
  #            aes(shape = is_sig, fill = group),
  #            position = position_dodge(width = 0.8), size = 2) +
    scale_color_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                     limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
     scale_fill_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                     limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
 scale_shape_manual(values = c(1,16)) +
  scale_y_log10(limits=c(orlimL,orlimR)) +
  # scale_x_discrete(labels=parse(text=unique(bootstrap_Bca_CI$regressor)))
  # ylim(1/35, 35) +
  guides(shape = FALSE) +
  theme_classic() +
  # theme(legend.position = c(0.8,0.9),
  #       legend.title = element_blank(),
  #       legend.box = "vertical",
  #       legend.text = element_text(size = 20),
  #       plot.title =  element_text(size = 24,
  #                                 hjust = 0),
  #       panel.grid.minor.y = element_blank(),
  #       axis.line = element_line(colour="black"),
  #       axis.title.y = element_blank(),
  #       axis.title.x = element_text(size = 24),
  #       axis.text = element_text(size = 22),
  #       panel.grid.major.y = element_blank()) +
  coord_flip() +
  xlab("Robust standardized coefficient") +
  labs(title = "PREDICTORS OF WORSENED SELF-RATED HEALTH \nBETWEEN QUESTIONNAIRES",
       subtitle = "") +
   theme(axis.title = element_text(size = 14),
        strip.text =  element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.background = element_rect(fill = "white"),
        strip.background.x = element_blank(),
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(size = 14, angle = 30,hjust = 0.8),
        plot.title =  element_text(size = 16),
        plot.subtitle = element_text(size = 16,
                                     hjust = 0.5))

p2
```



```{r}
p2 <- p2 +   labs(title = NULL,subtitle = NULL)
ggplot2::ggsave(filename = "../results/figures/exp_big_log_reg_results_b.pdf",
                plot= p2,
       width = 35,
       height = 55,
       dpi = 600,
       units = "cm")

```


#Separate plots for individual variables

```{r}
fig.width=15
fig.height=13
#Load questionnaire data
load(file = "../data/final_data.rdata") 
final_data <- data

```

## Change in physical condition color
```{r}

pdata <- final_data %>% 
  dplyr::select(donor, group, health, questionnaire) %>% 
  drop_na(group) %>% 
  filter( group != "Women_pre_menop_no_mens") %>% 
  spread(key = questionnaire, value = health) %>% 
  filter(First != "NA" & Second != "NA") %>%  
  mutate(Difference = case_when(First < Second ~ "Improved",
                                First == Second ~ "Stable",
                                TRUE ~ "Worsened")) %>% 
  inner_join(regression_data, by = c("donor", "group")) %>% 
  mutate(group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal \n women",
                           group == "Post_menopause_women" ~ "Post-menopausal \n women",
                           TRUE ~ "Men"),
         group = ordered(group, levels = c(  "Men", "Post-menopausal \n women", "Pre-menopausal \n women") ))
 
p <- ggplot(aes(x= phys_cond_d , fill = group), data=pdata) +
  # geom_hline(yintercept = 15, color = "red", linetype = 3)
  geom_bar()+
  #geom_boxplot(outlier.alpha = 0) +
  #geom_beeswarm(alpha = 0.5, shape = 1) +
  # scale_y_log10() +
  facet_grid(group ~ Difference,scales="free_y") +
  scale_fill_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                    limits = c("Pre-menopausal \n women", "Post-menopausal \n women","Men" )) +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("Evolution of self-rated health") +
  xlab("Change in physical condition") +
  ylab("Count") +
  geom_text(stat='count', aes(label=..count..),color="black", vjust=1)  
#scale_y_log10() 



ggplot2::ggsave(plot=p,filename = "../results/figures/exp_phys_cond_c.pdf",
       width = fig.width,
       height = fig.height,
       dpi = 600,
       units = "cm")

p
```




## Quality of life
```{r}

p <- ggplot(aes(x= life_qual), data=pdata) +
  # geom_hline(yintercept = 15, color = "red", linetype = 3)
  geom_bar(aes(fill=group))+
  #geom_boxplot(outlier.alpha = 0) +
  #geom_beeswarm(alpha = 0.5, shape = 1) +
  # scale_y_log10() +
  facet_grid(group ~ Difference,scales="free_y") +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("Evolution of self-rated health")+
  xlab("Quality of life") +
  ylab("Count") +
  geom_text(stat='count', aes(label=..count..),color="black", vjust=1) +
  scale_fill_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c("Pre-menopausal \n women", "Post-menopausal \n women","Men" ))

ggplot2::ggsave(plot=p,filename = "../results/figures/exp_life_qual_c.pdf",
       width = fig.width,
       height = fig.height,
       dpi = 600,
       units = "cm")

p
```


