---
title: "Low ferritin subgroup analysis"   
author: "Sofie Ekroos"
date: "21.11.2021"
output: pdf_document

---

# Summary

This code is based on Mikko Arvas' Bayesian model for iron deficiency, "Low ferritin model" (https://github.com/FRCBS/iron_measurement_comparisons/blob/master/src/low_ferritin_model_bayes.Rmd). In this analysis we analyse iron deficiency risk factors in pre-menopausal women. This subgroup analysis also allows us to compare the model to premenopausal women in Lobier et al. 2019 (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6692066/).

Version 1 includes the imputed blood donation observations for the general cohorts. These were imputed as random intervals (3-6 years, see low ferritin data code). 

All variables:

Age  
BMI  
Smoking
iron_complience (only blood donor cohort)
donation_count_2 (only blood donor cohort)
donation_count (only blood donor cohort)
log_last_donation
log_CRP 
Region (not in blood donor cohort)
Weigh
Menstruation
RedMeat_n (not in h2000 cohort)
PreviousChildbirth

```{r setup, , message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
# echo "rmarkdown::render('version1_low_ferritin_model_bayes_subgroup_analysis.Rmd', clean=TRUE,output_format='pdf_document',output_file='../results/bmodels_subgroup/version1_low_ferritin_model_bayes_subgroup_analysis/version1_low_ferritin_model_bayes_subgroup_analysis.pdf')" | R --slave

library(tidyverse)
library(brms)
library(bayesplot)
```



PART 2: BAYESIAN LOGISTIC REGRESSION

# Data loading and preparation 

Load data

```{r}
load("../data/ID_data_regression_cohorts.rdata")
all_cohorts <- regression_cohorts %>% filter(Group == "Pre_menopause_women")

all_cohorts <- all_cohorts %>% 
  filter(!is.na(CRP)) %>% 
  filter(! CRP >= 30)
```



# Build subgroups of pre-menopausal women (three cohorts: FinDonor, FinRisk97, Health2000) 

## Data
```{r}
# FinDonor
regression_data_blood <- all_cohorts %>% 
  filter(Cohort == "FINDONOR") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, iron_deficiency, Smoking, iron_complience, Region, Weight, Menstruation, RedMeat_n, PreviousChildbirth) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         Weight = scale(Weight, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# FinRisk97
regression_data_fr1997 <- all_cohorts %>% 
  filter(Cohort == "FINRISK97") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, iron_deficiency, Smoking, iron_complience, Region, Weight, Menstruation, RedMeat_n, PreviousChildbirth) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         Weight = scale(Weight, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 

# Health2000 
regression_data_h2000 <- all_cohorts %>% 
  filter(Cohort == "HEALTH00") %>% 
  dplyr::select(ID, log_ferritin, Age, log_CRP, BMI, donation_count, log_last_donation, iron_deficiency, Smoking, iron_complience, Region, Weight, Menstruation, RedMeat_n, PreviousChildbirth) %>%
  mutate(Age = scale(Age, scale = FALSE)[,1],
         Weight = scale(Weight, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         donation_count = scale(donation_count, scale = FALSE)[,1],
         log_last_donation = scale(log_last_donation, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         donation_count_2 = donation_count^2) 


```

## Models 

### FinDonor

The Region-variable only has one observation (HYKS) in FinDonor, since all blood donors are from Helsinki or Espoo. As the code will not run as long as Region is included, it was removed for the FinDonor cohort.

```{r blood}
file <- "../results/bmodels_subgroup/version1_low_ferritin_model_bayes_subgroup_analysis/logit_pre_blood_b"

logistic_pre_blood_b <- brm(iron_deficiency ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        iron_complience + 
                        donation_count_2 + 
                        donation_count +
                        log_last_donation + 
                        log_CRP +
                        #Region + 
                        Weight +
                        Menstruation + 
                        RedMeat_n +
                        PreviousChildbirth,
                      data = regression_data_blood, 
                      family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(logistic_pre_blood_b)

```

FinRisk97

```{r fr1997}
file <- "../results/bmodels_subgroup/version1_low_ferritin_model_bayes_subgroup_analysis/logit_pre_fr1997_b"

logistic_pre_fr1997_b <- brm(iron_deficiency ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        #iron_complience + 
                        #donation_count_2 + 
                        #donation_count + 
                        log_last_donation + 
                        log_CRP +
                        Region + 
                        Weight +
                        Menstruation + 
                        RedMeat_n +
                        PreviousChildbirth, 
                      data = regression_data_fr1997, 
                      family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(logistic_pre_fr1997_b)
```

Hmm, there are problems with this cohort. Warning message: 

There were 1481 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10. See
http://mc-stan.org/misc/warnings.html#maximum-treedepth-exceededExamine the pairs() plot to diagnose sampling problems
The largest R-hat is 1.45, indicating chains have not mixed.
Running the chains for more iterations may help. See
http://mc-stan.org/misc/warnings.html#r-hatBulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.
Running the chains for more iterations may help. See
http://mc-stan.org/misc/warnings.html#bulk-essTail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable.
Running the chains for more iterations may help. See
http://mc-stan.org/misc/warnings.html#tail-essParts of the model have not converged (some Rhats are > 1.05). Be careful when analysing the results! We recommend running more iterations and/or setting stronger priors. 

This was solved by removing the variables iron_complience, donation_count and donation_count_2 (diagnostics in the later stage of this analysis also suggested these variables don't work). Donation count were imputed as a randomized number between 3-6 years in the general population cohorts (see low_ferritin_data) and donations in the past two years as 0 as we didn't have this information. We also don't have data on iron supplementation in the general population cohorts so all observations were imputed as 0.

## Health00

```{r h2000}
file <- "../results/bmodels_subgroup/version1_low_ferritin_model_bayes_subgroup_analysis/logit_pre_h2000_b"

logistic_pre_h2000_b <- brm(iron_deficiency ~ 
                        Age + 
                        BMI + 
                        Smoking + 
                        #iron_complience + 
                        #donation_count_2 +
                        #donation_count + 
                        log_last_donation + 
                        log_CRP +
                        Region + 
                        Weight +
                        Menstruation + 
                        #RedMeat_n +
                        PreviousChildbirth, 
                      data = regression_data_h2000, 
                      family = bernoulli(),
                      file=file, 
                      cores = 4,
                      iter= 10000
)

summary(logistic_pre_h2000_b)
``` 

There are problems with this cohort as well, same warning message. Solved this as explained for FinRisk97. Red meat consumption was also removed as the Health2000 didn't contain information on this.

## Diagnostics

### Trace

```{r , message=FALSE}
mcmc_plot(logistic_pre_blood_b, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(logistic_pre_fr1997_b, type = "trace")
```

```{r , message=FALSE}
mcmc_plot(logistic_pre_h2000_b, type = "trace")
```

Chains do not seem to get stuck now that the variables iron_complience and donation_count_2 were removed in both general population cohorts. 

### ACF bars

```{r}
mcmc_plot(logistic_pre_blood_b, type = "acf_bar")
```


```{r}
mcmc_plot(logistic_pre_fr1997_b, type = "acf_bar")
```


```{r}
mcmc_plot(logistic_pre_h2000_b, type = "acf_bar")
```

Autocorrelation drops in a satisfactory way.

### Rhat

```{r}
mcmc_plot(logistic_pre_blood_b, type = "rhat")
```

```{r}
mcmc_plot(logistic_pre_fr1997_b, type = "rhat")
```

```{r}
mcmc_plot(logistic_pre_h2000_b, type = "rhat")
```
https://mc-stan.org/rstan/reference/Rhat.html


"The Rhat function produces R-hat convergence diagnostic, which compares the between- and within-chain estimates for model parameters and other univariate quantities of interest. If chains have not mixed well (ie, the between- and within-chain estimates don't agree), R-hat is larger than 1. We recommend running at least four chains by default and only using the sample if R-hat is less than 1.05."

Rhat looks good as well.

## Forest plots

```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(logistic_pre_blood_b, 
         type = "areas",
         prob = 0.95,
         transformations = "exp") +
  geom_vline(xintercept = 1, color = "grey")

p <- p + scale_x_log10()
#Ferritin needs to be removed
# plot.OR<-  tidy_logit_pre_menop_women %>%filter(!term=="log_Ferritin_beginning") %>%  dplyr::select(OR) %>% mutate(y=rev(2:(nrow(tidy_logit_pre_menop_women)))) %>% bind_rows(tibble(OR=NA,y=1 ))
# p <- p + geom_point(aes(y=y,x=OR),data=plot.OR, col="red")
p
```

```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(logistic_pre_fr1997_b, 
         type = "areas",
         prob=0.95,
         prob_est="median",
         transformations = "exp"
        ) +
  geom_vline(xintercept = 1, color = "grey")

p <- p + scale_x_log10()
p
```

```{r , message=FALSE, warning=FALSE}
p <- mcmc_plot(logistic_pre_h2000_b, 
         type = "areas",
         prob = 0.95,
         transformations = "exp") +
  geom_vline(xintercept = 1, color = "grey")

p <- p + scale_x_log10()
#Ferritin needs to be removed
# plot.OR<-  tidy_logit_pre_menop_women %>%filter(!term=="log_Ferritin_beginning") %>%  dplyr::select(OR) %>% mutate(y=rev(2:(nrow(tidy_logit_pre_menop_women)))) %>% bind_rows(tibble(OR=NA,y=1 ))
# p <- p + geom_point(aes(y=y,x=OR),data=plot.OR, col="red")
p
```

### All groups together

```{r}
regressor_values <- c(
 "Age" = "Age (5 years)",
   "log_CRP" = "CRP (log transformed)",
   "BMI" = "BMI",
   "log_last_donation" = "Days since last donation (log, /log(2))",
   "Smokingyes" ="Smoker",
    "iron_complience" = "Iron supplementation",
    "donation_count_2" = "Donations in last 2 years (quadratic)",
    "Menstruationregular_period" = "Regular periods",
    "Weight" = "Weight (10 kg)",
    "Region" = "Region",
    "RedMeat_n" = "Red meat",
    "PreviousChildbirth" = "Given birth in the past",
    "RegionTYKS" = "Region (Turku university hospital)",
    "RegionTAYS" = "Region (Tampere university hospital)",
    "RegionOYS" = "Region (Oulu university hospital)",
    "RegionKYS" = "Region (Kuopio university hospital)",
    "donation_count" = "Donations in last 2 years"
 
)
```

```{r }

intervals_pre_blood <- mcmc_intervals_data(
  logistic_pre_blood_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median",
  transformations = "exp"
  ) %>% mutate(Cohort='FINDONOR')

intervals_pre_fr1997 <- mcmc_intervals_data(
  logistic_pre_fr1997_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median",
  transformations = "exp"
  ) %>% mutate(Cohort='FINRISK97') %>% 
  arrange(m) # how to order the variables?

intervals_pre_h2000 <- mcmc_intervals_data(
  logistic_pre_h2000_b,
  prob=0.9,
  prob_outer = 0.95,
  point_est = "median",
  transformations = "exp"
  ) %>% mutate(Cohort='HEALTH00')

intervals_combined <- rbind(intervals_pre_blood,intervals_pre_fr1997,intervals_pre_h2000) %>% 
  filter(! (parameter == 'exp(b_Intercept)' | parameter == 'exp(lp__)')) %>% 
  mutate(parameter = gsub("\\)","",gsub("exp\\(b_","",parameter))) %>% 
  mutate(regressor = plyr::revalue(parameter, regressor_values)) %>% 
  select(-parameter)
  #arrange(m)


```

```{r , fig1, fig.height = 10, fig.width = 15 ,warning=FALSE}

orlower <- 1/10
orupper <- 10

# Truncate ORs for plotting
plotdata <- intervals_combined %>% 
  mutate(
    ll = case_when(
      ll < orlower ~ orlower,
      TRUE ~ ll
    ),
    hh = case_when(
      hh > orupper ~ orupper,
      TRUE ~ hh
    )
  )

#theme_set(bayesplot::theme_default())
pos <- position_nudge(y=as.numeric(as.factor(plotdata$Cohort))/5 -0.1) 

p <- ggplot(plotdata, aes(x = m, y = regressor, color = Cohort)) + 
  geom_point(position = pos,
             size=4) +
  geom_linerange(aes(xmin = ll, xmax = hh), 
                 position = pos,
                 size = 1, 
                 linetype = 1
                 ) +
  scale_color_manual(values = c("#84b56f", "#88b5bc", "#a2c4c9"),
                     limits = c("FINDONOR",  "FINRISK97","HEALTH00")) +
  scale_fill_manual(values = c("#84b56f", "#88b5bc", "#a2c4c9"),
                    limits = c("FINDONOR",  "FINRISK97","HEALTH00")) +
  scale_shape_manual(values = c(1,16)) +
  scale_x_log10(limits=c(orlower,orupper))  + #how to get the lines that drop out of the visible?  
  geom_vline(xintercept = 1 , color = "grey", linetype = "dotted",
             size = 1) +
  theme(axis.title = element_text(size = 14),
        strip.text =  element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.background = element_rect(fill = "white"),
        strip.background.x = element_blank(),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14)
  ) +
  guides(shape = "none") +
  theme_classic() +
  ylab("Coefficient") 

p

```

Compared to Lobier et al. 2019 CRP has a noticeably larger effect. This can be explained by the larger data set we have in this analysis. The blood donation variables that were imputed do not seem to distort the results, so for the purpose of merging our data we can keep the variable.  

```{r}
ggplot2::ggsave(filename = "../results/figures/subgroup_analysis/version1_forest_pre_women_all_cohorts_bayes.pdf",
                plot = p,
                width = 35,
                height = 25,
                dpi = 600,
                units = "cm")

# save ORs for comparison with other models 
OR_logistic_subgroup_pre <- plotdata
save(OR_logistic_subgroup_pre, file = "../data/OR_low_ferritin_model_subgroup_analysis.rdata")
```
